---
title: "DESeq2"
author: "allyson_demerlis"
date: "2024-11-25"
output: html_document
---

# Installing and loading packages
```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install(version = "3.15")
# BiocManager::install("DESeq2",dependencies=T)
# BiocManager::install("BiocParallel")
#BiocManager::install("arrayQualityMetrics",dependencies=T)  # requires Xquartz, xquartz.org

# install.packages("pheatmap")
# install.packages("VennDiagram")
# install.packages("gplots")
# install.packages("vegan")
# install.packages("plotrix")
# install.packages("ape")
# install.packages("ggplot2")

library(DESeq2)
library(tidyverse)
library(ape)
library(vegan)
library(Biobase)
library(arrayQualityMetrics)
```

# Read in counts matrix and initial tidying
```{r}
#read in counts
counts = readxl::read_xlsx("1_bioinformatics/allcounts_pcli.xlsx", sheet = "sym") #alignment to Avila-Magana et al. 2021 transcriptome 

column_to_rownames(counts, var ="...1") -> counts

colnames(counts) # sample IDs (i.e. Pcli-003). You need to keep track of column names and row names of data frames throughout this analysis because they MUST MATCH 
rownames(counts) #gene names (but not actual gene names, arbitrary gene names. To get "actual" gene names, need annotation files)

# how many genes we have total?
nrow(counts) #26253
ncol(counts) #48 samples

# how does the data look? 
head(counts)

# filtering out low-count genes
keep <- rowSums(counts) >= 10
countData <- counts[keep,]
nrow(countData) #13507
ncol(countData) #48
#write.csv(countData, file = "results_csv/Pcli_countdata.csv")
```


## import and tidy design file
```{r}
# importing a design .csv file
design = read.csv("../treatment_metadata.csv", head=TRUE)

design %>% 
  select(Species:Treatment) %>% 
  filter(Species == "Pcli") -> design

design %>% 
  select(!Species) -> design

column_to_rownames(design, var="Sample_ID") -> design
design$Genotype <- as.factor(design$Genotype)
design$Treatment <- as.factor(design$Treatment)

# check if order is the same
all(rownames(design) == colnames(countData))
#TRUE
```


# DESeq2 full model design
```{r}
#when making dds formula, the order of variables in the design matters. The design indicates how to model the samples, (here: design = ~batch + condition), that we want to measure the effect of the condition, controlling for batch differences. The two factor variables batch and condition should be columns of coldata.

dds = DESeqDataSetFromMatrix(countData=countData, colData=design, design= ~Genotype + Treatment) # following host design
dds=estimateSizeFactors(dds)
plot(sort(sizeFactors(dds))) 

totalCountshost=colSums(countData)
totalCountshost
barplot(totalCountshost)
as.data.frame(totalCountshost) %>% 
  arrange(totalCountshost)

min(totalCountshost)  #6605
max(totalCountshost) #386770
```

## variance stabilized transformation for visualization
```{r}
Vsd=varianceStabilizingTransformation(dds)
```

## Outlier detection using ArrayQualityMetrics
```{r}
e=ExpressionSet(assay(Vsd), AnnotatedDataFrame(as.data.frame(colData(Vsd))))

# running outlier detection
arrayQualityMetrics(e,intgroup=c("Treatment", "Genotype"),force=T)
# open the directory "arrayQualityMetrics report for e" in your working directory and open index.html
# Array metadata and outlier detection overview gives a report of all samples, and which are likely outliers according to the 3 methods tested.
#I typically remove the samples that violate *1 (distance between arrays).
# Figure 2 shows a bar plot of array-to-array distances and an outlier detection threshold based on your samples. 
#Samples above the threshold are considered outliers
# under Figure 3: Principal Components Analyses, look for any points far away from the rest of the sample cluster
# use the array number for removal in the following section

# if there were outliers:
outs=c(5, 10, 12, 23, 26, 37, 39, 41, 43, 44, 46) #these numbers were taken from the index.html report from arrayQualityMetrics Figure 2 and Figure 3
# number 5 in the array corresponds to Pcli-011, which was determined to be an outlier using PCA
# 23 corresponds to Pcli-066
# 39 corresponds to Pcli-114

countData=countData[,-outs]
Vsd=Vsd[,-outs]
design=design[-outs,]

# remaking model with outliers removed from dataset
dds = DESeqDataSetFromMatrix(countData=countData, colData=design, design=~Genotype + Treatment)

# save all these dataframes as an Rdata package so you don't need to rerun each time
save(dds,design,countData,Vsd,file="initial_fullddsdesigncountsVsdcounts.RData")
```


### PCA
```{r}
load("Rdata_files/symbiont/initial_fullddsdesigncountsVsdcounts.RData")

plot <- DESeq2::plotPCA(Vsd, intgroup=c("Genotype", "Treatment", "Sample.ID"))

Vsd$Sample.ID

plot + geom_label(aes(label = Sample.ID)) #sample Pcli-011 is an outlier

#re-run above code to remove it as an outlier from the data

DESeq2::plotPCA(Vsd, intgroup=c("Genotype", "Treatment", "Sample.ID")) + geom_label(aes(label = Sample.ID))

#Pcli-066 is also an outlier

DESeq2::plotPCA(Vsd, intgroup=c("Genotype", "Treatment", "Sample.ID")) + geom_label(aes(label = Sample.ID))

#Pcli-114 is also an outlier

DESeq2::plotPCA(Vsd, intgroup=c("Genotype", "Treatment")) 
# i think this is the best we're going to get
```

### PCoA
```{r}
ad.pcoa=pcoa(dist(t(assay(Vsd)),method="manhattan")/1000)
scores=ad.pcoa$vectors
ad.pcoa$values # the % variation explained by each axis is the Relative_eig column
conditions=design

#code by group
pdf("Symbiont_PCoA_Treatment_Genotype.pdf")
plot(scores[,1], scores[,2], pch=c(15,17,25)[as.numeric(as.factor(conditions$Genotype))], col=c("#00CCCC", "grey","#FF3333")[as.numeric(as.factor(conditions$Treatment))], xlab="PC1: 13%", ylab="PC2: 4%", main="Treatment")
ordispider(scores[,1:2], conditions$Treatment,col=c("#00CCCC", "grey", "#FF3333"))
legend("topright", legend=c("Control", "Initial", "Variable"), fill = c("#00CCCC", "grey", "#FF3333"), bty="n")
legend("topleft", legend=c("A", "B", "C"), pch=c(15,17,25), bty="n")


#analysis of variance 
ad=adonis2(t(assay(Vsd))~Genotype+Treatment,data=conditions,method="manhattan", permutations=1e6)
ad
as.data.frame(ad) %>% 
  write_csv("PERMANOVA_Pcli.csv")
```


## create vsd
```{r}
# generating normalized variance-stabilized data for PCoA, heatmaps, etc
vsd=assay(Vsd)
colnames(vsd)
# takes the sample IDs and factor levels from the design to create new column names for the dataframe
snames=paste(colnames(countData),design[,1], design[,4],sep="_")
snames

# renames the column names
colnames(vsd)=snames

save(vsd,design,file="Rdata_files/symbiont/vsd.RData")
```


# Run DESeq2 on full model
```{r}
# with multi-factor, multi-level design
load("Rdata_files/symbiont/initial_fullddsdesigncountsVsdcounts.RData")
library(DESeq2)
library(BiocParallel)

# Running full model for contrast statements
#dds = DESeqDataSetFromMatrix(countData=countData, colData=design, design=~ Genotype + Treatment)
rownames(design)
colnames(countData)
dds=DESeq(dds, parallel=TRUE)

# saving all models
save(dds,file="Rdata_files/symbiont/realModels_Pcli.RData")
```


# DESeq2 Contrasts

## import gene names
```{r}
load("Rdata_files/symbiont/realModels_Pcli.RData")
library(DESeq2)
library(tidyverse)

read.table(file = "1_bioinformatics/Breviolum_iso2geneName.tab",
                          sep = "\t",
                          quote="", fill=FALSE) %>%
  rename(gene = V1,
         annot = V2) -> iso2geneName
```

## Treatment
```{r}
# treatment
Treatment=results(dds) 
summary(Treatment, alpha = 0.05) 
resultsNames(dds)
#primary ones of interest: "Treatment_Variable_vs_Initial",  "Treatment_Control_vs_Initial",  "Treatment_Variable_vs_Control"
```

### Variable vs. Control
```{r}
# Variable vs. Control
Treatment_Variable_vs_Control=results(dds,contrast=c("Treatment","Variable","Control"))
summary(Treatment_Variable_vs_Control, alpha = 0.05)
degs_Treatment_Variable_vs_Control=row.names(Treatment_Variable_vs_Control)[Treatment_Variable_vs_Control$padj<0.05 & !(is.na(Treatment_Variable_vs_Control$padj))]
length(degs_Treatment_Variable_vs_Control) #0

degs_Treatment_Variable_vs_Control_L2FC1=row.names(Treatment_Variable_vs_Control)[Treatment_Variable_vs_Control$padj<0.05 & !(is.na(Treatment_Variable_vs_Control$padj)) & abs(Treatment_Variable_vs_Control$log2FoldChange)>1]
length(degs_Treatment_Variable_vs_Control_L2FC1) #0

Treatment_Variable_vs_Control %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange > 1) %>% 
  nrow() #0
Treatment_Variable_vs_Control %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange < -1) %>% 
  nrow() #0
```

### Variable vs. Initial
```{r}
load("Rdata_files/symbiont/realModels_Pcli.RData")
# Treated vs. Initial
Treatment_Variable_vs_Initial=results(dds,contrast=c("Treatment","Variable","Initial"))
summary(Treatment_Variable_vs_Initial, alpha = 0.05)
degs_Treatment_Variable_vs_Initial=row.names(Treatment_Variable_vs_Initial)[Treatment_Variable_vs_Initial$padj<0.05 & !(is.na(Treatment_Variable_vs_Initial$padj))]
length(degs_Treatment_Variable_vs_Initial) #1

degs_Treatment_Variable_vs_Initial_L2FC1=row.names(Treatment_Variable_vs_Initial)[Treatment_Variable_vs_Initial$padj<0.05 & !(is.na(Treatment_Variable_vs_Initial$padj)) & abs(Treatment_Variable_vs_Initial$log2FoldChange)>1]
length(degs_Treatment_Variable_vs_Initial_L2FC1) #1

Treatment_Variable_vs_Initial %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange > 1) %>% 
  nrow() #1
Treatment_Variable_vs_Initial %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange < -1) %>% 
  nrow() #0

Treatment_Variable_vs_Initial %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "gene") %>% 
  drop_na(padj) %>% 
  filter(padj<0.05 & abs(log2FoldChange) >1) %>% 
  full_join(., iso2geneName, by = "gene") %>% 
  drop_na(baseMean) %>% 
  select(gene, annot, baseMean:padj) %>% 
  write_csv("Symbiont_VariablevInitial_annotDGEs.csv")

Treatment_Variable_vs_Initial %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "gene") %>% 
  drop_na(padj) %>% 
  filter(padj<0.05 & abs(log2FoldChange) >1) %>% 
  full_join(., iso2geneName, by = "gene") %>% 
  drop_na(baseMean) %>% 
  select(gene, annot, baseMean:padj) %>% 
  mutate(Contrast = "Variable vs Initial") -> variable_vs_initial
```


### Control vs. Initial
```{r}
# Control vs. Initial
Treatment_Control_vs_Initial=results(dds,contrast=c("Treatment","Control","Initial"))
summary(Treatment_Control_vs_Initial, alpha = 0.05)
degs_Treatment_Control_vs_Initial=row.names(Treatment_Control_vs_Initial)[Treatment_Control_vs_Initial$padj<0.05 & !(is.na(Treatment_Control_vs_Initial$padj))]
length(degs_Treatment_Control_vs_Initial) #0

degs_Treatment_Control_vs_Initial_L2FC1=row.names(Treatment_Control_vs_Initial)[Treatment_Control_vs_Initial$padj<0.05 & !(is.na(Treatment_Control_vs_Initial$padj)) & abs(Treatment_Control_vs_Initial$log2FoldChange)>1]
length(degs_Treatment_Control_vs_Initial_L2FC1) #0


Treatment_Control_vs_Initial %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange > 1) %>% 
  nrow() #0
Treatment_Control_vs_Initial %>% 
  as.data.frame() %>% 
  filter(padj<0.05 & log2FoldChange < -1) %>% 
  nrow() #0
```


