---
title: "samplesizes"
author: "allyson_demerlis"
date: "2023-02-22"
output: html_document
---
#load dataset and packages
```{r}
metadata <- readxl::read_xlsx("metadata.xlsx", sheet = "tidy_data")

library(tidyverse)
library(janitor)
library(lubridate)
```

#beginning of experiment sample size (by species and genotype)
```{r}
metadata %>% 
  select(Species, ID, Colony) %>% 
  group_by(Species, Colony) %>% 
  summarise(count = n())
```
#beginning of experiment sample size after fragmenting mortality (by species, genotype, and treatment)
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment) %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n()) #treatment = NA means that number of fragments died before getting treatment assignment
```

# sample size going into temperature treatment
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment) %>% 
  drop_na(Treatment) %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n()) 
```

# sample size for CBASS (separated by CBASS temperature assignment)
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment, CBASS_tank) %>% 
  drop_na(Treatment) %>% 
  group_by(Species, Colony, Treatment, CBASS_tank) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x=CBASS_tank, y=count, fill = Treatment)) + geom_bar(stat="identity") + facet_wrap(~Species + Colony) + scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic()

metadata %>% 
  select(Species, ID, Colony, Treatment, CBASS_tank) %>% 
  drop_na(Treatment) %>% 
  group_by(CBASS_tank) %>% 
  summarise(count = n())
```


# sample size post-CBASS
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment, CBASS_recovery_tank) %>% 
  drop_na(Treatment, CBASS_recovery_tank) %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n()) #overall numbers went from 283 pre-cbass to 183 post-cbass

#how many of those came from the control CBASS tank (temp=28ºC)?
metadata %>% 
  select(Species, ID, Colony, Treatment, CBASS_tank, CBASS_recovery_tank) %>% 
  drop_na(Treatment, CBASS_recovery_tank) %>% 
  group_by(Species, Colony, Treatment, CBASS_tank) %>% 
  summarise(count = n()) %>% 
    ggplot(., aes(x=CBASS_tank, y=count, fill = Treatment)) + geom_bar(stat="identity") + facet_wrap(~Species + Colony) + scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic()
```

# sample size pre-slow burn heat-stress
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment, slowburn_heatstress_tank) %>% 
  drop_na(Treatment, slowburn_heatstress_tank) %>% 
  group_by(Species, Colony, Treatment, slowburn_heatstress_tank) %>% 
  summarise(count = n()) #sample size total = 163 corals going into slow burn (lost 20 more corals due to CBASS stress)

metadata %>% 
  select(Species, ID, Colony, Treatment, slowburn_heatstress_tank) %>% 
  drop_na(Treatment, slowburn_heatstress_tank) %>% 
  group_by( slowburn_heatstress_tank) %>% 
  summarise(count = n())
```
# sample size post-slow burn heat-stress
```{r}
metadata %>% 
  select(Species, ID, Colony, Treatment, heatstress_recovery_tank) %>% 
  drop_na(Treatment, heatstress_recovery_tank) %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n()) #19 corals made it to the end (or were never put in the slow burn because they weren't ready for it)
```




