---
title: "CBASS_DRC_modeling"
author: "allyson_demerlis"
date: "2023-12-22"
output: html_document
---

# Packages

```{r}
library(tidyverse)
library(plotrix)
library(glmmTMB)
library(emmeans)
library(lme4)
library(ggpubr)
library(drc) #for the function "drm" dose-response model
library(broom) #for the tidy function needed for dose response model
library(car)
library(rstatix)
library(rcompanion)
library(nlme)
library(multcomp)
library(MASS)
```

# Import and tidy data

```{r}
ipam_tidy_data <- read.csv("ipam_tidy_data.csv")

ipam_tidy_data %>% 
  dplyr::filter(Treatment_period == "CBASS") %>% 
  dplyr::select(!Treatment) -> ipam_tidy_CBASS

ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>% 
  dplyr::select(Puck, Colony, Treatment, Species) %>% 
  distinct() %>% #283 rows = one coral ID per treatment, no repeats
  full_join(ipam_tidy_CBASS, by = c("Puck", "Colony", "Species")) %>% 
  drop_na() %>% 
  dplyr::select(Puck:Species, fvfm, Tank:CBASS_temp) -> ipam_tidy_CBASS_treatments

ipam_tidy_CBASS_treatments %>% 
  dplyr::filter(fvfm < 0.75 & fvfm > 0) %>% #has to be in between 0 and 1
  dplyr::filter(Species  == "Acervicornis") %>% 
  mutate(CBASS_temp=as.factor(CBASS_temp)) -> Acer_ipam_CBASS
```

# Raw CBASS data + GLMMs

## 1. Raw CBASS Fv/Fm by species (boxplot)
```{r}
ipam_tidy_CBASS_treatments %>% 
  mutate(CBASS_temp = as.factor(CBASS_temp)) %>% 
  ggplot(., aes(x=CBASS_temp, y= fvfm, fill = Treatment)) + 
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  geom_boxplot() +
  facet_wrap(~Species) + 
  theme_classic()
```

## 2. Raw CBASS Fv/Fm by species (curve)
```{r}
ipam_tidy_CBASS_treatments %>%  
  mutate(CBASS_temp=as.numeric(CBASS_temp)) %>% 
ggplot(., aes(x=CBASS_temp, y = fvfm, color = Treatment)) +
  geom_point() +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  geom_smooth(formula = y ~ x, method = "loess", span = 0.9)  + 
  facet_wrap(~Species) +
  theme_classic() +
  labs(x="Temperature (ºC)", y = "Fv/Fm")
```

### Statistics 

#### GLM Acer (colony as random effect)
```{r}
ipam_tidy_CBASS_treatments %>% 
  dplyr::filter(fvfm < 0.75 & fvfm > 0) %>% #has to be in between 0 and 1
  dplyr::filter(Species  == "Acervicornis") %>% 
  mutate(CBASS_temp=as.factor(CBASS_temp)) -> Acer_ipam_CBASS

#model 1 = full model Treatment*time_point + (1|Colony) + (1|Tank)
glmm_AcerCBASS_full <- glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Colony) + (1|Tank), family=beta_family(link = "logit"), data=Acer_ipam_CBASS)

summary(glmm_AcerCBASS_full) 

qqnorm(resid(glmm_AcerCBASS_full))
qqline(resid(glmm_AcerCBASS_full))

boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$Treatment)
boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$CBASS_temp)
boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$Colony)
boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$Tank)

#model 2 = Treatment*time_point + (1|Colony)
AcerCBASS_GLMM_Colony <- glmmTMB::glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Colony), family=beta_family(link = "logit"), data=Acer_ipam_CBASS)
summary(AcerCBASS_GLMM_Colony)

#need to check normality and variance visually (boxplots of fixed effects only)
qqnorm(resid(AcerCBASS_GLMM_Colony))
qqline(resid(AcerCBASS_GLMM_Colony))

boxplot(resid(AcerCBASS_GLMM_Colony)~Acer_ipam_CBASS$Treatment) 
boxplot(resid(AcerCBASS_GLMM_Colony)~Acer_ipam_CBASS$CBASS_temp)
boxplot(resid(AcerCBASS_GLMM_Colony)~Acer_ipam_CBASS$Colony)

#model 3 = Treatment*Date + (1|Tank)
Acer_CBASS_GLMM_Tank <- glmmTMB::glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Tank), family=beta_family(link = "logit"), data=Acer_ipam_CBASS)
summary(Acer_CBASS_GLMM_Tank)

qqnorm(resid(Acer_CBASS_GLMM_Tank)) #looks great
qqline(resid(Acer_CBASS_GLMM_Tank))

boxplot(resid(Acer_CBASS_GLMM_Tank)~Acer_ipam_CBASS$Treatment) 
boxplot(resid(Acer_CBASS_GLMM_Tank)~Acer_ipam_CBASS$CBASS_temp)
boxplot(resid(Acer_CBASS_GLMM_Tank)~Acer_ipam_CBASS$Tank)

lrt(AcerCBASS_GLMM_Colony, glmm_AcerCBASS_full)

lrt(Acer_CBASS_GLMM_Tank, glmm_AcerCBASS_full)

# GLMM Acer Full: AIC =  -428.5 
# GLMM Acer Colony: AIC =  -430.5, LRT to full model p-value = 1
# GLMM Acer Tank: AIC =  -423.3, LRT to full model p-value = 0.007379927

#based on AIC and LRT p-values (lowest AIC is best, if p-value is significant that means there is a significant difference in the full model versus the reduced model) ...

#the model with just colony as a fixed effect is not significantly different than the full model. However, tank model is significantly different from full. If I'm interpreting this correctly, it means that the fixed effect of colony is not important to the model, but tank is important to the model. You don't want to have too many variables to try to explain variance.  But the AIC of the full model is better than the AIC of the tank model. I will move forward with the full model.
```


```{r}
Anova(glmm_AcerCBASS_full, type = "II")

Acer_CBASS_emm <- emmeans(glmm_AcerCBASS_full, specs = ~ Treatment*CBASS_temp)

as.data.frame(pairs(Acer_CBASS_emm)) -> Acer_CBASS_treatment_posthoc

list <- c("Control CBASS_temp28 - Variable CBASS_temp28", "Control CBASS_temp30 - Variable CBASS_temp30", "Control CBASS_temp32 - Variable CBASS_temp32","Control CBASS_temp33 - Variable CBASS_temp33","Control CBASS_temp34 - Variable CBASS_temp34" ,"Control CBASS_temp35 - Variable CBASS_temp35","Control CBASS_temp36 - Variable CBASS_temp36","Control CBASS_temp37 - Variable CBASS_temp37")
list <- as.vector(list)

Acer_CBASS_treatment_posthoc %>% 
  filter(contrast %in% list)
```

#### GLM Pcli (colony as random effect)

```{r}
ipam_tidy_CBASS_treatments %>% 
  dplyr::filter(fvfm < 0.75) %>% 
  dplyr::filter(Species  == "Pclivosa") %>% 
  mutate(CBASS_temp=as.factor(CBASS_temp)) -> Pcli_ipam_CBASS

#model 1 = full model Treatment*time_point + (1|Colony) + (1|Tank)
glmm_PcliCBASS_full <- glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Colony) + (1|Tank), family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)

summary(glmm_PcliCBASS_full) 

qqnorm(resid(glmm_PcliCBASS_full))
qqline(resid(glmm_PcliCBASS_full))

boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$Treatment)
boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$CBASS_temp)
boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$Colony)
boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$Tank)

#model 2 = Treatment*time_point + (1|Colony)
PcliCBASS_GLMM_Colony <- glmmTMB::glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Colony), family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)
summary(PcliCBASS_GLMM_Colony)

#need to check normality and variance visually (boxplots of fixed effects only)
qqnorm(resid(PcliCBASS_GLMM_Colony))
qqline(resid(PcliCBASS_GLMM_Colony))

boxplot(resid(PcliCBASS_GLMM_Colony)~Pcli_ipam_CBASS$Treatment) 
boxplot(resid(PcliCBASS_GLMM_Colony)~Pcli_ipam_CBASS$CBASS_temp)
boxplot(resid(PcliCBASS_GLMM_Colony)~Pcli_ipam_CBASS$Colony)

#model 3 = Treatment*Date + (1|Tank)
Pcli_CBASS_GLMM_Tank <- glmmTMB::glmmTMB(fvfm ~ Treatment*CBASS_temp + (1|Tank), family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)
summary(Pcli_CBASS_GLMM_Tank)

qqnorm(resid(Pcli_CBASS_GLMM_Tank))
qqline(resid(Pcli_CBASS_GLMM_Tank))

boxplot(resid(Pcli_CBASS_GLMM_Tank)~Pcli_ipam_CBASS$Treatment) 
boxplot(resid(Pcli_CBASS_GLMM_Tank)~Pcli_ipam_CBASS$CBASS_temp)
boxplot(resid(Pcli_CBASS_GLMM_Tank)~Pcli_ipam_CBASS$Tank)

lrt(PcliCBASS_GLMM_Colony, glmm_PcliCBASS_full)

lrt(Pcli_CBASS_GLMM_Tank, glmm_PcliCBASS_full)

# GLMM Pcli Full: AIC =  -504.0
# GLMM Pcli Colony: AIC =  -506, LRT to full model p-value = 1
# GLMM Pcli Tank: AIC =  -477.2, LRT to full model p-value = 8.420974e-08

#based on AIC and LRT p-values (lowest AIC is best, if p-value is significant that means there is a significant difference in the full model versus the reduced model) ...

#the model with just colony as a fixed effect is not significantly different than the full model. However, tank model is significantly different from full. If I'm interpreting this correctly, it means that the fixed effect of colony is not important to the model, but tank is important to the model. You don't want to have too many variables to try to explain variance.  But the AIC of the full model is better than the AIC of the tank model. I will move forward with the full model.

```

```{r}
Anova(glmm_PcliCBASS_full, type = "II")

Pcli_CBASS_emm <- emmeans(glmm_PcliCBASS_full, specs = ~ Treatment*CBASS_temp)

as.data.frame(pairs(Pcli_CBASS_emm)) -> Pcli_CBASS_treatment_posthoc

list <- c("Control CBASS_temp28 - Variable CBASS_temp28", "Control CBASS_temp30 - Variable CBASS_temp30", "Control CBASS_temp32 - Variable CBASS_temp32","Control CBASS_temp33 - Variable CBASS_temp33","Control CBASS_temp34 - Variable CBASS_temp34" ,"Control CBASS_temp35 - Variable CBASS_temp35","Control CBASS_temp36 - Variable CBASS_temp36","Control CBASS_temp37 - Variable CBASS_temp37")
list <- as.vector(list)

Pcli_CBASS_treatment_posthoc %>% 
  filter(contrast %in% list)
```



## 3c. Raw CBASS Fv/Fm by colony (boxplot)

```{r}
ipam_tidy_data %>% 
  dplyr::filter(Treatment_period == "CBASS") %>% 
  dplyr::select(!Treatment) -> ipam_tidy_CBASS

ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>% 
  dplyr::select(Puck, Colony, Treatment, Species) %>% 
  distinct() %>% #283 rows = one coral ID per treatment, no repeats
  full_join(ipam_tidy_CBASS, by = c("Puck", "Colony", "Species")) %>% 
  drop_na() %>% 
  dplyr::select(Puck:Species, fvfm, Tank:CBASS_temp) -> ipam_tidy_CBASS_treatments

ipam_tidy_CBASS_treatments$Colony <- factor(ipam_tidy_CBASS_treatments$Colony, levels = c('MB-B', 'BC-8b', 'SI-C', 'A', 'B', 'C'))

ipam_tidy_CBASS_treatments %>% 
  mutate(CBASS_temp = as.factor(CBASS_temp)) %>% 
  ggplot(., aes(x=CBASS_temp, y= fvfm, fill = Treatment)) + 
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  geom_boxplot() +
  facet_wrap(~Colony+Species) +
  theme_classic()
```

## 3d. Raw CBASS Fv/Fm by colony (curve)
```{r}
ipam_tidy_CBASS_treatments %>%  
  mutate(CBASS_temp=as.numeric(CBASS_temp)) %>% 
ggplot(., aes(x=CBASS_temp, y = fvfm, color = Treatment)) +
  geom_point() +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  geom_smooth(formula = y ~ x, method = "loess", span = 0.9)  + 
  facet_wrap(~Colony+Species) +
  theme_classic() +
  labs(x="Temperature (ºC)", y = "Fv/Fm")
```


### Statistics 

#### GLM Acer colony co-variate

First need to make column for combination of colony_treatment
```{r}
ipam_tidy_CBASS_treatments %>% 
  dplyr::filter(fvfm < 0.75 & fvfm > 0) %>% #has to be in between 0 and 1
  dplyr::filter(Species  == "Acervicornis") %>% 
  mutate(CBASS_temp=as.factor(CBASS_temp)) -> Acer_ipam_CBASS

Acer_ipam_CBASS$colony_treatment<-interaction(Acer_ipam_CBASS$Colony, Acer_ipam_CBASS$Treatment)
```

```{r}
#model 1 = full model colony_treatment + (1|Tank)
glmm_Acer_full <- glmmTMB(fvfm ~ colony_treatment + (1|Tank), family=beta_family(link = "logit"), data=Acer_ipam_CBASS)

summary(glmm_Acer_full) 

qqnorm(resid(glmm_Acer_full))
qqline(resid(glmm_Acer_full))

boxplot(resid(glmm_Acer_full)~Acer_ipam_CBASS$Treatment)
boxplot(resid(glmm_Acer_full)~Acer_ipam_CBASS$Colony)
boxplot(resid(glmm_Acer_full)~Acer_ipam_CBASS$Tank)

#model 2 = Treatment*time_point + (1|Colony)
Acer_fvfm_GLMM_notank <- glmmTMB::glmmTMB(fvfm ~ colony_treatment, family=beta_family(link = "logit"), data=Acer_ipam_CBASS)
summary(Acer_fvfm_GLMM_notank)

qqnorm(resid(Acer_fvfm_GLMM_notank)) #big yikes
qqline(resid(Acer_fvfm_GLMM_notank))

boxplot(resid(Acer_fvfm_GLMM_notank)~Acer_ipam_CBASS$Treatment) 
boxplot(resid(Acer_fvfm_GLMM_notank)~Acer_ipam_CBASS$Colony)

lrt(Acer_fvfm_GLMM_notank, glmm_Acer_full) #significant

# GLMM Acer Full: AIC =   -405.7
# GLMM Acer no tank: AIC =  -131.5

#based on AIC and LRT p-values (lowest AIC is best, if p-value is significant that means there is a significant difference in the full model versus the reduced model), the full model is the best model.
```


```{r}
Anova(glmm_Acer_full, type = "II")

emmeans(glmm_Acer_full, pairwise ~ colony_treatment)

summary(glht(glmm_Acer_full, linfct=mcp(colony_treatment="Tukey")), test = adjusted(type = "bonferroni"))
```


#### GLM Pcli colony co-variate

First need to make column for combination of colony_treatment
```{r}
ipam_tidy_CBASS_treatments %>% 
  dplyr::filter(fvfm < 0.75) %>% 
  dplyr::filter(Species  == "Pclivosa") %>% 
  mutate(CBASS_temp=as.factor(CBASS_temp)) -> Pcli_ipam_CBASS

Pcli_ipam_CBASS$colony_treatment<-interaction(Pcli_ipam_CBASS$Colony, Pcli_ipam_CBASS$Treatment)
```

```{r}
#model 1 = full model colony_treatment + (1|Tank)
glmm_Pcli_full <- glmmTMB(fvfm ~ colony_treatment + (1|Tank), family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)

summary(glmm_Pcli_full) 

qqnorm(resid(glmm_Pcli_full))
qqline(resid(glmm_Pcli_full))

boxplot(resid(glmm_Pcli_full)~Pcli_ipam_CBASS$Treatment)
boxplot(resid(glmm_Pcli_full)~Pcli_ipam_CBASS$Colony)
boxplot(resid(glmm_Pcli_full)~Pcli_ipam_CBASS$Tank)

#model 2 = Treatment*time_point + (1|Colony)
Pcli_fvfm_GLMM_notank <- glmmTMB::glmmTMB(fvfm ~ colony_treatment, family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)
summary(Pcli_fvfm_GLMM_notank)

qqnorm(resid(Pcli_fvfm_GLMM_notank)) #big yikes
qqline(resid(Pcli_fvfm_GLMM_notank))

boxplot(resid(Pcli_fvfm_GLMM_notank)~Pcli_ipam_CBASS$Treatment) 
boxplot(resid(Pcli_fvfm_GLMM_notank)~Pcli_ipam_CBASS$Colony)

lrt(Pcli_fvfm_GLMM_notank, glmm_Pcli_full) #significant

# GLMM Pcli Full: AIC =   -491.5
# GLMM Pcli no tank: AIC =  -202.9

#based on AIC and LRT p-values (lowest AIC is best, if p-value is significant that means there is a significant difference in the full model versus the reduced model), the full model is the best model.
```


```{r}
Anova(glmm_Pcli_full, type = "II")

emmeans(glmm_Pcli_full, pairwise ~ colony_treatment)

summary(glht(glmm_Acer_full, linfct=mcp(colony_treatment="Tukey")), test = adjusted(type = "bonferroni"))
```





# Ross Cunning's DRC analysis

```{r}
# Define dose-response curves using same constraints
ll3 <- function(data) {
  drm(fvfm ~ CBASS_temp, data = data,
      fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.01, 30))} 

# I changed the max value from 0.55 to 0.01 so I don't get NaN for std.error, statistic, or p-value.
#these are the limits Ross had set in his code:
#upperl = c(120, 0.72, 40),
#lowerl = c(10, 0.55, 30))

tryll3 <- possibly(ll3, otherwise = NULL)

# Fit model to each coral, get parameters, fitted values, and residuals
initmods_species <- ipam_tidy_CBASS_treatments %>%
  dplyr::select(!c(Colony, Tank, Date, Treatment_period)) %>% 
  nest(data = c(Puck, CBASS_temp, fvfm)) %>% #this line wasn't working for the longest time because I hadn't added "Puck" to the list for the nested data. 
  mutate(ll3 = map(data, tryll3)) %>% # Fit the model to each coral
  # Get model parameters and fitted values/residuals
  mutate(pars = map(ll3, tidy), #tidy comes from the broom package I think?
         pred = map2(ll3, data, ~augment(.x, drop_na(.y, fvfm))))

# Extract ed50 parameter values for each Species and Treatment combination from model fits
ed50_species <- initmods_species %>% 
  unnest(data) %>% 
  dplyr::select(Species, Treatment, pars) %>%
  unnest(pars) %>%
  filter(term == "ED50")

ed50_species %>% distinct() -> ed50_species

# Collect raw data, fitted values, and ed50 estimates
fvals_species <- initmods_species %>%
  dplyr::select(Species, Treatment, pred) %>%
  unnest(pred) %>%
  full_join(ed50_species) %>%
  rename(ed50 = estimate)
```


```{r}
ggplot(data = fvals_species, aes(x=CBASS_temp, y=fvfm)) +
  geom_point(data = fvals_species, aes(color=Treatment)) +
    geom_line(data = drop_na(fvals_species, .fitted),
              aes(y = .fitted)) +
    geom_vline(data = distinct(fvals_species, Species, Treatment, ed50), 
               aes(xintercept = ed50), 
               lwd = 0.2, lty = 2) +
    geom_text(data = distinct(fvals_species, Species, Treatment, ed50),
              aes(x = ed50, y = 0.05, label = round(ed50, 2)), 
              hjust = 1, nudge_x = -0.2, size = 3) +
    facet_grid(vars(Treatment), vars(Species)) +
  labs(x = "Temperature (C)", y = "Fv/Fm") +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_bw() +
  theme(text = element_text(size = 14)) 
```

## 3f. Hill slope (i wrote this)

### DRC for Species
```{r}
# Define dose-response curves using same constraints
ll3 <- function(data) {
  drm(fvfm ~ CBASS_temp, data = data,
      fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.01, 30))} 

# I changed the max value from 0.55 to 0.01 so I don't get NaN for std.error, statistic, or p-value.
#these are the limits Ross had set in his code:
#upperl = c(120, 0.72, 40),
#lowerl = c(10, 0.55, 30))

tryll3 <- possibly(ll3, otherwise = NULL)

# Fit model to each coral, get parameters, fitted values, and residuals
initmods_species <- ipam_tidy_CBASS_treatments %>%
  dplyr::select(!c(Colony, Tank, Date, Treatment_period)) %>% 
  nest(data = c(Puck, CBASS_temp, fvfm)) %>% #this line wasn't working for the longest time because I hadn't added "Puck" to the list for the nested data. 
  mutate(ll3 = map(data, tryll3)) %>% # Fit the model to each coral
  # Get model parameters and fitted values/residuals
  mutate(pars = map(ll3, tidy), #tidy comes from the broom package I think?
         pred = map2(ll3, data, ~augment(.x, drop_na(.y, fvfm))))

# Extract ed50 parameter values for each Species and Treatment combination from model fits
hill_species <- initmods_species %>% 
  unnest(data) %>% 
  dplyr::select(Species, Treatment, pars) %>%
  unnest(pars) %>%
  filter(term == "hill")

hill_species %>% distinct() -> hill_species

# Collect raw data, fitted values, and ed50 estimates
fvals_species_hill <- initmods_species %>%
  dplyr::select(Species, Treatment, pred) %>%
  unnest(pred) %>%
  full_join(hill_species) %>%
  rename(hill = estimate)
```

## 3f. ED50 DRM curves for raw fv/fm with colony 

### DRC for Colony
```{r, include = F}
# Define dose-response curves using same constraints
ll3 <- function(data) {
  drm(fvfm ~ CBASS_temp, data = data,
      fct = LL.3(names = c("hill", "max", "ED50")),
      upperl = c(120, 0.72, 40),
      lowerl = c(10, 0.01, 30))} 

# when I remove the lower limit from the ll3 function, SI-C variable ED50 is 36.25712. But when I keep both limits, it is 35.96952 but the std.error, statistic, and p-value are NaN. I narrowed it down and it is an issue with the "max" value. I changed it from 0.55 to 0.01, and I get 36.25712 for the ED50. Below are the upper and lower limits set by Ross in his code.
#upperl = c(120, 0.72, 40),
#lowerl = c(10, 0.55, 30))

tryll3 <- possibly(ll3, otherwise = NULL)

# Fit model to each coral, get parameters, fitted values, and residuals
initmods_colony <- ipam_tidy_CBASS_treatments %>%
  dplyr::select(!c(Tank, Date, Treatment_period)) %>% 
  nest(data = c(Puck, CBASS_temp, fvfm)) %>% #this line wasn't working for the longest time because I hadn't added "Puck" to the list for the nested data. 
  mutate(ll3 = map(data, tryll3)) %>% # Fit the model to each coral
  # Get model parameters and fitted values/residuals
  mutate(pars = map(ll3, tidy), #tidy comes from the broom package I think?
         pred = map2(ll3, data, ~augment(.x, drop_na(.y, fvfm))))

# Extract ed50 parameter values for each Colony and Treatment combination from model fits
ed50_colony <- initmods_colony%>% 
  unnest(data) %>% 
  dplyr::select(Colony, Treatment, pars) %>%
  unnest(pars) %>%
  filter(term == "ED50")

ed50_colony %>% distinct() -> ed50_colony

# Collect raw data, fitted values, and ed50 estimates
fvals_colony <- initmods_colony %>%
  dplyr::select(Species, Treatment, Colony, pred) %>%
  unnest(pred) %>%
  full_join(ed50_colony) %>%
  rename(ed50 = estimate)
```


```{r}
fvals_colony$Colony <- factor(fvals_colony$Colony, levels = c('MB-B', 'BC-8b', 'SI-C', 'A', 'B', 'C'))

fvals_colony

ggplot(data = fvals_colony, aes(x=CBASS_temp, y=fvfm)) +
  geom_point(data = fvals_colony, aes(color=Treatment)) +
    geom_line(data = drop_na(fvals_colony, .fitted),
              aes(y = .fitted)) +
    geom_vline(data = distinct(fvals_colony, Colony, Treatment, ed50), 
               aes(xintercept = ed50), 
               lwd = 0.2, lty = 2) +
    geom_text(data = distinct(fvals_colony, Colony, Treatment, ed50),
              aes(x = ed50, y = 0.05, label = round(ed50, 2)), 
              hjust = 1, nudge_x = -0.2, size = 3) +
    facet_grid(vars(Treatment), vars(Colony)) +
  labs(x = "Temperature (C)", y = "Fv/Fm") +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_bw() +
  theme(text = element_text(size = 14)) 
```

### ED50 Statistical test 

Take all genotypes together and compare the mean ED50 of treatment (that way you have multiple samples to add to the total group mean)

Kruskall-wallis test
```{r}
#Acer
ed50_colony %>% 
  filter(Colony == "BC-8b" | Colony == "MB-B" | Colony == "SI-C") %>% 
kruskal_test(data = ., estimate ~ Treatment) #not significant


#Pcli
ed50_colony %>% 
  filter(Colony == "A" | Colony == "B" | Colony == "C") %>% 
kruskal_test(data = ., estimate ~ Treatment) #not significant
```

### Independent t-test for Acer
```{r}
ed50_colony %>% 
  filter(Colony == "BC-8b" | Colony == "MB-B" | Colony == "SI-C") %>% 
  group_by(Treatment) %>% 
  shapiro_test(estimate) #normality test not significant

ed50_colony %>% 
  filter(Colony == "BC-8b" | Colony == "MB-B" | Colony == "SI-C") %>% 
  levene_test(estimate ~ Treatment) # variances are equal (p-value not significant)

ed50_colony %>% 
  filter(Colony == "BC-8b" | Colony == "MB-B" | Colony == "SI-C") %>% 
  t_test(estimate ~ Treatment) #ed50s are not significantly different for Acer
```

### Independent t-test for Pcli
```{r}
ed50_colony %>% 
  filter(Colony == "A" | Colony == "B" | Colony == "C") %>% 
  group_by(Treatment) %>% 
  shapiro_test(estimate) #normality test not significant

ed50_colony %>% 
   filter(Colony == "A" | Colony == "B" | Colony == "C") %>% 
  levene_test(estimate ~ Treatment) # variances are equal (p-value not significant)

ed50_colony %>% 
   filter(Colony == "A" | Colony == "B" | Colony == "C") %>% 
  t_test(estimate ~ Treatment) #ed50s are not significantly different for Pcli
```





# Rachel Alderice DRC analysis

```{r}
#curve ID treatment - filter for control
ipam_tidy_CBASS_treatments %>% filter(Treatment == "Control") -> control

DRCpam_control = drm(fvfm ~ CBASS_temp, data = control, curveid = Species,
             fct = LL.3(names = c('hill', 'max', 'ed50')))
summary(DRCpam_control)
compParm(DRCpam_control, 'ed50')
compParm(DRCpam_control, 'ed50', "-")
plot(DRCpam_control)

#significant!!

compParm(DRCpam_control, 'hill')
compParm(DRCpam_control, 'hill', "-")

#curve ID treatment - filter for variable
ipam_tidy_CBASS_treatments %>% filter(Treatment == "Variable") -> variable

DRCpam_variable = drm(fvfm ~ CBASS_temp, data = variable, curveid = Species,
             fct = LL.3(names = c('hill', 'max', 'ed50')))
summary(DRCpam_variable)
compParm(DRCpam_variable, 'ed50')
#significant!
compParm(DRCpam_variable, 'ed50', "-")
plot(DRCpam_variable)

compParm(DRCpam_variable, 'hill')
compParm(DRCpam_variable, 'hill', "-")
```


# Evensen DRC analysis

```{r}
Acer_ipam_CBASS -> Acer_CBASS
Acer_CBASS$CBASS_temp <- as.numeric(as.character(Acer_CBASS$CBASS_temp))
CBASS_Acer_DRC <- drm(fvfm ~ CBASS_temp, data = Acer_CBASS, curveid = Treatment, fct = LL.3(names = c('hill', 'max', 'ed50')))
mselect(CBASS_Acer_DRC, list(LL.2(), LL.4(), LL.5(), LL2.2(), LL2.3(), LL2.4(), LL2.5(), AR.2(), AR.3(), EXD.2(), EXD.3()), icfct = AIC)
modelFit(CBASS_Acer_DRC)
summary(CBASS_Acer_DRC)
plot(CBASS_Acer_DRC)

#extract ED50
Acer_CBASS_variable_coeff<-CBASS_Acer_DRC$coefficients[5]
Acer_CBASS_control_coeff<-CBASS_Acer_DRC$coefficients[6]

Acer_coeff <- data.frame(Acer_CBASS_variable_coeff,Acer_CBASS_control_coeff)

control_CBASS <- drm(fvfm ~ CBASS_temp, data = Acer_CBASS[Acer_CBASS$Treatment=="Control",], fct = LL.3())
summary(control_CBASS)
plot(control_CBASS)

variable_CBASS <- drm(fvfm ~ CBASS_temp, data = Acer_CBASS[Acer_CBASS$Treatment=="Variable",], fct = LL.3())
summary(variable_CBASS)
plot(variable_CBASS)

control_CBASS_preddata = data.frame(temp = seq(27,39, length.out = 100))
control_CBASS_pred = as.data.frame(predict(control_CBASS, newdata = control_CBASS_preddata, interval = 'confidence'))
control_CBASS_preddata = data.frame(control_CBASS_preddata, fvfm = control_CBASS_pred$Prediction, Lower = control_CBASS_pred$Lower, Upper = control_CBASS_pred$Upper)

variable_CBASS_preddata = data.frame(temp = seq(27,39, length.out = 100))
variable_CBASS_pred = as.data.frame(predict(variable_CBASS, newdata = variable_CBASS_preddata, interval = 'confidence'))
variable_CBASS_preddata = data.frame(variable_CBASS_preddata, fvfm = variable_CBASS_pred$Prediction, Lower = variable_CBASS_pred$Lower, Upper = variable_CBASS_pred$Upper)


ggplot() +
  geom_jitter(data = Acer_CBASS, aes(x = CBASS_temp, y = fvfm, color = Treatment), size = 1, width = 0.25) +
  scale_x_continuous(limits=c(26,40), breaks=c(26,28,30,32,34,36,38)) +
  scale_y_continuous(limits=c(-0.1, 0.7), breaks=c(0, 0.2, 0.4, 0.6)) +
  geom_line(data = control_CBASS_preddata, aes(x = temp, y = fvfm), color = '#66ccfe', show.legend = FALSE) +
  geom_ribbon(data = control_CBASS_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = '#66ccfe', linetype=2, alpha = 0.2) +
  geom_vline(data = Acer_coeff, aes(xintercept = Acer_CBASS_control_coeff), color = '#66ccfe', show.legend = FALSE) +
  geom_text(data = Acer_coeff, aes(label = Acer_CBASS_control_coeff), x = 30, y = 0.2, show.legend = FALSE, color = '#66ccfe') +
 
  geom_line(data = variable_CBASS_preddata, aes(x = temp, y = fvfm), color = '#e92000', show.legend = FALSE) +
  geom_ribbon(data = variable_CBASS_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = '#e92000', linetype=2, alpha = 0.2) +
  geom_vline(data = Acer_coeff, aes(xintercept = Acer_CBASS_variable_coeff), color = '#e92000', show.legend = FALSE) +
  geom_text(data = Acer_coeff, aes(label = Acer_CBASS_variable_coeff), x = 30, y = 0.1, show.legend = FALSE, color = '#e92000') +

  scale_color_manual(values=c('#66ccfe','#e92000')) +
  ylab("Fv/Fm") +
  xlab("Temperature (°C)") +
  theme_bw() +
  labs(title = "Acervicornis")
```

Pcli
```{r}
Pcli_ipam_CBASS -> Pcli_CBASS
Pcli_CBASS$CBASS_temp <- as.numeric(as.character(Pcli_CBASS$CBASS_temp))
CBASS_Pcli_DRC <- drm(fvfm ~ CBASS_temp, data = Pcli_CBASS, curveid = Treatment, fct = LL.3(names = c('hill', 'max', 'ed50')))
mselect(CBASS_Pcli_DRC, list(LL.2(), LL.4(), LL.5(), LL2.2(), LL2.3(), LL2.4(), LL2.5(), AR.2(), AR.3(), EXD.2(), EXD.3()), icfct = AIC)
modelFit(CBASS_Pcli_DRC)
summary(CBASS_Pcli_DRC)
plot(CBASS_Pcli_DRC)

#extract ED50
Pcli_CBASS_variable_coeff<-CBASS_Pcli_DRC$coefficients[5]
Pcli_CBASS_control_coeff<-CBASS_Pcli_DRC$coefficients[6]

Pcli_coeff <- data.frame(Pcli_CBASS_variable_coeff,Pcli_CBASS_control_coeff)

control_CBASS <- drm(fvfm ~ CBASS_temp, data = Pcli_CBASS[Pcli_CBASS$Treatment=="Control",], fct = LL.3())
summary(control_CBASS)
plot(control_CBASS)

variable_CBASS <- drm(fvfm ~ CBASS_temp, data = Pcli_CBASS[Pcli_CBASS$Treatment=="Variable",], fct = LL.3())
summary(variable_CBASS)
plot(variable_CBASS)

control_CBASS_preddata = data.frame(temp = seq(27,39, length.out = 100))
control_CBASS_pred = as.data.frame(predict(control_CBASS, newdata = control_CBASS_preddata, interval = 'confidence'))
control_CBASS_preddata = data.frame(control_CBASS_preddata, fvfm = control_CBASS_pred$Prediction, Lower = control_CBASS_pred$Lower, Upper = control_CBASS_pred$Upper)

variable_CBASS_preddata = data.frame(temp = seq(27,39, length.out = 100))
variable_CBASS_pred = as.data.frame(predict(variable_CBASS, newdata = variable_CBASS_preddata, interval = 'confidence'))
variable_CBASS_preddata = data.frame(variable_CBASS_preddata, fvfm = variable_CBASS_pred$Prediction, Lower = variable_CBASS_pred$Lower, Upper = variable_CBASS_pred$Upper)


ggplot() +
  geom_jitter(data = Pcli_CBASS, aes(x = CBASS_temp, y = fvfm, color = Treatment), size = 1, width = 0.25) +
  scale_x_continuous(limits=c(26,40), breaks=c(26,28,30,32,34,36,38)) +
  scale_y_continuous(limits=c(-0.1, 0.7), breaks=c(0, 0.2, 0.4, 0.6)) +
  geom_line(data = control_CBASS_preddata, aes(x = temp, y = fvfm), color = '#66ccfe', show.legend = FALSE) +
  geom_ribbon(data = control_CBASS_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = '#66ccfe', linetype=2, alpha = 0.2) +
  geom_vline(data = Pcli_coeff, aes(xintercept = Pcli_CBASS_control_coeff), color = '#66ccfe', show.legend = FALSE) +
  geom_text(data = Pcli_coeff, aes(label = Pcli_CBASS_control_coeff), x = 30, y = 0.2, show.legend = FALSE, color = '#66ccfe') +
 
  geom_line(data = variable_CBASS_preddata, aes(x = temp, y = fvfm), color = '#e92000', show.legend = FALSE) +
  geom_ribbon(data = variable_CBASS_preddata, aes(x = temp, ymin=Lower, ymax=Upper), color = '#e92000', linetype=2, alpha = 0.2) +
  geom_vline(data = Pcli_coeff, aes(xintercept = Pcli_CBASS_variable_coeff), color = '#e92000', show.legend = FALSE) +
  geom_text(data = Pcli_coeff, aes(label = Pcli_CBASS_variable_coeff), x = 30, y = 0.1, show.legend = FALSE, color = '#e92000') +

  scale_color_manual(values=c('#66ccfe','#e92000')) +
  ylab("Fv/Fm") +
  xlab("Temperature (°C)") +
  theme_bw() +
  labs(title = "Pclivosa")
```

























# 4. Normalized to initial measurement (treatment day 0 - T1/T0) - CBASS

```{r}
ipam_tidy_data %>% 
  filter(Date < "2022-03-22") %>% 
  dplyr::select(fvfm, Puck, Colony, Date, Treatment,Species) %>% 
  mutate(fvfm_T0 = fvfm) %>% 
  dplyr::select(!fvfm) -> preCBASSfvfm

ipam_tidy_data %>% 
  dplyr::filter(Treatment_period == "CBASS") %>% 
  dplyr::select(fvfm, Puck:Date, CBASS_temp, Species) %>% 
  mutate(fvfm_T1 = fvfm) %>% 
  dplyr::select(!fvfm) %>% 
  left_join(preCBASSfvfm, by = c("Puck", "Colony", "Species")) %>% 
  dplyr::select(!Date.x) %>% 
  dplyr::select(!Date.y) %>% 
  filter(fvfm_T0 < 0.75) %>% #there were some weird outliers where the fvfm = 2.00
  filter(fvfm_T1 < 0.75) %>% 
  mutate(CBASS_normalize_fvfm = fvfm_T1/fvfm_T0) %>% 
  filter(CBASS_normalize_fvfm > 0) %>% 
  filter(CBASS_normalize_fvfm < 1) %>% 
  mutate(CBASS_temp = as.factor(CBASS_temp))-> ipam_CBASS_normalized

  ggplot(ipam_CBASS_normalized, aes(x=CBASS_temp, y = CBASS_normalize_fvfm, fill=Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species) + 
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic()
```

```{r}
ipam_CBASS_normalized %>% 
  mutate(CBASS_temp=as.numeric(as.character(CBASS_temp))) %>% 
ggplot(., aes(x=CBASS_temp, y = CBASS_normalize_fvfm, color = Treatment)) +
  geom_point() +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  geom_smooth(formula = y ~ x, method = "loess", span = 0.9)  + 
  facet_wrap(~Species) +
  theme_classic() +
  labs(x="Temperature (ºC)", y = "Normalized Fv/Fm")
```

### GLM Acer

```{r}
ipam_CBASS_normalized %>% 
  filter(Species == "Acervicornis") -> Acer_ipam_CBASS

#model 1 = full model
glmm_AcerCBASS_full <- glmmTMB(CBASS_normalize_fvfm ~ Treatment*CBASS_temp + (1|Colony), family=beta_family(link = "logit"), data=Acer_ipam_CBASS)

summary(glmm_AcerCBASS_full) 

qqnorm(resid(glmm_AcerCBASS_full))
qqline(resid(glmm_AcerCBASS_full))

boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$Treatment)
boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$CBASS_temp)
boxplot(resid(glmm_AcerCBASS_full)~Acer_ipam_CBASS$Colony)

#model 2 
AcerCBASS_GLMM <- glmmTMB::glmmTMB(CBASS_normalize_fvfm ~ Treatment*CBASS_temp, family=beta_family(link = "logit"), data=Acer_ipam_CBASS)
summary(AcerCBASS_GLMM)

#need to check normality and variance visually (boxplots of fixed effects only)
qqnorm(resid(AcerCBASS_GLMM))
qqline(resid(AcerCBASS_GLMM))

boxplot(resid(AcerCBASS_GLMM)~Acer_ipam_CBASS$Treatment) 
boxplot(resid(AcerCBASS_GLMM)~Acer_ipam_CBASS$CBASS_temp)

lrt(AcerCBASS_GLMM, glmm_AcerCBASS_full)

lrt(Acer_CBASS_GLMM_Tank, glmm_AcerCBASS_full)

# GLMM Acer Full: AIC =  -323.6 
# GLMM Acer no Colony: AIC =  -324.1

# no colony is better based on AIC, and lrt p-value is not significant
```


```{r}
Anova(AcerCBASS_GLMM, type = "II")

Acer_CBASS_emm <- emmeans(AcerCBASS_GLMM, specs = ~ Treatment*CBASS_temp)

as.data.frame(pairs(Acer_CBASS_emm))
```


### GLM Pcli

```{r}
ipam_CBASS_normalized %>% 
  filter(Species == "Pclivosa") -> Pcli_ipam_CBASS

#model 1 = full model
glmm_PcliCBASS_full <- glmmTMB(CBASS_normalize_fvfm ~ Treatment*CBASS_temp + (1|Colony), family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)

summary(glmm_PcliCBASS_full) 

qqnorm(resid(glmm_PcliCBASS_full))
qqline(resid(glmm_PcliCBASS_full))

boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$Treatment)
boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$CBASS_temp)
boxplot(resid(glmm_PcliCBASS_full)~Pcli_ipam_CBASS$Colony)

#model 2 
PcliCBASS_GLMM <- glmmTMB::glmmTMB(CBASS_normalize_fvfm ~ Treatment*CBASS_temp, family=beta_family(link = "logit"), data=Pcli_ipam_CBASS)
summary(PcliCBASS_GLMM)

#need to check normality and variance visually (boxplots of fixed effects only)
qqnorm(resid(PcliCBASS_GLMM))
qqline(resid(PcliCBASS_GLMM))

boxplot(resid(PcliCBASS_GLMM)~Pcli_ipam_CBASS$Treatment) 
boxplot(resid(PcliCBASS_GLMM)~Pcli_ipam_CBASS$CBASS_temp)

lrt(PcliCBASS_GLMM, glmm_PcliCBASS_full)

# GLMM Pcli Full: AIC =  -355.8 
# GLMM Pcli no Colony: AIC =  -344.3  , LRT p-value = 0.0002403821

# full model is better based on AIC and p-value
```


```{r}
Anova(glmm_PcliCBASS_full, type = "II")

Pcli_CBASS_emm <- emmeans(glmm_PcliCBASS_full, specs = ~ Treatment*CBASS_temp)

as.data.frame(pairs(Pcli_CBASS_emm))
```






# 5. Normalized to final treatment measurement (treatment day 28) - CBASS

```{r}
ipam_tidy_data %>% 
  filter(Date == "2022-04-20") %>% 
  dplyr::select(fvfm, Puck, Colony, Date, Treatment,Species) %>% 
  mutate(fvfm_T0 = fvfm) %>% 
  dplyr::select(!fvfm) -> preCBASSfvfm_endoftreatment

ipam_tidy_data %>% 
  dplyr::filter(Treatment_period == "CBASS") %>% 
  dplyr::select(fvfm, Puck:Date, CBASS_temp, Species) %>% 
  mutate(fvfm_T1 = fvfm) %>% 
  dplyr::select(!fvfm) %>% 
  left_join(preCBASSfvfm_endoftreatment, by = c("Puck", "Colony", "Species")) %>% 
  dplyr::select(!Date.x) %>% 
  dplyr::select(!Date.y) %>% 
  filter(fvfm_T0 < 0.75) %>% #there were some weird outliers where the fvfm = 2.00
  filter(fvfm_T1 < 0.75) %>% 
  mutate(CBASS_normalize_fvfm = fvfm_T1/fvfm_T0) %>% 
  #filter(CBASS_normalize_fvfm > 0) %>% 
  #filter(CBASS_normalize_fvfm < 1) %>% 
  mutate(CBASS_temp = as.factor(CBASS_temp))-> ipam_CBASS_normalized_endoftreatment

  ggplot(ipam_CBASS_normalized_endoftreatment, aes(x=CBASS_temp, y = CBASS_normalize_fvfm, fill=Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species) + 
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic()
  
  #there are a lot of values > 1 here so I can't just filter it out so the data is 0<x<1 
```

