---
title: "GO_terms_tidying"
author: "allyson_demerlis"
date: "2024-02-22"
output: html_document
---

```{r}
library(tidyverse)
```

CC Treated vs. Initial
```{r}
cc_TreatedvInitial <- read.csv("CC_Treated_vs_Initial_fc.csv", sep = "")

cc_TreatedvInitial %>% 
  rename(lfc = value) %>% 
  rename (level = lev) -> cc_TreatedvInitial

MWU_result_CC_TreatedvInitial <- read.csv("MWU_CC_Treated_vs_Initial_fc.csv", sep = "")

full_join(cc_TreatedvInitial, MWU_result_CC_TreatedvInitial, by = c("level", "term", "name")) %>% view()
  filter(p.adj < 0.05) %>% 
  distinct(seq)
  
# these tables are going to be way to big -- CC for Treated v Initial alone is like 30,000 rows...
```

Instead lets just make a table for each dds contrast with significant (p.adj < 0.05) BP, MF, and CC terms in one table

Treated vs Initial
```{r}
MWU_result_CC_TreatedvInitial <- read.csv("MWU_CC_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_CC_TreatedvInitial %>% 
  filter(p.adj <0.05) %>%  #3 GO terms
  mutate(GO_category = "CC") -> MWU_result_CC_TreatedvInitial_sig

MWU_result_CC_TreatedvInitial_sig %>% 
  write_csv("TreatedvsInitial_GOMWU_sigterms.csv")

MWU_result_MF_TreatedvInitial <- read.csv("MWU_MF_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_MF_TreatedvInitial %>% 
  filter(p.adj <0.05)  #0 GO terms

MWU_result_BP_TreatedvInitial <- read.csv("MWU_BP_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_BP_TreatedvInitial %>%
  filter(p.adj <0.05)   #0 GO terms

```


Treated vs Untreated
```{r}
MWU_result_CC_TreatedvUntreated <- read.csv("MWU_CC_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_CC_TreatedvUntreated %>% 
  filter(p.adj <0.05)  #0 GO terms

MWU_result_MF_TreatedvUntreated <- read.csv("MWU_MF_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_MF_TreatedvUntreated %>% 
  filter(p.adj <0.05)  #0 GO terms

MWU_result_BP_TreatedvUntreated <- read.csv("MWU_BP_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_BP_TreatedvUntreated %>%
  filter(p.adj <0.05) %>%  #0 GO terms
```

Untreated vs Initial

```{r}
MWU_result_CC_UntreatedvInitial <- read.csv("MWU_CC_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_CC_UntreatedvInitial %>% 
  filter(p.adj <0.05)  #0 GO terms
  mutate(GO_category = "CC") -> MWU_result_CC_UntreatedvInitial_sig

MWU_result_MF_UntreatedvInitial <- read.csv("MWU_MF_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_MF_UntreatedvInitial %>% 
  filter(p.adj <0.05) #0 GO terms

MWU_result_BP_UntreatedvInitial <- read.csv("MWU_BP_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_BP_UntreatedvInitial %>%
  filter(p.adj <0.05) #0 GO terms
```

