---
title: "tank_temp_figures"
author: "allyson_demerlis"
date: "2024-11-11"
output: html_document
---

# Load libraries and data
```{r}
library(tidyverse)
library(plotrix) #for standard error function
library(cowplot)

feb4<- readxl::read_xlsx("Data from LabVIEW/20220204_TemperatureData edited.xlsx", col_types = c("date", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric"))
feb23 <- readxl::read_xlsx("Data from LabVIEW/20220223_TemperatureData edited.xlsx")
march7 <- readxl::read_xlsx("Data from LabVIEW/20220307_TemperatureData - Copy.xlsx")
march29 <- readxl::read_xlsx("Data from LabVIEW/20220329_TemperatureData.xlsx")
april15 <- readxl::read_xlsx("Data from LabVIEW/20220415_TemperatureData.xlsx")

april25 <- readxl::read_xlsx("Data from LabVIEW/20220425_TemperatureData (1).xlsx") #had to convert to csv because dates weren't reading in correctly

CBASS <- readxl::read_xlsx("Data from LabVIEW/CBASS_labview.xlsx")
```


Creating subsets for different stages of experiment
```{r}
acclimation <- bind_rows(feb4, feb23) 
temptreatment<- bind_rows(march7, march29, april15) #treatment started march 24 and ended april 20
temptreatment %>% filter(Time >= as.Date("2022-03-24") & Time <= as.Date("2022-04-22")) -> temptreatment
#filter out one anomalous data point
  temptreatment %>% 
    filter(T1<23) 
# Set the specific data point to NA
temptreatment[temptreatment$T1 < 22.5, "T1"]= NA
  temptreatment %>% 
    filter(T1<23) #its gone now

CBASS #april 22
```

# Acclimation period
```{r}
#tanks 1-4 had Pclivosa in them since Feb 4, so they were maintained at 24 C. Tanks 5-8 didn't get Acer until Feb 24, so that's why their temps fluctuate before then.
#need to remove tank temps for tanks 5-8 for dates before feb 24 and before feb 17 for tanks 1-4
#ended up doing this manually in Excel

acclimation %>% 
  filter(Time >= as.Date("2022-02-17")) %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  ggplot(data = .,
       aes(x = Time, y = Temp, color = Tank)) +
  geom_line() + 
  theme_classic() +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Date") +
  scale_y_continuous(limits = c(22, 26)) +
  theme(text = element_text(size = 16)) #all font sizes
```

```{r}
#need to clean up this graph,summarize all the tanks together
acclimation$Time <- lubridate::round_date(acclimation$Time, unit = "15 min")

acclimation %>% 
  filter(Time >= as.Date("2022-02-17")) %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  drop_na(Temp) %>% 
  dplyr::group_by(Time) %>% 
  dplyr::summarise(avg_temp = mean(Temp), se = std.error(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp)) + 
  geom_point() +
  geom_smooth() +
  theme_classic() +
   geom_line() +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Date") +
  scale_y_continuous(limits = c(22, 26)) +
  theme(text = element_text(size = 16)) #all font sizes
```

# Treatment period (Mar 24 - Apr 20)
```{r}
#all tanks separated
temptreatment %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  ggplot(data = .,
       aes(x = Time, y = Temp, color = Tank)) +
  geom_point() + 
  theme_classic() +
  facet_wrap(~Tank) +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Date")
#ggsave("temptreatment_28d.pdf", width = 10, height = 6)
```

```{r}
#summary figure of average tank temp during 24-hr period over the 28 days
temptreatment$Time <- lubridate::round_date(temptreatment$Time, unit = "15 min")
temptreatment %>% drop_na() -> temptreatment

temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable", 
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control",
                               Tank == "T8" ~ "variable")) %>% 
  dplyr::group_by(Time, treatment) %>% 
  dplyr::summarise(avg_temp = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = treatment)) + 
  theme_classic() +
  geom_ribbon(aes(fill=treatment, group = treatment, ymin = avg_temp - sd, ymax = avg_temp + sd), alpha = 0.2) +
   geom_line(aes(group = treatment)) +
  scale_x_discrete(breaks = c("03:00:00", "09:00:00", "15:00:00", "21:00:00")) +
  labs(y="Temperature (ºC)") +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))+
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 15)) 
#ggsave("summarytemptreatment_stdeviation.pdf", width = 10, height = 6)
```

Calculate the mean daily min, max, and variance for each tank/treatment over the 28-d period.
```{r}
temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "Untreated",
                               Tank == "T2" ~ "Treated",
                               Tank == "T3" ~ "Treated",
                               Tank == "T4" ~ "Untreated",
                               Tank == "T5" ~ "Treated", 
                               Tank == "T6" ~ "Untreated",
                               Tank == "T7" ~ "Untreated",
                               Tank == "T8" ~ "Treated")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ungroup() %>% 
  group_by(Tank, Treatment, Species) %>% 
  summarise(mean_daily_max=mean(daily_max),sd_daily_max=sd(daily_max), mean_daily_min=mean(daily_min),sd_daily_min=sd(daily_min),mean_daily_range=mean(daily_max - daily_min), sd_daily_range=sd(daily_max - daily_min)) -> meanmaxminrange_summarytable

meanmaxminrange_summarytable %>% 
  group_by(Treatment) %>% summarise(mean_max=mean(mean_daily_max),sd_max=sd(mean_daily_max), mean_min=mean(mean_daily_min),sd_min=sd(mean_daily_min),mean_range=mean(mean_daily_max - mean_daily_min), sd_daily_range=sd(mean_daily_max - mean_daily_min)) %>% 
  write_csv("meanmaxminrange.csv")

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_max, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_max - sd_daily_max, ymax = mean_daily_max + sd_daily_max), width = 0.2) +
  theme_classic()

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_min, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_min - sd_daily_min, ymax = mean_daily_min + sd_daily_min), width = 0.2) +
  theme_classic()

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_range, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_range - sd_daily_range, ymax = mean_daily_range + sd_daily_range), width = 0.2) +
  theme_classic()
```


Calculate the mean min, max, and variance for each treatment over the 28-d period.
```{r}
temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "Untreated",
                               Tank == "T2" ~ "Treated",
                               Tank == "T3" ~ "Treated",
                               Tank == "T4" ~ "Untreated",
                               Tank == "T5" ~ "Treated", 
                               Tank == "T6" ~ "Untreated",
                               Tank == "T7" ~ "Untreated",
                               Tank == "T8" ~ "Treated")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Date, Treatment, Species) %>% 
  summarise(max = max(Temp), min = min(Temp), range = (max(Temp)-min(Temp))) %>% 
  group_by(Treatment) %>% 
  summarise(count=n(), mean_daily_max=mean(max),sd_daily_max=sd(max), mean_daily_min=mean(min),sd_daily_min=sd(min),mean_daily_range=mean(max - min), sd_daily_range=sd(max - min)) 
```





Combined plot 
```{r}
p1=temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "Untreated",
                               Tank == "T2" ~ "Treated",
                               Tank == "T3" ~ "Treated",
                               Tank == "T4" ~ "Untreated",
                               Tank == "T5" ~ "Treated", 
                               Tank == "T6" ~ "Untreated",
                               Tank == "T7" ~ "Untreated",
                               Tank == "T8" ~ "Tank 8")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ggplot(., aes(x=Treatment, y=daily_max, color=Tank)) +
  geom_boxplot() + 
  theme_classic() +
  labs(y="Daily Maximum Temperature (ºC)") + 
  theme(legend.position = "none") +
  theme(text = element_text(size = 15)) 

p2=temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "Untreated",
                               Tank == "T2" ~ "Treated",
                               Tank == "T3" ~ "Treated",
                               Tank == "T4" ~ "Untreated",
                               Tank == "T5" ~ "Treated", 
                               Tank == "T6" ~ "Untreated",
                               Tank == "T7" ~ "Untreated",
                               Tank == "T8" ~ "Tank 8")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ggplot(., aes(x=Treatment, y=daily_min, color=Tank)) +
  geom_boxplot() + 
  theme_classic() +
  labs(y="Daily Minimum Temperature (ºC)") + 
  ylim(27,28.5) + 
  theme(legend.position = "none") +
  theme(text = element_text(size = 15)) 

p3=temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "Untreated",
                               Tank == "T2" ~ "Treated",
                               Tank == "T3" ~ "Treated",
                               Tank == "T4" ~ "Untreated",
                               Tank == "T5" ~ "Treated", 
                               Tank == "T6" ~ "Untreated",
                               Tank == "T7" ~ "Untreated",
                               Tank == "T8" ~ "Tank 8")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ggplot(., aes(x=Treatment, y=daily_range, color=Tank)) +
  geom_boxplot() + 
  theme_classic() +
  labs(y="Daily Range of Temperature (ºC)") +
  theme(text = element_text(size = 15)) 

plot_grid(p1, p2, p3, nrow = 1, align = 'v', rel_widths = c(1, 1, 1))
#ggsave("combined_maxminrange.pdf", width = 15, height = 5)
```

Test for significant differences 
```{r}
maxminrange_summarytable <- temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable", 
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control",
                               Tank == "T8" ~ "tank_8")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ungroup()

str(maxminrange_summarytable)

maxminrange_summarytable <- maxminrange_summarytable %>% 
  mutate(across(c(Tank,Date,Treatment,Species), as.factor))
```

Daily max
```{r}
maxminrange_summarytable %>% 
  aov(daily_max ~ Tank, data = .) -> model1

summary(model1)
#tank is significant

qqnorm(residuals(model1)) 
qqline(residuals(model1)) # not normal

#Non parametric test
maxminrange_summarytable %>% 
  kruskal_test(., daily_max ~ Tank) #significant

dunn.test(maxminrange_summarytable$daily_max, maxminrange_summarytable$Tank, method = "bonferroni")
# tank 5 (True treatment tank) is significantly different from all tanks

#dunn's test assumption is that distribution of groups is similar
ggqqplot(maxminrange_summarytable, "daily_max", facet.by = "Tank")
```

Daily min
```{r}
maxminrange_summarytable %>% 
  aov(daily_min ~ Tank, data = .) -> model2

summary(model2)
#tank is significant

qqnorm(residuals(model2)) 
qqline(residuals(model2)) # not normal

#Non parametric test
maxminrange_summarytable %>% 
  kruskal_test(., daily_min ~ Tank) #significant

dunn.test(maxminrange_summarytable$daily_min, maxminrange_summarytable$Tank, method = "bonferroni")
# tank 5 (True treatment tank) is significantly different from all tanks

#dunn's test assumption is that distribution of groups is similar
ggqqplot(maxminrange_summarytable, "daily_min", facet.by = "Tank")
```


Daily range
```{r}
maxminrange_summarytable %>% 
  aov(daily_range ~ Tank, data = .) -> model3

summary(model3)
#tank is significant

qqnorm(residuals(model3)) 
qqline(residuals(model3)) # not normal

#Non parametric test
maxminrange_summarytable %>% 
  kruskal_test(., daily_range ~ Tank) #significant

dunn.test(maxminrange_summarytable$daily_range, maxminrange_summarytable$Tank, method = "bonferroni")
# tank 5 (True treatment tank) is significantly different from all tanks

#dunn's test assumption is that distribution of groups is similar
ggqqplot(maxminrange_summarytable, "daily_range", facet.by = "Tank")
```



# CBASS (April 22)
```{r}
CBASS %>% 
  dplyr::select(Time:`8`) %>% 
  pivot_longer(`1`:`8`, names_to="Tank", values_to ="Temp") %>% 
  ggplot(data = .,
       aes(x = Time, y = Temp, color = Tank)) +
  geom_point() + 
  theme_classic() +
  facet_wrap(~Tank) +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Date")
```

```{r}
CBASS %>% 
  dplyr::select(Time:`8`) %>% 
  pivot_longer(`1`:`8`, names_to="Tank", values_to ="Temp") %>% 
  ggplot(data = .,
       aes(x = Time, y = Temp, color = Tank)) +
  geom_line() + 
  theme_classic() +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Hour") +
  theme(text = element_text(size = 20)) 
#ggsave("CBASStemps.pdf", width = 10)
```

summarise tank temps 
```{r}
CBASS %>% 
  dplyr::select(Time:`8`) %>% 
  pivot_longer(`1`:`8`, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Time = hms(Time)) %>% 
  filter(Time > "14 H 30 M 0 S" & Time < "17 H 30 M 0 S") %>% 
  ggplot(., aes(x=Time, y=Temp, fill = Tank)) +
  geom_line() +
  scale_x_time()

CBASS %>% 
  dplyr::select(Time:`8`) %>% 
  pivot_longer(`1`:`8`, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Time = hms(Time)) %>% 
  filter(Time > "14 H 30 M 0 S" & Time < "17 H 30 M 0 S") %>% 
  group_by(Tank) %>% 
  summarise(count = n(), mean_temp = mean(Temp), sd = sd(Temp))
```




