---
title: "immunecellassay"
author: "allyson_demerlis"
date: "2024-08-16"
output: html_document
---

# Load libraries and datasets
```{r}
library(tidyverse)
library(plotrix)
library(ggpubr)
library(rstatix)
library(cowplot)
library(dunn.test)
library(Rmisc)
library(plyr)
library(dplyr)
library(lme4)
library(emmeans)
library(modelsummary)
library(car)
library(multcomp)

percent_cells_df <- read_csv("immunecellsdata.csv")

percent_cells_df %>% 
  drop_na() %>% #one technical replicate is NA for one coral ID
  group_by(num_days, Species, Genotype, ID, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perID = mean(Percent)) %>% 
  ungroup() %>% 
  dplyr::select(Species:ID, Tank:num_days, mean_replicate_percent_perID) %>% 
  distinct() -> comb_tech_reps
```

# boxplot of percent change over treatment
```{r}
#taking the mean percent change per colony per tank, so the sample size for each colony in each treatment is only 2. because of this, stats cannot be performed at the colony level.
plot1<-comb_tech_reps %>% 
  group_by(num_days, Tank, Species, Genotype, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perGenet = mean(mean_replicate_percent_perID)) %>% 
  dplyr::select(Species, Genotype, Tank, Treatment, num_days, mean_replicate_percent_perGenet) %>% 
  distinct() %>% 
  filter(num_days == "0" | num_days == "28") %>% 
  pivot_wider(names_from = "num_days", values_from = "mean_replicate_percent_perGenet") %>% 
  mutate(relative_activity = ((`28`-`0`)/`0`)*100) %>% #T1-T0/T0*100
  ggplot(., aes(x=Species, y=relative_activity, fill= Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species,ncol = 1, nrow = 2, scales="free_x") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  ylab("% Change Immune Cell Activity")+
  theme_classic() +
  theme(text = element_text(size = 15)) +
  theme(legend.position="none")
```

## linear model stats
```{r}
# try including colony as a random effect since there isn't enough data for it to be a fixed effect
comb_tech_reps %>% 
  group_by(num_days, Tank, Species, Genotype, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perGenet = mean(mean_replicate_percent_perID)) %>% 
  dplyr::select(Species, Genotype, Tank, Treatment, num_days, mean_replicate_percent_perGenet) %>% 
  distinct() %>% 
  filter(num_days == "0" | num_days == "28") %>% 
  pivot_wider(names_from = "num_days", values_from = "mean_replicate_percent_perGenet") %>% 
  mutate(relative_activity = ((`28`-`0`)/`0`)*100) -> percent_change

percent_change %>% 
  filter(Species == "Acer") -> Acer_percent_change

percent_change %>% 
  filter(Species == "Pcli") -> Pcli_percent_change

Acer_model <- lmer(relative_activity ~ Treatment + (1|Genotype) + (1|Tank), data = Acer_percent_change)
summary(Acer_model)
anova(Acer_model)

Pcli_model <- lmer(relative_activity ~ Treatment + (1|Genotype) + (1|Tank), data = Pcli_percent_change)
summary(Pcli_model)
anova(Pcli_model)
```

#### Model fitting and assumptions diagnostic

Acer
```{r}
x = residuals(Acer_model)
plot(fitted(Acer_model), x) 
leveneTest(x ~ Treatment, data=Acer_percent_change, center=mean) # formal statistical test for homogenity of variance
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test

Anova(Acer_model,  type = "II", test.statistic = "F") #you need to manually specify Satterthwaite's method, because otherwise it will give you the Wald/Chi-squared approximation and that is only appropriate for binomial or poisson distributions. Satterthwaite's method is for normal/gaussian
capture.output(Anova(Acer_model,  type = "II", test.statistic = "F"), file = "Acer_model_ANOVA.csv")
```


Pcli
```{r}
x = residuals(Pcli_model)
plot(fitted(Pcli_model), x) 
leveneTest(x ~ Treatment, data=Pcli_percent_change, center=mean) # formal statistical test for homogeneity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test 

Anova(Pcli_model, type = "II", test.statistic = "F") #you need to manually specify Satterthwaite's method, because otherwise it will give you the Wald/Chi-squared approximation and that is only appropriate for binomial or poisson distributions. Satterthwaite's method is for normal/gaussian

capture.output(Anova(Pcli_model, type = "II", test.statistic = "F"), file = "Pcli_model_ANOVA.csv")
```


# percent immune activity over the entire experiment period (mean technical replicate)

## species x treatment
```{r}
plot2<-percent_cells_df %>% 
  drop_na() %>% #one technical replicate is NA for one coral ID
  group_by(num_days, Species, Genotype, ID, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perID = mean(Percent)) %>% 
  ungroup() %>% 
  dplyr::select(Species:ID, Tank:num_days, mean_replicate_percent_perID) %>% 
  ggplot(., aes(x=num_days, y=mean_replicate_percent_perID, color = Treatment)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species,ncol = 1, nrow = 2) +
  ylab("% Immune Cell Activity")+
  theme_classic() +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) +
  labs(x="Days in Treatment") +
  theme(legend.position="none")
#ggsave("immunecellactivityovertime_bothspecies.pdf", width = 10, height = 7)
```

## species x colony x treatment
```{r}
plot3<-percent_cells_df %>% 
  drop_na() %>% #one technical replicate is NA for one coral ID
  group_by(num_days, Species, Genotype, ID, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perID = mean(Percent)) %>% 
  ungroup() %>% 
  dplyr::select(Species:ID, Tank:num_days, mean_replicate_percent_perID) %>% 
  ggplot(., aes(x=num_days, y=mean_replicate_percent_perID, color = Treatment)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species + Genotype) +
  ylab("% Immune Cell Activity")+
  theme_classic() +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) +
  labs(x="Days in Treatment")+
  theme(legend.position="none")
#ggsave("immunecellactivityovertime_bothspecies_genotypes.pdf", width = 10, height = 7)
```
  
## combined plot
```{r}
plot_grid(plot1, plot2, plot3, nrow = 1, rel_widths = c(1, 1, 2))
ggsave("percentimmunecellactivity_boxplot_linear.pdf", width = 12, height = 5)
```

  
# Mixed effects GLMM (fixed: treatment, colony, timepoint; random: tank)

First, make the dataset with percentages as the mean technical replicates per coral fragment (r1, r2, and r3)
```{r}
percent_cells_df %>% 
  drop_na() %>% #one technical replicate is NA for one coral ID
  group_by(num_days, Species, Genotype, ID, Treatment) %>% 
  dplyr::mutate(mean_replicate_percent_perID = mean(Percent)) %>% 
  ungroup() %>% 
  dplyr::select(Species:ID, Tank:num_days, mean_replicate_percent_perID) -> immmunecells_meanreps

immmunecells_meanreps %>% 
  dplyr::mutate(across(c(Species, Genotype, ID, Tank, Treatment, num_days), as.factor)) -> immmunecells_meanreps

immmunecells_meanreps %>% 
  filter(Species == "Acer") %>% 
  distinct() -> immmunecells_meanreps_Acer

immmunecells_meanreps %>% 
  filter(Species == "Pcli") %>% 
  distinct() -> immmunecells_meanreps_Pcli
```


```{r}
#although this was measured over time, it is not repeated measures because different corals were measured at each time point. so coral ID is not included in this formula
Acer_model <- lmer(mean_replicate_percent_perID~Treatment*Genotype*num_days + (1|Tank), data = immmunecells_meanreps_Acer)
summary(Acer_model)
Anova(Acer_model, type = "II", test.statistic = "F")

Pcli_model <- lmer(mean_replicate_percent_perID~Treatment*Genotype*num_days + (1|Tank), data = immmunecells_meanreps_Pcli)
summary(Pcli_model)
Anova(Pcli_model, type = "II", test.statistic = "F")
```

## Model fitting and assumptions diagnostic

Acer
```{r}
x = residuals(Acer_model)
plot(fitted(Acer_model), x) 
leveneTest(x ~ Treatment*Genotype*num_days, data=immmunecells_meanreps_Acer, center=mean) # formal statistical test for homogeneity of variance #very significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # not significant

#log transformation helped
Acer_model_2 <- lmer(log(mean_replicate_percent_perID)~Treatment*Genotype*num_days + (1|Tank), data = Acer_immunecells_noout)
x = residuals(Acer_model_2)
plot(fitted(Acer_model_2), x) 
leveneTest(x ~ Treatment*Genotype*num_days, data=Acer_immunecells_noout, center=mean) 
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test

Anova(Acer_model_2, type = "II", test.statistic = "F")
capture.output(Anova(Acer_model_2, type = "II", test.statistic = "F"), file = "Acer_immunecells_model_ANOVA.csv")
```

Pcli
```{r}
x = residuals(Pcli_model)
plot(fitted(Pcli_model), x) 
leveneTest(x ~ Treatment*Genotype*num_days, data=immmunecells_meanreps_Pcli, center=mean) 
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test

Anova(Pcli_model, type = "II", test.statistic = "F")
capture.output(Anova(Pcli_model, type = "II", test.statistic = "F"), file = "Pcli_immunecells_model_ANOVA.csv")
```

## Statistical pairwise comparisons

Acer
```{r}
#colony + treatment + timepoint
Acer_emms <- emmeans(Acer_model_2, pairwise ~ Treatment*Genotype*num_days)
pairs(Acer_emms)
capture.output(pairs(Acer_emms), file = "Acer_Tukey_colonytreatmenttimepoint.csv")
```

Pcli
```{r}
#colony + treatment + timepoint
Pcli_emms <- emmeans(Pcli_model, pairwise ~ Treatment*Genotype*num_days)
pairs(Pcli_emms)
capture.output(pairs(Pcli_emms), file = "Pcli_Tukey_colonytreatmenttimepoint.csv")
```




