---
title: "immunecellassay"
author: "allyson_demerlis"
date: "2023-02-28"
output: html_document
---
Load libraries and datasets
```{r}
library(tidyverse)
adjusted_cells_data <- read_csv("adjusted_cell_proportions.csv")
percent_cells_metadata <- read_csv("stresshardening_master.csv") #this only has replicate 1 in it
#the controls for normalization were taken from replicate 1. 


```
Raw percent comparison by treatment over time
(first technical replicate only)
```{r}
#only look at first technical replicate
percent_cells_metadata %>%
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = "free") + 
  theme_classic() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34"))

```

Raw percent comparison by treatment over time
(average of all technical replicates)
```{r}
percent_cells_metadata %>%
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  group_by(TimePoint, Species, Genotype, Treatment) %>% 
  select(!c("Tank","Average Control (for that given genotype and timepoint)", "Adjusted value (Percent -- Average Control)")) %>% 
  mutate(average_percent = mean(Percent)) %>% 
  select(!Percent) %>% 
  distinct() %>% 
  arrange(TimePoint, Species, Genotype)
  ggplot(., aes(x=TimePoint, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = "free") + 
  theme_classic() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34"))

```




Raw comparison by treatment x genotype over time
```{r}

```



we can't compare percentages across treatments and time because of several limitations:
1) we don't know what a "normal" level of phagocytic cells is, and the "normal" range for individuals can be quite large.
2) the cytometer itself may influence time point
so what we want to do is use the control group as the "baseline" and calculate the percent difference of the variable group to the control
#and we can do this for each time point







# subset to compare to Lys's data 
```{r}
adjusted_cells_data %>% 
  filter(Species == "Pcli") %>% 
  filter(TimePoint == "T0") %>%  #only look at before they were temp treated because in Lys's experiment she combined treated and untreated coral fragments
  dplyr::rename(Avg_Control = `Average Control (for that given genotype and timepoint)`,
                Adj_percent = `Adjusted value (Percent -- Average Control)`)
  group_by(Genotype) %>% 
  dplyr::summarise(mean_percent = mean(Adj_percent), sd = sd(Adj_percent), count = n())
```


```{r}
#only look at first technical replicate
percent_cells_metadata %>%
  filter(Replicate == "r1") %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = "free") + 
  theme_classic() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34"))

#we can't compare percentages across treatments and time because of several limitations:
# 1) we don't know what a "normal" level of phagocytic cells is, and the "normal" range for individuals can be quite large.
# 2) the cytometer itself may influence time point

```


```{r}
#without follow up (because that wasn't during the treatment period, it was like a month later)

percent_cells_metadata %>%
  filter(Replicate == "r1") %>% 
  filter(!TimePoint=="FollowUp") %>% 
  mutate(TimePoint = case_when(TimePoint == "T0" ~ "Day 1",
                               TimePoint == "T1" ~ "Day 7",
                               TimePoint == "T2" ~ "Day 28")) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "Day 1", "Day 7", "Day 28")) %>% 
  ggplot(., aes(x=TimePoint, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = "free") + 
  theme_classic() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 15)) 
```

```{r}
#so what we want to do is use the control group as the "baseline" and calculate the percent difference of the variable group to the control
#and we can do this for each time point

#maybe try doing a ratio? like variable / control?

percent_cells_metadata %>%
  filter(Replicate == "r1") %>% 
  filter(!TimePoint=="FollowUp") %>% 
  group_by(TimePoint, Species, Treatment) %>% 
  summarise(avg_percent = mean(Percent)) %>% 
  pivot_wider(names_from = "Treatment", values_from = "avg_percent") %>% 
  mutate(TimePoint = case_when(TimePoint == "T0" ~ "Day 1",
                               TimePoint == "T1" ~ "Day 7",
                               TimePoint == "T2" ~ "Day 28")) %>% 
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "Day 1", "Day 7", "Day 28")) %>% 
  mutate("VC_Ratio" = Variable/Control) %>% 
  mutate(VC_Ratio = VC_Ratio - 1) %>% 
  mutate(color_assignment = case_when(VC_Ratio > 0 ~ "Variable",
                                      VC_Ratio < 0 ~ "Control"
                                      )) %>% 
  ggplot(., aes(x=TimePoint, y=VC_Ratio, fill = color_assignment)) +
  geom_col(position = position_dodge()) + 
  theme_classic() +
  facet_wrap(~Species) +
  labs(x = "Time Point", y = "Ratio of Average Phagocytosis Percentage", fill = "Treatment") +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 15))
  
```


```{r}
percent_cells_metadata %>%
  filter(Replicate == "r1") %>% 
  filter(!TimePoint=="FollowUp") %>% 
  select(TimePoint, Species, Percent, Treatment) %>% 
  pivot_wider(names_from = "TimePoint", values_from = "Percent")
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  mutate("VC_Ratio" = Variable/Control)
```

using Grace/Benyamin's method of adjusting values based on the average control value for that timepoint+genotype
```{r}
jitter <- position_jitter(width = 0.1, height = 0.1)

adjusted_cells_data %>% 
  group_by(TimePoint, Species, Genotype, Treatment) %>% 
  summarise(avg_percentage = mean(`Adjusted value (Percent -- Average Control)`), sd = sd(`Adjusted value (Percent -- Average Control)`)) %>% 
  filter(Species == "Acer") %>% 
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=avg_percentage, fill = Treatment)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = avg_percentage - sd, ymax = avg_percentage + sd), width =.2, position = jitter) +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic() +
  facet_wrap(~Genotype)

```

```{r}
adjusted_cells_data %>% 
  group_by(TimePoint, Species, Genotype, Treatment) %>% 
  summarise(avg_percentage = mean(`Adjusted value (Percent -- Average Control)`), sd = sd(`Adjusted value (Percent -- Average Control)`)) %>% 
  ungroup() %>% 
  filter(Species == "Pcli") %>% 
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=avg_percentage, fill = Treatment)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = avg_percentage - sd, ymax = avg_percentage + sd), width =.2, position = jitter) +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic() +
  facet_wrap(~Genotype)
```

```{r}
adjusted_cells_data %>% 
  group_by(TimePoint, Species, Genotype, Treatment) %>% 
  filter(Species == "Acer") %>% 
  filter(!TimePoint == "FollowUp") %>% 
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=`Adjusted value (Percent -- Average Control)`, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic() +
  facet_wrap(~Genotype) +
  theme(text = element_text(size = 13))
```


```{r}
adjusted_cells_data %>% 
  group_by(TimePoint, Species, Genotype, Treatment) %>% 
  filter(!TimePoint=="FollowUp") %>% 
  filter(Species == "Pcli") %>% 
  mutate(TimePoint = as.factor(TimePoint)) %>% 
  mutate(TimePoint = fct_relevel(TimePoint, 
            "T0", "T1", "T2", 
            "FollowUp")) %>% 
  ggplot(., aes(x=TimePoint, y=`Adjusted value (Percent -- Average Control)`, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme_classic() +
  facet_wrap(~Genotype) +
  theme(text = element_text(size = 13))
```



Stats (attempt to use lmer)
```{r}
adjusted_cells_data %>% 
  filter(Species == "Acer") -> Acer_FACS_df

#model 1 = full model
Acer_FACS_GLMM_full <- lmer(`Adjusted value (Percent -- Average Control)` ~ Treatment*TimePoint + (1|Genotype) + (1|Tank), data=Acer_FACS_df)
summary(Acer_FACS_GLMM_full)

qqnorm(resid(Acer_FACS_GLMM_full))
qqline(resid(Acer_FACS_GLMM_full))

Acer_FACS_aov <- aov(`Adjusted value (Percent -- Average Control)` ~ Treatment*TimePoint*Genotype, data=Acer_FACS_df)
summary(Acer_FACS_aov)

qqnorm(resid(Acer_FACS_aov))
```

```{r}
adjusted_cells_data %>% 
  filter(Species == "Pcli") -> Pcli_FACS_df

Pcli_FACS_aov <- aov(`Adjusted value (Percent -- Average Control)` ~ Treatment*TimePoint*Genotype, data=Pcli_FACS_df)
summary(Pcli_FACS_aov)

qqnorm(resid(Pcli_FACS_aov))

pivot_longer(cols = `Trypan )

```

