---
title: "technicalreplicates"
author: "allyson_demerlis"
date: "2024-06-18"
output: html_document
---

# Load libraries and datasets
```{r}
library(tidyverse)
library(plotrix)
library(ggpubr)
library(rstatix)
library(cowplot)
library(dunn.test)
library(Rmisc)
library(plyr)
library(dplyr)

percent_cells_df <- read_csv("Grace_percentcells_alldata.csv") 

percent_cells_df%>% 
    mutate(Treatment = case_when(Treatment == "Control" ~ "Untreated",
                               Treatment == "Variable" ~ "Treated")) -> percent_cells_df

#there's one outlier to remove right off the bat: T1	Pcli	B	P9	r2	Percent=111.50	T1	Untreated
percent_cells_df %>% 
  filter(!(TimePoint=="T1" & Species == "Pcli" & Genotype == "B" & Treatment == "Untreated" & ID == "P9" & Replicate == "r2")) -> percent_cells_df

#replicates are technical replicates of the same coral fragment
#time points:
#T0 = initial pre-treatment (March 22)
#T1 = one week into treatment (March 30)
#T2 = end of treatment (April 20)
#follow-up = post CBASS and treatment follow-up (June 1)

percent_cells_df %>% 
  mutate(num_days = case_when(TimePoint == "T0" ~ "0",
                              TimePoint == "T1" ~ "7",
                              TimePoint == "T2" ~ "28",
                              TimePoint == "FollowUp" ~ "71")) %>% 
  mutate(num_days = as.factor(num_days)) %>% 
  mutate(num_days = fct_relevel(num_days, "0", "7", "28", "71")) -> percent_cells_df
```


# Raw percent at each time point (compare reps 1-3) and testing if significant differences between technical replicates 

None of the technical replicates are significantly different than one another within a given species:genotype:treatment. So I feel confident taking the mean percent across the three technical replicates and using that moving forward.

## T0
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0") %>% 
  ggplot(., aes(x=Replicate, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species+Genotype) + 
  theme_classic() +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```
### Statistics

#### Acer BC-8b 

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Acer" & Genotype == "BC-8b") -> T0_Acer_BC8b_df

str(T0_df)

#make factors
T0_Acer_BC8b_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Acer_BC8b_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Acer_BC8b_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Acer_BC8b_df)
T0_Acer_BC8b_model_metrics <- augment(T0_Acer_BC8b_model)
plot(T0_Acer_BC8b_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Acer_BC8b_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Acer_BC8b_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T0_Acer_BC8b_df)

summary(aov(Percent ~ Treatment*Replicate, data = T0_Acer_BC8b_df)) #replicate is not significant. wooooo
```

#### Acer MB-B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Acer" & Genotype == "MB-B") -> T0_Acer_MBB_df

#make factors
T0_Acer_MBB_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Acer_MBB_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Acer_MBB_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Acer_MBB_df)
T0_Acer_MBB_model_metrics <- augment(T0_Acer_MBB_model)
plot(T0_Acer_MBB_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Acer_MBB_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Acer_MBB_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T0_Acer_MBB_df)

summary(aov(Percent ~ Treatment*Replicate, data = T0_Acer_MBB_df)) #replicate is not significant. wooooo
```

#### Acer SI-C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Acer" & Genotype == "SI-C") -> T0_Acer_SIC_df

#make factors
T0_Acer_SIC_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Acer_SIC_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Acer_SIC_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Acer_SIC_df)
T0_Acer_SIC_model_metrics <- augment(T0_Acer_SIC_model)
plot(T0_Acer_SIC_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Acer_SIC_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Acer_SIC_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T0_Acer_SIC_df)

summary(aov(Percent ~ Treatment*Replicate, data = T0_Acer_SIC_df)) #replicate is not significant. wooooo
```

#### Pcli A

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Pcli" & Genotype == "A") -> T0_Pcli_A_df

#make factors
T0_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Pcli_A_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Pcli_A_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Pcli_A_df)
T0_Pcli_A_model_metrics <- augment(T0_Pcli_A_model)
plot(T0_Pcli_A_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Pcli_A_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Pcli_A_model_metrics) #significant

#make interaction term
T0_Pcli_A_df$Treatment_Replicate <- interaction(T0_Pcli_A_df$Treatment, T0_Pcli_A_df$Replicate)

#run Welch's anova
oneway.test(Percent ~ Treatment_Replicate, data = T0_Pcli_A_df, var.equal = FALSE) #not significant
```


#### Pcli B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Pcli" & Genotype == "B") -> T0_Pcli_B_df

#make factors
T0_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Pcli_B_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Pcli_B_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Pcli_B_df)
T0_Pcli_B_model_metrics <- augment(T0_Pcli_B_model)
plot(T0_Pcli_B_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Pcli_B_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Pcli_B_model_metrics) #significant

#make interaction term
T0_Pcli_B_df$Treatment_Replicate <- interaction(T0_Pcli_B_df$Treatment, T0_Pcli_B_df$Replicate)

#run Welch's anova
oneway.test(Percent ~ Treatment_Replicate, data = T0_Pcli_B_df, var.equal = FALSE) #not significant
```


#### Pcli C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T0" & Species == "Pcli" & Genotype == "C") -> T0_Pcli_C_df

#make factors
T0_Pcli_C_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T0_Pcli_C_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T0_Pcli_C_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T0_Pcli_C_df)
T0_Pcli_C_model_metrics <- augment(T0_Pcli_C_model)
plot(T0_Pcli_C_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T0_Pcli_C_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T0_Pcli_C_model_metrics) #significant

#make interaction term
T0_Pcli_C_df$Treatment_Replicate <- interaction(T0_Pcli_C_df$Treatment, T0_Pcli_C_df$Replicate)

#run Welch's anova
oneway.test(Percent ~ Treatment_Replicate, data = T0_Pcli_C_df, var.equal = FALSE) #not significant
```


## T1
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1") %>% 
  ggplot(., aes(x=Replicate, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species+Genotype) + 
  theme_classic() +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```

### Statistics

#### Acer BC-8b 

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Acer" & Genotype == "BC-8b") -> T1_Acer_BC8b_df

#make factors
T1_Acer_BC8b_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Acer_BC8b_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Acer_BC8b_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Acer_BC8b_df)
T1_Acer_BC8b_model_metrics <- augment(T1_Acer_BC8b_model)
plot(T1_Acer_BC8b_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Acer_BC8b_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Acer_BC8b_model_metrics) #significant

#make interaction term
T1_Acer_BC8b_df$Treatment_Replicate <- interaction(T1_Acer_BC8b_df$Treatment, T1_Acer_BC8b_df$Replicate)

#run Welch's anova
oneway.test(Percent ~ Treatment_Replicate, data = T1_Acer_BC8b_df, var.equal = FALSE) #not significant
```

#### Acer MB-B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Acer" & Genotype == "MB-B") -> T1_Acer_MBB_df

#make factors
T1_Acer_MBB_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Acer_MBB_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Acer_MBB_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Acer_MBB_df)
T1_Acer_MBB_model_metrics <- augment(T1_Acer_MBB_model)
plot(T1_Acer_MBB_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Acer_MBB_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Acer_MBB_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T1_Acer_MBB_df)

summary(aov(Percent ~ Treatment*Replicate, data = T1_Acer_MBB_df)) #replicate is not significant. wooooo
```

#### Acer SI-C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Acer" & Genotype == "SI-C") -> T1_Acer_SIC_df

#make factors
T1_Acer_SIC_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Acer_SIC_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Acer_SIC_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Acer_SIC_df)
T1_Acer_SIC_model_metrics <- augment(T1_Acer_SIC_model)
plot(T1_Acer_SIC_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Acer_SIC_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Acer_SIC_model_metrics) #significant

#make interaction term
T1_Acer_SIC_df$Treatment_Replicate <- interaction(T1_Acer_SIC_df$Treatment, T1_Acer_SIC_df$Replicate)

#run kruskal wallis test
kruskal_test(Percent ~ T1_Acer_SIC_df$Treatment_Replicate, data = T1_Acer_SIC_df) #not significant
```

#### Pcli A

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Pcli" & Genotype == "A") -> T1_Pcli_A_df

#make factors
T1_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Pcli_A_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Pcli_A_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Pcli_A_df)
T1_Pcli_A_model_metrics <- augment(T1_Pcli_A_model)
plot(T1_Pcli_A_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Pcli_A_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Pcli_A_model_metrics) #significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T1_Pcli_A_df)

summary(aov(Percent ~ Treatment*Replicate, data = T1_Pcli_A_df)) #replicate is not significant. wooooo
```


#### Pcli B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Pcli" & Genotype == "B") -> T1_Pcli_B_df

#make factors
T1_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Pcli_B_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Pcli_B_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Pcli_B_df)
T1_Pcli_B_model_metrics <- augment(T1_Pcli_B_model)
plot(T1_Pcli_B_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Pcli_B_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Pcli_B_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T1_Pcli_B_df)

summary(aov(Percent ~ Treatment*Replicate, data = T1_Pcli_B_df)) #replicate is not significant. wooooo
```


#### Pcli C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T1" & Species == "Pcli" & Genotype == "C") -> T1_Pcli_C_df

#make factors
T1_Pcli_C_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T1_Pcli_C_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T1_Pcli_C_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T1_Pcli_C_df)
T1_Pcli_C_model_metrics <- augment(T1_Pcli_C_model)
plot(T1_Pcli_C_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T1_Pcli_C_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T1_Pcli_C_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T1_Pcli_C_df)

summary(aov(Percent ~ Treatment*Replicate, data = T1_Pcli_C_df)) #replicate is not significant. wooooo
```




## T2
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2") %>% 
  ggplot(., aes(x=Replicate, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species+Genotype) + 
  theme_classic() +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```


### Statistics

#### Acer BC-8b 

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Acer" & Genotype == "BC-8b") -> T2_Acer_BC8b_df

#make factors
T2_Acer_BC8b_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Acer_BC8b_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Acer_BC8b_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Acer_BC8b_df)
T2_Acer_BC8b_model_metrics <- augment(T2_Acer_BC8b_model)
plot(T2_Acer_BC8b_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Acer_BC8b_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Acer_BC8b_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Acer_BC8b_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Acer_BC8b_df)) #replicate is not significant. wooooo
```

#### Acer MB-B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Acer" & Genotype == "MB-B") -> T2_Acer_MBB_df

#make factors
T2_Acer_MBB_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Acer_MBB_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Acer_MBB_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Acer_MBB_df)
T2_Acer_MBB_model_metrics <- augment(T2_Acer_MBB_model)
plot(T2_Acer_MBB_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Acer_MBB_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Acer_MBB_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Acer_MBB_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Acer_MBB_df)) #replicate is not significant. wooooo
```

#### Acer SI-C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Acer" & Genotype == "SI-C") -> T2_Acer_SIC_df

#make factors
T2_Acer_SIC_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Acer_SIC_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Acer_SIC_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Acer_SIC_df)
T2_Acer_SIC_model_metrics <- augment(T2_Acer_SIC_model)
plot(T2_Acer_SIC_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Acer_SIC_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Acer_SIC_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Acer_SIC_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Acer_SIC_df)) #replicate is not significant. wooooo
```

#### Pcli A

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Pcli" & Genotype == "A") -> T2_Pcli_A_df

#make factors
T2_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Pcli_A_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Pcli_A_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Pcli_A_df)
T2_Pcli_A_model_metrics <- augment(T2_Pcli_A_model)
plot(T2_Pcli_A_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Pcli_A_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Pcli_A_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Pcli_A_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Pcli_A_df)) #replicate is not significant. wooooo
```


#### Pcli B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Pcli" & Genotype == "B") -> T2_Pcli_B_df

#make factors
T2_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Pcli_B_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Pcli_B_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Pcli_B_df)
T2_Pcli_B_model_metrics <- augment(T2_Pcli_B_model)
plot(T2_Pcli_B_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Pcli_B_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Pcli_B_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Pcli_B_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Pcli_B_df)) #replicate is not significant. wooooo
```


#### Pcli C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "T2" & Species == "Pcli" & Genotype == "C") -> T2_Pcli_C_df

#make factors
T2_Pcli_C_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> T2_Pcli_C_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
T2_Pcli_C_model <- lm(Percent ~ Treatment*Replicate + Tank, data = T2_Pcli_C_df)
T2_Pcli_C_model_metrics <- augment(T2_Pcli_C_model)
plot(T2_Pcli_C_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(T2_Pcli_C_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = T2_Pcli_C_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = T2_Pcli_C_df)

summary(aov(Percent ~ Treatment*Replicate, data = T2_Pcli_C_df)) #replicate is not significant. wooooo
```



## FollowUp
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp") %>% 
  ggplot(., aes(x=Replicate, y=Percent, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species+Genotype) + 
  theme_classic() +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```



### Statistics

#### Acer BC-8b 

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Acer" & Genotype == "BC-8b") -> FollowUp_Acer_BC8b_df

#make factors
FollowUp_Acer_BC8b_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Acer_BC8b_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Acer_BC8b_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Acer_BC8b_df)
FollowUp_Acer_BC8b_model_metrics <- augment(FollowUp_Acer_BC8b_model)
plot(FollowUp_Acer_BC8b_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Acer_BC8b_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Acer_BC8b_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_BC8b_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_BC8b_df)) #replicate is not significant. wooooo
```

#### Acer MB-B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Acer" & Genotype == "MB-B") -> FollowUp_Acer_MBB_df

#make factors
FollowUp_Acer_MBB_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Acer_MBB_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Acer_MBB_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Acer_MBB_df)
FollowUp_Acer_MBB_model_metrics <- augment(FollowUp_Acer_MBB_model)
plot(FollowUp_Acer_MBB_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Acer_MBB_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Acer_MBB_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_MBB_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_MBB_df)) #replicate is not significant. wooooo
```

#### Acer SI-C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Acer" & Genotype == "SI-C") -> FollowUp_Acer_SIC_df

#make factors
FollowUp_Acer_SIC_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Acer_SIC_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Acer_SIC_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Acer_SIC_df)
FollowUp_Acer_SIC_model_metrics <- augment(FollowUp_Acer_SIC_model)
plot(FollowUp_Acer_SIC_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Acer_SIC_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Acer_SIC_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_SIC_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Acer_SIC_df)) #replicate is not significant. wooooo
```

#### Pcli A

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Pcli" & Genotype == "A") -> FollowUp_Pcli_A_df

#make factors
FollowUp_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Pcli_A_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Pcli_A_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Pcli_A_df)
FollowUp_Pcli_A_model_metrics <- augment(FollowUp_Pcli_A_model)
plot(FollowUp_Pcli_A_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Pcli_A_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Pcli_A_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_A_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_A_df)) #replicate is not significant. wooooo
```


#### Pcli B

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Pcli" & Genotype == "B") -> FollowUp_Pcli_B_df

#make factors
FollowUp_Pcli_A_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Pcli_B_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Pcli_B_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Pcli_B_df)
FollowUp_Pcli_B_model_metrics <- augment(FollowUp_Pcli_B_model)
plot(FollowUp_Pcli_B_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Pcli_B_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Pcli_B_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_B_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_B_df)) #replicate is not significant. wooooo
```


#### Pcli C

do within each species:genotype because you are interested in whether replicates within the species:genotype comparison are significantly different
```{r}
percent_cells_df %>%
  filter(TimePoint == "FollowUp" & Species == "Pcli" & Genotype == "C") -> FollowUp_Pcli_C_df

#make factors
FollowUp_Pcli_C_df %>% 
  mutate_at(vars(TimePoint, Species, Genotype, Replicate, Tank, Treatment), factor) -> FollowUp_Pcli_C_df
```


Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
FollowUp_Pcli_C_model <- lm(Percent ~ Treatment*Replicate + Tank, data = FollowUp_Pcli_C_df)
FollowUp_Pcli_C_model_metrics <- augment(FollowUp_Pcli_C_model)
plot(FollowUp_Pcli_C_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(FollowUp_Pcli_C_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Replicate, data = FollowUp_Pcli_C_model_metrics) #not significant

#run two-way anova
aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_C_df)

summary(aov(Percent ~ Treatment*Replicate, data = FollowUp_Pcli_C_df)) #replicate is not significant. wooooo
```


None of the technical replicates are significantly different than one another within a given species:genotype:treatment. So I feel confident taking the mean percent across the three technical replicates and using that moving forward.