---
title: "survivorship"
author: "allyson_demerlis"
date: "2023-02-20"
output: html_document
---
# Load data and packages
```{r}
library(tidyverse)
library(janitor)
library(rstatix)
library(cowplot)
library(plotrix)

metadata <- readxl::read_xlsx("../metadata.xlsx", sheet = "tidy_data")

str(metadata) #295 corals

# metadata %>% 
#   filter(!Removed_Date == "unknown") -> metadata #244 corals... 

metadata$Removed_Date <- as.numeric(metadata$Removed_Date)
metadata$Removed_Date <- as.Date(metadata$Removed_Date, origin = "1899-12-30")

metadata$mortality_date <- as.numeric(metadata$mortality_date)
metadata$mortality_date <- as.Date(metadata$mortality_date, origin = "1899-12-30")
```

Tidy data 

```{r}
metadata %>% 
  mutate(CBASS_temp = case_when(CBASS_tank == "1" ~ "36",
                               CBASS_tank == "2" ~ "37",
                               CBASS_tank == "3" ~ "32",
                               CBASS_tank == "4" ~ "33",
                               CBASS_tank == "5" ~ "35",
                               CBASS_tank == "6" ~ "30",
                               CBASS_tank == "7" ~ "34",
                               CBASS_tank == "8" ~ "28")) -> tidy_metadata
```


There are two columns of significance for survivorship, Removed_Date and Mortality_Date. At some point, tank 7 become the recovery tank for corals. If the coral had a lot of tissue on it but looked stressed, it was removed from either the CBASS tank or the slow burn heat stress assay (high temp) into the recovery tank at 28ºC. That date of transfer was recorded in the column "Removed_Date". The column "Mortality_Date" refers to corals that were removed from any tank and died. 

```{r}
metadata %>% 
  filter(is.na(Removed_Date) & is.na(mortality_date)) #43 corals don't have either. But if they weren't assigned a heat-stress tank, that should indicate that they were removed/died before 06/15/2022

metadata %>% 
  filter(is.na(Removed_Date) & is.na(mortality_date) & is.na(slowburn_heatstress_tank)) # 5 corals. looks like the four Acer were tissue sampled and the Pcli who knows, it got buoyant weighed on April 21. they probably died sometime after CBASS and before heat stress.
```


# CBASS Mortality

```{r}
tidy_metadata %>%
  filter(Removed_Date < "2022-05-20") %>% 
  arrange(desc(Removed_Date))

#April 26 was the last coral reported with a mortality date until the slow burn heat stress started. So have April 26 be the cut-off point for CBASS mortality plots.

ggplot(tidy_metadata, aes(x=Removed_Date, fill = Treatment)) +
  geom_bar() + 
  facet_wrap(~Species)

ggplot(tidy_metadata, aes(x=mortality_date, fill = Treatment)) +
  geom_bar() + 
  facet_wrap(~Species)

CBASS_mortality <- tidy_metadata %>% 
  dplyr::filter(Removed_Date >= "2022-04-20" & Removed_Date <= "2022-04-26") 
#for this time period, Removed_Date and mortality_date match up so there's no need to use the coalesce function
```


## Number of corals that died due to CBASS
```{r}
tidy_metadata %>% 
  dplyr::filter(Removed_Date >= "2022-04-20" & Removed_Date <= "2022-04-26") %>%
  drop_na(Species, Treatment, Reason_removed) %>% 
  group_by(Reason_removed, Species, Treatment) %>% 
  summarize(count = n()) %>% 
  ggplot(., aes(x=Reason_removed, y=count, fill = Treatment)) +
  geom_col(position = position_dodge()) +
  theme_classic() +
  facet_wrap(~Species)  +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(x="Category", y="Number of corals") + 
  coord_flip()
```

```{r}
CBASS_mortality %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "post-CBASS RTL" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL")) %>% 
  drop_na(Species, Treatment) %>% 
  group_by(Species, Treatment) %>% 
  summarize(count_RTL = n()) -> RTL_species_treatment_count

CBASS_mortality %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "post-CBASS RTL" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL")) %>% 
  drop_na(Species, Treatment) %>% 
  group_by(Species, Treatment, CBASS_temp) %>% 
  summarize(count_RTL = n()) -> RTL_species_treatment_CBASStemps
```


```{r}
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment) %>% 
  summarize(count_total = n()) %>% 
  full_join(., RTL_species_treatment_count) %>% 
  mutate(count_survived = count_total-count_RTL) %>% 
  select(!count_total) %>% 
  pivot_longer(count_RTL:count_survived, names_to="label", values_to="count") %>% 
  ggplot(., aes(x=Treatment, y=count, fill = label)) +
  geom_col() +
  theme_classic() +
  scale_fill_manual(labels=c("RTL", "Survived"), values = c("black", "grey"))  +
  theme(text = element_text(size = 12)) +
  facet_wrap(~Species, scales = "free_x") +
  labs(y="Count", fill = "Category") 
```

```{r}
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment) %>% 
  summarize(count_total = n()) %>% 
  full_join(., RTL_species_treatment_count) %>% 
  mutate(count_survived = count_total-count_RTL) %>% 
  select(!count_total) %>% 
  
  write_csv("RTL_vs_survived_CBASS_speciestreatments.csv")
```


```{r}
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment, CBASS_temp) %>% 
  summarize(count_total = n()) %>% 
  full_join(., RTL_species_treatment_CBASStemps) %>% 
  drop_na(CBASS_temp) %>% 
  mutate(count_RTL = if_else(is.na(count_RTL),0,count_RTL)) %>%
  mutate(count_survived = count_total-count_RTL) %>% 
  mutate(count_survived = if_else(is.na(count_survived),0,count_survived)) %>% 
  select(!count_total) %>% 
  pivot_longer(count_RTL:count_survived, names_to="label", values_to="count") %>% 
  ggplot(., aes(x=CBASS_temp, y=count, fill = label)) +
  geom_col() +
  theme_classic() +
  scale_fill_manual(labels=c("RTL", "Survived"), values = c("black", "grey"))  +
  theme(text = element_text(size = 12)) +
  facet_wrap(~Species*Treatment, scales = "free_x") +
  labs(y="Count", fill = "Category") 
```

```{r}
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment, CBASS_temp) %>% 
  summarize(count_total = n()) %>% 
  full_join(., RTL_species_treatment_CBASStemps) %>% 
  drop_na(CBASS_temp) %>% 
  mutate(count_RTL = if_else(is.na(count_RTL),0,count_RTL)) %>%
  mutate(count_survived = count_total-count_RTL) %>% 
  mutate(count_survived = if_else(is.na(count_survived),0,count_survived)) %>% 
  select(!count_total) %>% 
  write_csv("RTL_vs_survived_CBASS_byCBASStemp.csv")
```

## stats test on counts (not standardized)

```{r}
#chi squared test for comparing mortality numbers
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment) %>% 
  summarize(count_total = n()) %>%
  full_join(., RTL_species_treatment_count) %>% 
  mutate(count_survived = count_total-count_RTL) %>% 
  select(!count_total) %>% 
  unite("Species.Treatment", Species,Treatment, sep = ".") %>% 
  column_to_rownames(var="Species.Treatment") %>% 
  as.data.frame() %>% 
  chisq_test() #not significant

tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment,CBASS_temp) %>% 
  summarize(count_total = n()) %>%
  drop_na(CBASS_temp) %>% 
  full_join(., RTL_species_treatment_CBASStemps) %>% 
  mutate(count_RTL = if_else(is.na(count_RTL),0,count_RTL)) %>%
  mutate(count_survived = count_total-count_RTL) %>% 
  select(!count_total) %>% 
  unite("Species.Treatment.CBASS_temp", Species,Treatment,CBASS_temp, sep = ".") %>% 
  column_to_rownames(var="Species.Treatment.CBASS_temp") %>%  
  as.data.frame() %>% 
  chisq_test() #very significant 
```

## post-hoc test for pairwise comparisons
```{r}
tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment,CBASS_temp) %>% 
  summarize(count_total = n()) %>%
  drop_na(CBASS_temp) %>% 
  full_join(., RTL_species_treatment_CBASStemps) %>% 
  mutate(count_RTL = if_else(is.na(count_RTL),0,count_RTL)) %>%
  mutate(count_survived = count_total-count_RTL) %>% 
  select(!count_total) %>% 
  unite("Species.Treatment.CBASS_temp", Species,Treatment,CBASS_temp, sep = ".") %>% 
  column_to_rownames(var="Species.Treatment.CBASS_temp") %>% 
  as.data.frame() %>% 
  pairwise_prop_test() #%>% 
  # write_csv("pairwise_prop_test_RTLvsSurvivedCounts.csv")
```

## Proportions 

```{r}
p1<- tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment) %>% 
  summarize(count_total = n()) %>% 
  full_join(., RTL_species_treatment_count) %>% 
  mutate(count_survived = count_total-count_RTL) %>% 
  mutate(proportion_survived = count_survived/count_total) %>% 
  mutate(proportion_RTL = 1-proportion_survived) %>% 
  pivot_longer(proportion_survived:proportion_RTL, names_to="label", values_to="proportion") %>% 
  ggplot(., aes(x=Treatment, y=proportion, fill = label)) +
  geom_col() +
  theme_classic() +
  scale_fill_manual(labels=c("RTL", "Survived"), values = c("black", "grey"))  +
  theme(text = element_text(size = 15)) +
  facet_wrap(~Species, scales = "free_x") +
  labs(y="Proportion", fill = "Category")  +
  theme(legend.position="none")

#ggsave("coralmortalityduetoCBASS.pdf", width=8, height=5)
```


```{r}
p2<-tidy_metadata %>% 
  dplyr::filter(!Removed_Date <= "2022-04-20" | is.na(Removed_Date)) %>% #remove everything that was removed before CBASS so you only account for RTL due to CBASS
  group_by(Species, Treatment,CBASS_temp) %>% 
  summarize(count_total = n()) %>% 
  drop_na(CBASS_temp) %>% 
  full_join(., RTL_species_treatment_CBASStemps) %>% 
  mutate(count_RTL = if_else(is.na(count_RTL),0,count_RTL)) %>% 
  mutate(count_survived = count_total-count_RTL) %>% 
  mutate(count_survived = if_else(is.na(count_survived),0,count_survived)) %>% 
  mutate(proportion_survived = count_survived/count_total) %>% 
  mutate(proportion_RTL = 1-proportion_survived) %>% 
  pivot_longer(proportion_survived:proportion_RTL, names_to="label", values_to="proportion") %>%
  ggplot(., aes(x=CBASS_temp, y=proportion, fill = label)) +
  geom_col() +
  theme_classic() +
  scale_fill_manual(labels=c("RTL", "Survived"), values = c("black", "grey"))  +
  theme(text = element_text(size = 12)) +
  facet_wrap(~Species*Treatment, scales = "free_x") +
  labs(y="Proportion", fill = "Category", x = "CBASS Temperature")
#ggsave("coralmortalityduetoCBASS_bytemp.pdf", width = 10, height =6)
```
combine plots
```{r}
plot_grid(p1, p2, ncol=2, rel_widths = c(1.5, 2))
#ggsave("coralmortalityduetoCBASS_bytreatandtemp.pdf", width = 10, height = 5)
```




# Slow-Burn Mortality

Heat-stress at 32ºC was reached on June 25. 
```{r}
metadata %>% 
  filter(!is.na(slowburn_heatstress_tank)) #163 corals

metadata %>% 
  filter(!is.na(slowburn_heatstress_tank)) %>% 
  select(Species:heatstress_recovery_tank) %>% 
  mutate(CBASS_temp = case_when(CBASS_tank == "1" ~ "36",
                               CBASS_tank == "2" ~ "37",
                               CBASS_tank == "3" ~ "32",
                               CBASS_tank == "4" ~ "33",
                               CBASS_tank == "5" ~ "35",
                               CBASS_tank == "6" ~ "30",
                               CBASS_tank == "7" ~ "34",
                               CBASS_tank == "8" ~ "28")) %>% 
  mutate(slowburn= case_when(slowburn_heatstress_tank == "7" ~ "ambient",
                             TRUE ~ "heatstress_32")) -> slowburn_metadata
```
## days at 32ºC
```{r}
slowburn_metadata %>% 
  filter(Removed_Date >= "2022-06-25") %>% 
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>% 
  mutate(days_to_removed = Removed_Date - start_date) %>% 
  ggplot(., aes(x=Treatment, y=days_to_removed, fill = Treatment)) +
  geom_boxplot() +
  facet_wrap(~Species)
```

```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25") %>% 
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>% 
  mutate(days_to_removed = Removed_Date - start_date) %>% 
  mutate(days_to_removed = as.numeric(days_to_removed)) %>% 
  group_by(Species, Treatment) %>% 
  summarise(mean=mean(days_to_removed), sd = sd(days_to_removed)) %>% 
  as.data.frame()
```

```{r}
b1<- slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  filter(!days_to_removed == "NA days") %>% 
  group_by(Species, Treatment) %>% 
  summarise(averagedays = mean(days_to_removed), se = std.error(days_to_removed)) %>%
  ggplot(., aes(x = Treatment, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - se, ymax = averagedays + se), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  facet_wrap(~Species, scales = "free_x") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  labs( y="Number of Days at 32ºC") +
  theme(text = element_text(size = 15)) +
  theme(legend.position="none")

```

```{r}
b2<- slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  filter(!days_to_removed == "NA days") %>% 
  group_by(Species, Treatment, Colony) %>% 
  summarise(averagedays = mean(days_to_removed), se = std.error(days_to_removed)) %>%
  ggplot(., aes(x = Colony, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - se, ymax = averagedays + se), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  facet_wrap(~Species, scales = "free_x") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  labs( y="Number of Days at 32ºC") +
  theme(text = element_text(size = 15))
```

create combined plot
```{r}
plot_grid(b1, b2, rel_widths = c(1,2))
#ggsave("daysat32_barplot_speciesgenet.pdf", width = 12, height = 5)
```


## 2-way anova for species and treatment

Change to factors
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>%
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  select(Species, ID, Colony, Treatment, CBASS_temp, slowburn, days_to_removed) -> slowburn_summary_data

str(slowburn_summary_data)
#make factors: Colony, Puck, Tank, Treatment, Species

slowburn_summary_data %>% 
  mutate_at(vars(Species, ID, Colony, Treatment, CBASS_temp), factor) %>% 
  mutate(days_to_removed = as.numeric(days_to_removed)) -> slowburn_summary_data
```

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
treat_model <- lm(days_to_removed ~ Treatment*Species + Colony, data = slowburn_summary_data)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics) #not significant
```

Run Welch's ANOVA
```{r}
#make interaction term
slowburn_summary_data$Species_Treatment <- interaction(slowburn_summary_data$Species, slowburn_summary_data$Treatment)

#run Welch's anova
oneway.test(days_to_removed ~ Species_Treatment, data = slowburn_summary_data, var.equal = FALSE) #significant

#posthoc test: Games-Howell

slowburn_summary_data %>% 
  ungroup() %>% 
games_howell_test(., days_to_removed ~ Species_Treatment)
```

### looking at just 28C CBASS corals in slowburn
```{r}
slowburn_summary_data %>% 
  filter(CBASS_temp == "28") #6 corals, only Acer

slowburn_metadata %>% 
  filter(Species == "Pseudodiploria clivosa" & CBASS_temp == "28") #none of the Pcli that were at the "ambient' CBASS temperature made it to the slowburn heat stress.
```

##separated by max CBASS temp
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  filter(!days_to_removed == "NA days") %>% 
  group_by(Species, Treatment, CBASS_temp) %>% 
  summarise(averagedays = mean(days_to_removed), se = std.error(days_to_removed)) %>%
  ggplot(., aes(x = CBASS_temp, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - se, ymax = averagedays + se), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  facet_wrap(~Species, scales = "free_x") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  labs(x="CBASS Temperature", y="Number of Days at 32ºC")
``` 

```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  filter(!days_to_removed == "NA days") %>% 
  group_by(Species, Treatment, CBASS_temp, Colony) %>% 
  summarise(averagedays = mean(days_to_removed), se = std.error(days_to_removed)) %>%
  ggplot(., aes(x = CBASS_temp, y = averagedays, fill = Treatment)) + 
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = averagedays - se, ymax = averagedays + se), width =
                  .2, position=position_dodge(.9)) +
  theme_classic() +
  facet_wrap(~Species+Colony, scales = "free_x") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  labs(x="CBASS Temperature", y="Number of Days at 32ºC")
```

## 3-way anova for species, treatment, and CBASS temp

Change to factors
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>%
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Removed_Date = ymd(Removed_Date)) %>%
  mutate(mortality_date = ymd(mortality_date)) %>%
  mutate(end_date = coalesce(Removed_Date, mortality_date)) %>% #if removed_date is NA, then look and use the date from mortality date
  mutate(start_date="2022-06-25") %>% 
  mutate(start_date = ymd(start_date)) %>% 
  mutate(days_to_removed = end_date - start_date) %>% 
  select(Species, ID, Colony, Treatment, CBASS_temp, slowburn, days_to_removed) -> slowburn_summary_data

str(slowburn_summary_data)
#make factors: Colony, Puck, Tank, Treatment, Species

slowburn_summary_data %>% 
  mutate_at(vars(Species, ID, Colony, Treatment, CBASS_temp), factor) %>% 
  mutate(days_to_removed = as.numeric(days_to_removed)) -> slowburn_summary_data
```

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
treat_model <- lm(days_to_removed ~ Treatment*Species*CBASS_temp + Colony, data = slowburn_summary_data)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species*CBASS_temp, data = treat_model_metrics) #not significant
```
Run 3-way ANOVA
```{r}
aov(days_to_removed ~ Treatment*Species*CBASS_temp, data = slowburn_summary_data)

summary(aov(days_to_removed ~ Treatment*Species*CBASS_temp, data = slowburn_summary_data))
 
#capture.output(summary(aov(days_to_removed ~ Treatment*Species*CBASS_temp, data = slowburn_summary_data)), file = "daysuntilremoved_slowburn_anova_table.csv")

TukeyHSD(aov(days_to_removed ~ Treatment*Species*CBASS_temp, data = slowburn_summary_data))
         
#capture.output(TukeyHSD(aov(days_to_removed ~ Treatment*Species*CBASS_temp, data = slowburn_summary_data)), file = "daysuntilremoved_slowburn_anova_tukey.csv")
```


## RTL vs Bleaching
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached"))  %>% 
  
  ggplot(., aes(x=Reason_removed, y=count, fill = Treatment)) +
  geom_col(position = position_dodge()) +
  theme_classic() +
  facet_wrap(~Species)  +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(x="Category", y="Number of corals")
```
## proportion of bleaching vs RTL

```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species,Treatment, Reason_removed) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x=Treatment, y=count, fill = Reason_removed)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Species)
#ggsave("proportion_bleaching_RTL_species.pdf")
```

```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species,Treatment, Colony, Reason_removed) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x=Colony, y=count, fill = Reason_removed)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Species+Treatment, scales = "free_x")
#ggsave("proportion_bleaching_RTL_bycolony.pdf")
```


```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species,Treatment, Reason_removed, CBASS_temp) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x=CBASS_temp, y=count, fill = Reason_removed)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Species+Treatment)
```
## chi squared test for counts of RTL vs bleaching (Species*Treatment)
```{r}
#chi squared test for comparing mortality numbers
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species,Treatment, Reason_removed) %>% 
  summarise(count = n()) %>% 
  drop_na() %>% 
  unite("Species.Treatment", Species,Treatment, sep = ".") %>% 
  pivot_wider(names_from="Reason_removed", values_from = "count") %>% 
  column_to_rownames(var="Species.Treatment") %>% 
  as.data.frame() %>% 
  chisq_test() # significant
```

## post-hoc test for pairwise comparisons (Species*Treatment)
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species,Treatment, Reason_removed) %>% 
  summarise(count = n()) %>% 
  unite("Species.Treatment", Species,Treatment, sep = ".") %>% 
  pivot_wider(names_from="Reason_removed", values_from = "count") %>% 
  column_to_rownames(var="Species.Treatment") %>% 
  as.data.frame() %>% 
  pairwise_prop_test() #%>% 
  #write_csv("pairwise_prop_test_RTLvsbleaching_slowburn.csv")
```


## chi squared test for counts of RTL vs bleaching (Species*Colony*Treatment)
```{r}
#chi squared test for comparing mortality numbers
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species, Colony,Treatment, Reason_removed) %>% 
  summarise(count = n()) %>% 
  drop_na() %>% 
  unite("Species.Colony.Treatment", Species,Colony,Treatment, sep = ".") %>% 
  pivot_wider(names_from="Reason_removed", values_from = "count") %>% 
  mutate(bleached = if_else(is.na(bleached),0,bleached)) %>% 
  mutate(RTL = if_else(is.na(RTL),0,RTL)) %>% 
  column_to_rownames(var="Species.Colony.Treatment") %>% 
  as.data.frame() %>% 
  chisq_test() # significant
```
## post-hoc test for pairwise comparisons (Species*Colony*Treatment)
```{r}
slowburn_metadata %>% 
  filter(slowburn == "heatstress_32") %>% 
  filter(Removed_Date >= "2022-06-25" | is.na(Removed_Date)) %>% 
  mutate(Reason_removed = case_when(Reason_removed == "dead" ~ "RTL",
                             Reason_removed == "RTL" ~ "RTL",
                             Reason_removed == "paling" ~ "bleached", 
                             Reason_removed == "bleaching" ~ "bleached")) %>% 
  group_by(Species, Colony,Treatment, Reason_removed) %>% 
  summarise(count = n()) %>% 
  unite("Species.Colony.Treatment", Species,Colony,Treatment, sep = ".") %>% 
  pivot_wider(names_from="Reason_removed", values_from = "count") %>% 
  mutate(bleached = if_else(is.na(bleached),0,bleached)) %>% 
  mutate(RTL = if_else(is.na(RTL),0,RTL)) %>% 
  column_to_rownames(var="Species.Colony.Treatment") %>% 
  as.data.frame() %>% 
  pairwise_prop_test() #%>% 
  #write_csv("pairwise_prop_test_RTLvsbleaching_slowburn.csv")
```



