---
title: "FvFm"
author: "allyson_demerlis"
date: "2023-08-28"
output: html_document
---
# Packages

```{r}
library(tidyverse)
library(plotrix)
library(emmeans)
library(lme4) 
library(lmerTest)
library(ggpubr)
library(drc) #for the function "drm" dose-response model
library(broom) #for the tidy function needed for dose response model
library(car)
library(rstatix)
library(rcompanion)
library(nlme)
library(multcomp)
library(dunn.test)
```


# Import data and create tidy datasets
```{r}
ipam_tidy_data <- read.csv("ipam_tidy_data.csv")

treatment_time_bothspecies <- ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>%
  drop_na(Colony) %>% #this is because there are some corals that never matched up to IPAM values so they're just NA
  dplyr::select(Date, Colony, Puck, Tank, fvfm, Treatment, Species) %>% 
  dplyr::filter(Date <= "2022-04-20") %>% 
  pivot_wider(names_from = Date, values_from = fvfm) %>% 
   mutate(fvfm_t0 = case_when(Species == "Acervicornis" ~ `2022-03-16`,
                                  Species == "Pclivosa" ~ `2022-03-21`)) %>% 
   mutate(fvfm_t1 = case_when(Species == "Acervicornis" ~ `2022-04-06`,
                                  Species == "Pclivosa" ~ `2022-04-06`)) %>% 
   mutate(fvfm_t2 = case_when(Species == "Acervicornis" ~ `2022-04-20`,
                                  Species == "Pclivosa" ~ `2022-04-20`)) %>% 
  pivot_longer(fvfm_t0:fvfm_t2, names_to = "fvfm_timepoint", values_to = "fvfm") %>% 
  mutate(numDays = case_when(fvfm_timepoint == "fvfm_t0" ~ 0,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t1" ~ 16,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t1" ~ 21,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t2" ~ 30,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t2" ~ 35,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t3" ~ 65,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t3" ~ 70))

treatment_normalized_bothspecies <- ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>%
  drop_na(Colony) %>% #this is because there are some corals that never matched up to IPAM values so they're just NA
  dplyr::select(Date, Colony, Puck, Tank, fvfm, Treatment, Species) %>% 
  dplyr::filter(Date <= "2022-04-20") %>% #pre-CBASS time points
  pivot_wider(names_from = Date, values_from = fvfm) %>% 
  mutate(fvfm_loss = case_when(Species == "Acervicornis" ~ `2022-03-16` - `2022-04-20`,
                               Species == "Pclivosa" ~ `2022-03-21` - `2022-04-20`)) %>% 
  mutate(fvfm_loss_norm = case_when(Species == "Acervicornis" ~ fvfm_loss/`2022-03-16`,
                               Species == "Pclivosa" ~ fvfm_loss/`2022-03-21`))
```


# 1. (for figure) Raw Fv/Fm over time during the 28-d temperature treatment period.
```{r, warning = F}
treatment_time_bothspecies %>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment, color = Treatment)) + 
  geom_boxplot(color = "black", outlier.shape = NA, alpha=0.3) + #not removing outliers, just removing the black dots because they are already represented in the geom_point() line
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Species) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 14)) 
#ggsave("treatment_bothspecies_overtimeboxplots.pdf")
```

By genotype 
```{r, warning = F}
treatment_time_bothspecies%>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment, color = Treatment)) + 
  geom_boxplot(color = "black", outlier.shape = NA,  alpha=0.3) + #not removing outliers, just removing the black dots because they are already represented in the geom_point() line
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Species + Colony) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 14))
```
By genotype and tank
```{r, warning = F}
#Acer is the only one that was in tank 8
treatment_time_bothspecies%>% 
  filter(Species == "Acervicornis") %>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment, color = Treatment)) + 
  geom_boxplot(color = "black", outlier.shape = NA,  alpha=0.3) + #not removing outliers, just removing the black dots because they are already represented in the geom_point() line
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Colony + Tank) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 14))

#look at Pcli and see if there is any tank pattern
treatment_time_bothspecies%>% 
  filter(Species == "Pclivosa") %>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment, color = Treatment)) + 
  geom_boxplot(color = "black", outlier.shape = NA,  alpha=0.3) + #not removing outliers, just removing the black dots because they are already represented in the geom_point() line
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Colony + Tank) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 14))
#no pattern like in Acer
```


# 2. (for stats) Normalized to initial measurement (day 0) - Treatment Fv/Fm

## 2a. Reduction of Fv/Fm over 28-d Treatment, normalized to initial measurement (all genotypes combined)
```{r, warning = F}
treatment_normalized_bothspecies%>% 
  ggplot(., aes(x=Treatment, y= fvfm_loss_norm, color= Treatment)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  theme_classic() + 
  labs(y = "Normalized Reduction of Fv/Fm", x="Treatment") +
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 12)) + 
  facet_wrap(~Species, scales = "free_x")
#ggsave("treatment_normalized_bothspecies_boxplot.pdf")
```



### Statistics

```{r}
str(treatment_normalized_bothspecies)
#make factors: Colony, Puck, Tank, Treatment, Species

treatment_normalized_bothspecies %>% 
  mutate_at(vars(Colony, Puck, Tank, Treatment, Species), factor) %>% 
  select(!`2022-04-06`:`2022-04-20`) %>% 
  drop_na() -> treatment_normalized_bothspecies

str(treatment_normalized_bothspecies)
```


Test of normality and heteroschedascity, outlier removal
```{r}
treat_model <- lm(fvfm_loss_norm ~ Treatment*Species + Tank + Colony, data = treatment_normalized_bothspecies)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics) #not significant
```


Comparing GLMs and picking one with best fit

model generation and AICs
```{r}
full_model <- lme4::lmer(fvfm_loss_norm ~ Treatment*Species + (1|Tank) + (1|Colony), data = treatment_normalized_bothspecies)
AIC(full_model) #-912.7244
logLik(full_model) #463.3622 (df=7)
summary(full_model) # to get variance of random effects
capture.output(Anova(full_model), file = "treatment_normalized_bothspecies_lmer.csv")

no_tank_model <- lme4::lmer(fvfm_loss_norm ~ Treatment*Species + (1|Colony), data = treatment_normalized_bothspecies)
AIC(no_tank_model) #-735.5007
logLik(no_tank_model) #373.7503 (df=6)
summary(no_tank_model) # to get variance of random effects

no_colony_model <- lme4::lmer(fvfm_loss_norm ~ Treatment*Species + (1|Tank), data = treatment_normalized_bothspecies)
AIC(no_colony_model)#-894.6614
logLik(no_colony_model) #453.3307 (df=6)
summary(no_colony_model)
```


Likelihood Ratio Test function (from Kevin Wong)
```{r}
## Model comparison function

#two *nested* models as input, in this case obj1 is the model without the predictor variable and obj2 is the model with the predictor variable 
lrt <- function (obj1, obj2) {
  L0 <- logLik(obj1)
  L1 <- logLik(obj2)
  L01 <- as.vector(- 2 * (L0 - L1))
  df <- attr(L1, "df") - attr(L0, "df")
  list(L01 = L01, df = df,
       "p-value" = pchisq(L01, df, lower.tail = FALSE))
}
```


```{r}
lrt(no_tank_model, full_model) #significant
lrt(no_colony_model, full_model) #significant
```

Based on AICs and p-values, I am selecting the full model for use.

```{r}
Anova(full_model)

#recreate full model but with Treatment*Species as one variable so that the next function can interpret results
treatment_normalized_bothspecies$treat_species <- interaction(treatment_normalized_bothspecies$Treatment,treatment_normalized_bothspecies$Species)

full_model_combTreatSpecies <- lme4::lmer(fvfm_loss_norm ~ treat_species + (1|Tank) + (1|Colony), data = treatment_normalized_bothspecies)
#double check you get the same AIC as before
AIC(full_model_combTreatSpecies)
# yes = -912.7244

capture.output(summary(glht(full_model_combTreatSpecies,linfct=mcp(treat_species="Tukey")), test = adjusted(type = "bonferroni")), file = "treatment_normalized_bothspecies_glht.csv")

# Control Pclivosa vs. Control Acer significant (this result goes with figure)
```




## 2b. Reduction of Fv/Fm over 28-d Treatment, normalized to initial measurement (genotypes separated)
```{r, warning = F}
treatment_normalized_bothspecies%>% 
  ggplot(., aes(x=Colony, y= fvfm_loss_norm, color= Treatment)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  theme_classic() + labs(y = "Normalized Reduction of Fv/Fm", x="Genotype") +
  scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34"))  +
  scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 12)) + 
  facet_wrap(~Species, scales = "free_x")
#ggsave("treatment_normalized_genotypes_boxplot.pdf")
```

### Statistics

Test of normality and heteroschedascity, outlier removal
```{r}
str(treatment_normalized_bothspecies)
#make factors: Colony, Puck, Tank, Treatment, Species

treatment_normalized_bothspecies %>% 
  mutate_at(vars(Colony, Puck, Tank, Treatment, Species), factor) -> treatment_normalized_bothspecies
```

Testing inclusion of colony as a fixed effect
```{r}
treat_model_colonyfixed <- lm(fvfm_loss_norm ~ Treatment*Species*Colony + Tank, data = treatment_normalized_bothspecies)
treat_model_colonyfixed_metrics <- augment(treat_model_colonyfixed)
plot(treat_model_colonyfixed)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_colonyfixed_metrics$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_colonyfixed_metrics) #not significant

full_model <- lme4::lmer(fvfm_loss_norm ~ Treatment*Species*Colony + (1|Tank), data = treatment_normalized_bothspecies)
AIC(full_model) #-865.908
logLik(full_model) #446.954 (df=14)
summary(full_model) # to get variance of random effects
#capture.output(Anova(full_model), file = "treatment_normalized_bothspecies_lmer.csv")

Anova(full_model)

#recreate full model but with Treatment*Species as one variable so that the next function can interpret results
treatment_normalized_bothspecies$treat_species_colony <- interaction(treatment_normalized_bothspecies$Treatment,treatment_normalized_bothspecies$Species, treatment_normalized_bothspecies$Colony)

full_model_combTreatSpeciesColony <- lme4::lmer(fvfm_loss_norm ~ treat_species_colony + (1|Tank), data = treatment_normalized_bothspecies)
#double check you get the same AIC as before
AIC(full_model_combTreatSpeciesColony)
# yes = -865.908

summary(glht(full_model_combTreatSpeciesColony,linfct=mcp(treat_species_colony="Tukey")), test = adjusted(type = "bonferroni"))

# Control Pclivosa vs. Control Acer significant (this result goes with figure)
```



Just Acer genotypes
```{r}
#subset to test genotypes within each species
Acer <- treatment_normalized_bothspecies[treatment_normalized_bothspecies$Species=='Acervicornis',]

Acer_model <- lm(fvfm_loss_norm ~ Treatment*Colony + Tank, data = Acer)
plot(Acer_model)
Acer_model_metrics <- augment(Acer_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_model_metrics$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Acer_model_metrics) #significant

#visually inspect
acer_model <-lmer(fvfm_loss_norm ~ Treatment*Colony + (1|Tank), data=Acer) #there is only one possible random effect, so I can't really do model selection.
sjPlot::plot_model(acer_model, type="diag")
```


