---
title: "colorscores"
author: "allyson_demerlis"
date: "2022-11-30"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(rstatix)
library(plotrix)
library(ggpubr)
library(cowplot)
library(dunn.test)

Rscoredata <- read_xlsx("data_rscores.xlsx")

Rscoredata %>% 
  rename(`Photo ID` = `Image #`) %>% 
  rename(Date = `Date (Folder Name)`) -> Rscoredata

initialphotos <- read_xlsx("photo_metadata.xlsx")
april18 <- read_xlsx("photo_metadata.xlsx", sheet = "april 18")
june21 <- read_xlsx("photo_metadata.xlsx", sheet = "june 21")
july8 <- read_xlsx("photo_metadata.xlsx", sheet = "july 8")

coral_metadata <- readxl::read_xlsx("../metadata.xlsx")

coral_metadata %>% 
  dplyr::select(Species:CBASS_tank) -> coral_metadata
```


```{r}
str(initialphotos) 
initialphotos %>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> initialphotos

str(april18)
april18 %>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> april18

str(june21) 
june21%>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> june21

str(july8)
july8 %>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> july8

full_join(initialphotos, april18) %>% 
  full_join(., june21) %>% 
  full_join(., july8) -> allRscores

allRscores

#Tank column depends on time point, as the tank assignments changed throughout timeline. But, should be able to use the Puck ID + Species as the unique identifier

allRscores %>% 
  mutate(Species = case_when(Species == "Acer" ~ "Acropora cervicornis",
                             Species == "Pcli" ~ "Pseudodiploria clivosa")) -> allRscores

allRscores %>% 
  rename(ID = `Puck ID`) -> allRscores #821 rows

full_join(allRscores, coral_metadata, by = c("Species", "ID")) %>% 
  drop_na(`Photo ID`)  -> allmetadata #821 rows 

full_join(allmetadata, Rscoredata, by = c("Date", "Photo ID", "Coral Position")) -> Rintensity_withmetadata

Rintensity_withmetadata %>% 
  dplyr::select(!Genotype) -> Rintensity_withmetadata

Rintensity_withmetadata$month <- format(Rintensity_withmetadata$Date, "%m")

Rintensity_withmetadata$Date <- as.factor(Rintensity_withmetadata$Date)

Rintensity_withmetadata$month <- as.factor(Rintensity_withmetadata$month)

Rintensity_withmetadata %>% 
  group_by(ID, Species, Colony, Treatment, month) %>% 
  mutate(mean_Rintensity_percoral = mean(Mean)) -> Rintensity_withmetadata

# Rintensity_withmetadata %>% 
#   filter(is.na(Species)) #photo 4709 is a duplicate from photo 4711 so I already have these values in the spreadsheet

Rintensity_withmetadata %>% 
  drop_na(ID) -> Rintensity_withmetadata
``` 


```{r}
  ggplot(Rintensity_withmetadata,aes(x = month, y= mean_Rintensity_percoral, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  ylim(0, 255) +
  labs(y = "R Intensity") +
  theme(text = element_text(size = 13)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```

# Calculate percent change for the treatment period 
```{r}
Rintensity_withmetadata %>% 
  ungroup() %>% 
  filter(month == "03" | month == "04") %>% 
  dplyr::select(ID, Species, Colony, Treatment, Treatment_Tank, month, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  pivot_wider(names_from = "month", values_from = "mean_Rintensity_percoral") %>% 
  mutate(percent_change = ((`04`-`03`)/`03`)*100) -> percentchange

percentchange %>% 
  drop_na() -> percentchange #7 corals were not measured in the second time point 

#sample sizes
percentchange %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n())
```


```{r}
ggplot(percentchange,aes(x = Treatment, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(legend.position="none")
```

```{r}
ggplot(percentchange,aes(x = Colony, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```


# for figure: dotplots
```{r}
p1<-percentchange %>% 
  group_by(Species, Treatment) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Treatment, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Treatment, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Treatment, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Treatment")+
  ylab("% Change R-Intensity")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15)) +
  theme(legend.position="none")
```


```{r}
p2<-percentchange %>% 
  group_by(Species, Treatment,Colony) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Colony, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Colony, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Colony, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Colony")+
  ylab("% Change R-Intensity")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15))
```

```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5,2))

#ggsave("percentchangeRintensity_species_genets_treatmentperiod_dotplot.pdf", width = 12, height = 5)
```

## Statistics on percent change

### species x treatment

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
str(percentchange) #275 rows

percentchange %>% 
  mutate_at(vars(Colony, Treatment, Species), factor) -> percentchange

treat_model_1 <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = percentchange)
treat_model_metrics_1 <- augment(treat_model_1)
plot(treat_model_1)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics_1$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics_1) #significant

# identifying outliers
treat_model_metrics_1 %>% 
  filter(abs(.std.resid) > 3)
# 7 outliers 

percentchange %>% 
  arrange(desc(percent_change)) # everything greater than 180 is an outlier

#remove outliers
percentchange  %>% 
  filter(percent_change < 180) -> percentchange

#rerun
treat_model_1 <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = percentchange)
treat_model_metrics_1 <- augment(treat_model_1)

shapiro_test(treat_model_metrics_1$.resid) #still significant
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics_1) #still significant

#run kruskal-wallis test

#make interaction term
percentchange$Treatment_Species <- interaction(percentchange$Treatment, percentchange$Species)

kruskal_test(percent_change ~ Treatment_Species, data = percentchange) #significant

#posthoc test
as.data.frame(dunn.test(percentchange$percent_change, percentchange$Treatment_Species, method = "bonferroni")) #%>% 
  #write_csv("changeinRintensity_posthoc_species.csv")

```


### within species colony X treatment

#### Acer
```{r}
percentchange %>% 
  filter(Species == "Acropora cervicornis") -> Acer_R_percentchange

Acer_Rtreat_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Acer_R_percentchange)
Acer_Rtreat_model_metrics <- augment(Acer_Rtreat_model)
plot(Acer_Rtreat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_Rtreat_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Acer_Rtreat_model_metrics) #significant

# identifying outliers
Acer_Rtreat_model_metrics %>% 
  filter(abs(.std.resid) > 3) #0 outliers

#run non parametric test

#make interaction term
Acer_R_percentchange$Treatment_Colony <- interaction(Acer_R_percentchange$Treatment, Acer_R_percentchange$Colony)

kruskal_test(percent_change ~ Treatment_Colony, data = Acer_R_percentchange)#significant

#posthoc test
as.data.frame(dunn.test(Acer_R_percentchange$percent_change, Acer_R_percentchange$Treatment_Colony, method = "bonferroni")) 
```

#### Pcli
```{r}
percentchange %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_R_percentchange

Pcli_Rtreat_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Pcli_R_percentchange)
Pcli_Rtreat_model_metrics <- augment(Pcli_Rtreat_model)
plot(Pcli_Rtreat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Pcli_Rtreat_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Pcli_Rtreat_model_metrics) #not significant

# two-way anova
summary(aov(percent_change ~ Treatment*Colony, data = Pcli_R_percentchange)) #significant

TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Pcli_R_percentchange))
```


```{r}
percentchange %>% 
  dplyr::group_by(Species, Treatment) %>% 
  dplyr::mutate(mean_percent_change = mean(percent_change), se_percent_change = sd(percent_change)) %>% 
  ggplot(., aes(x = Species, y = mean_percent_change))+
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.2, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Species, ymax = mean_percent_change+se_percent_change, ymin = mean_percent_change-se_percent_change, color = Treatment), width = .2, position = position_dodge(width=0.8)) +
  #scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Species, fill = Treatment), size = 2, pch = 21, color = "black", position = position_dodge(width=0.8))+
  #scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Day")+
  #facet_wrap(~Species, scales = "free_x", dir = "v") +
  theme(text = element_text(size = 15))
```



# Calculate percent change for the slow-burn heat stress assay
```{r}
Rintensity_withmetadata %>% 
  ungroup() %>% 
  filter(month == "06" | month == "07") %>% 
  dplyr::select(ID, Species, Colony, Treatment, Treatment_Tank, month, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  pivot_wider(names_from = "month", values_from = "mean_Rintensity_percoral") %>% 
  mutate(percent_change = ((`07`-`06`)/`06`)*100) -> slowburn_Rintensity

slowburn_Rintensity %>% 
  drop_na() -> slowburn_Rintensity #~60 corals were not measured in the second time point 
```


```{r}
p1<- ggplot(slowburn_Rintensity,aes(x = Treatment, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(legend.position="none")
```

```{r}
p2<-ggplot(slowburn_Rintensity,aes(x = Colony, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```

```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5,2))

#ggsave("percentchangeRintensity_species_genets_slowburn.pdf", width = 12, height = 5)
```


## Statistics on percent change

### species x treatment

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
str(slowburn_Rintensity) #99 rows

slowburn_Rintensity %>% 
  mutate_at(vars(Colony, Treatment, Species), factor) -> slowburn_Rintensity

treat_model <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = slowburn_Rintensity)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics) #not significant

# anova

summary(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity)) #significant for species

TukeyHSD(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity))

#capture.output(TukeyHSD(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity)), file = "slowburn_Rintensity_treatmentspecies.csv")
```


### within species colony X treatment

#### Acer
```{r}
slowburn_Rintensity %>% 
  filter(Species == "Acropora cervicornis") -> Acer_slowburn_Rintensity

Acer_R_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Acer_slowburn_Rintensity)
Acer_R_model_model_metrics <- augment(Acer_R_model)
plot(Acer_R_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_R_model_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Acer_R_model_model_metrics) #not significant

# anova

summary(aov(percent_change ~ Treatment*Colony, data = Acer_slowburn_Rintensity)) 

TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Acer_slowburn_Rintensity))
```

#### Pcli
```{r}
slowburn_Rintensity %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_slowburn_Rintensity

Pcli_R_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Pcli_slowburn_Rintensity)
Pcli_R_model_metrics <- augment(Pcli_R_model)
plot(Pcli_R_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Pcli_R_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Pcli_R_model_metrics) #not significant

# two-way anova
summary(aov(percent_change ~ Treatment*Colony, data = Pcli_slowburn_Rintensity)) #significant

TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Pcli_slowburn_Rintensity))
```



 