---
title: "FvFm"
author: "allyson_demerlis"
date: "2023-08-28"
output: html_document
---
# Packages

```{r}
library(tidyverse)
library(plotrix)
library(ggpubr)
library(rstatix)
library(cowplot)
```


# Import data and create tidy datasets
```{r}
ipam_tidy_data <- read.csv("ipam_tidy_data.csv")

treatment_time_bothspecies <- ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>%
  drop_na(Colony) %>% #this is because there are some corals that never matched up to IPAM values so they're just NA
  dplyr::select(Date, Colony, Puck, Tank, fvfm, Treatment, Species) %>% 
  mutate(Date = mdy(Date)) %>% 
  dplyr::filter(Date <= "2022-04-20") %>% 
  pivot_wider(names_from = Date, values_from = fvfm) %>% 
   mutate(fvfm_t0 = case_when(Species == "Acervicornis" ~ `2022-03-16`,
                                  Species == "Pclivosa" ~ `2022-03-21`)) %>% 
   mutate(fvfm_t1 = case_when(Species == "Acervicornis" ~ `2022-04-06`,
                                  Species == "Pclivosa" ~ `2022-04-06`)) %>% 
   mutate(fvfm_t2 = case_when(Species == "Acervicornis" ~ `2022-04-20`,
                                  Species == "Pclivosa" ~ `2022-04-20`)) %>% 
  pivot_longer(fvfm_t0:fvfm_t2, names_to = "fvfm_timepoint", values_to = "fvfm") %>% 
  mutate(numDays = case_when(fvfm_timepoint == "fvfm_t0" ~ 0,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t1" ~ 16,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t1" ~ 21,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t2" ~ 30,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t2" ~ 35,
                             Species == "Pclivosa" & fvfm_timepoint == "fvfm_t3" ~ 65,
                             Species == "Acervicornis" & fvfm_timepoint == "fvfm_t3" ~ 70))

treatment_normalized_bothspecies <- ipam_tidy_data %>% 
  dplyr::filter(!Treatment_period == "CBASS") %>%
  drop_na(Colony) %>% #this is because there are some corals that never matched up to IPAM values so they're just NA
  dplyr::select(Date, Colony, Puck, Tank, fvfm, Treatment, Species)  %>% 
  mutate(Date = mdy(Date)) %>% 
  dplyr::filter(Date <= "2022-04-20") %>% #pre-CBASS time points
  pivot_wider(names_from = Date, values_from = fvfm) %>% 
  mutate(fvfm_loss = case_when(Species == "Acervicornis" ~ `2022-04-20` - `2022-03-16`,
                               Species == "Pclivosa" ~ `2022-04-20` - `2022-03-21`)) %>% 
  mutate(fvfm_loss_norm = case_when(Species == "Acervicornis" ~ (fvfm_loss/`2022-03-16`)*100,
                               Species == "Pclivosa" ~ (fvfm_loss/`2022-03-21`)*100))

#save data for WGCNA
treatment_normalized_bothspecies %>% 
  select(Colony, Puck, Treatment, Species, fvfm_loss, fvfm_loss_norm) %>% 
  write_csv("treatment_fvfm.csv")

```

sample sizes for corals measured during 28-d treatment
```{r}
treatment_time_bothspecies %>% 
  select(Colony:Species, fvfm_timepoint, fvfm) %>% 
  filter(fvfm_timepoint == "fvfm_t2") %>% #282 corals
  group_by(Species, Colony, Treatment) %>% 
  dplyr::summarise(count = n())
```



# 1. (for figure) Raw Fv/Fm over time during the 28-d temperature treatment period.
```{r, warning = F}
treatment_time_bothspecies %>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment, color = Treatment)) + 
  geom_boxplot(color = "black", outlier.shape = NA, alpha=0.3) + #not removing outliers, just removing the black dots because they are already represented in the geom_point() line
  geom_point(position = position_jitterdodge(), alpha=0.3) +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Species) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) 
#ggsave("treatment_bothspecies_overtimeboxplots.pdf")
```

By genotype 
```{r, warning = F}
treatment_time_bothspecies%>% 
  dplyr::select(!c(`2022-04-06`:`2022-04-20`)) %>% 
  mutate(numDays = as.factor(numDays)) %>% 
  ggplot(., aes(x=numDays, y=fvfm, fill = Treatment)) + 
  geom_boxplot(color = "black") +
  stat_summary(fun=mean, geom="line", aes(group=Treatment, color = Treatment), position = position_dodge(width = 0.5)) +
  facet_wrap(~Species + Colony) + 
  theme_classic() + 
  labs(y = "Fv/Fm", x = "Days in Treatment") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 14))
```

# 2. (for stats) Normalized to initial measurement (day 0) - Treatment Fv/Fm

## 2a. Reduction of Fv/Fm over 28-d Treatment, normalized to initial measurement (all genotypes combined)
```{r, warning = F}
treatment_normalized_bothspecies%>% 
  ggplot(., aes(x=Species, y= fvfm_loss_norm, fill= Treatment)) +
  geom_boxplot() + 
  theme_classic() + 
  labs(y = "% Decline in Fv/Fm", x="Treatment") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) + 
  theme(legend.position="none")
  
#ggsave("treatment_normalized_bothspecies_boxplot.pdf")
```

### Statistics

Change to factors
```{r}
str(treatment_normalized_bothspecies)
#make factors: Colony, Puck, Tank, Treatment, Species

treatment_normalized_bothspecies %>% 
  mutate_at(vars(Colony, Puck, Tank, Treatment, Species), factor) -> treatment_normalized_bothspecies
```

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
treat_model_1 <- lm(fvfm_loss_norm ~ Treatment*Species + Colony + Tank, data = treatment_normalized_bothspecies)
treat_model_metrics_1 <- augment(treat_model_1)
plot(treat_model_1)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics_1$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics_1) #not significant

# identifying outliers
treat_model_metrics_1 %>% 
  filter(abs(.std.resid) > 3) -> treat_outliers_1
# 2 outliers

#remove outliers
treatment_normalized_bothspecies_2wayanova <- treatment_normalized_bothspecies[!rownames(treatment_normalized_bothspecies) %in% treat_outliers_1$.rownames,]

#two corals were indeed removed from data frame
```

Run 2-way ANOVA
```{r}
aov(fvfm_loss_norm ~ Treatment*Species, data = treatment_normalized_bothspecies_2wayanova)

summary(aov(fvfm_loss_norm ~ Treatment*Species, data = treatment_normalized_bothspecies_2wayanova))

#capture.output(summary(aov(fvfm_loss_norm ~ Treatment*Species, data = treatment_normalized_bothspecies)), file = "fvfm_loss_norm_bothspecies_anova_table.csv")

TukeyHSD(aov(fvfm_loss_norm ~ Treatment*Species, data = treatment_normalized_bothspecies_2wayanova))
         
#capture.output(TukeyHSD(aov(fvfm_loss_norm ~ Treatment*Species, data = treatment_normalized_bothspecies)), file = "fvfm_loss_norm_bothspecies_anova_tukey.csv")
```


## 2b. Reduction of Fv/Fm over 28-d Treatment, normalized to initial measurement (genotypes separated)
```{r, warning = F}
treatment_normalized_bothspecies%>% 
  ggplot(., aes(x=Colony, y= fvfm_loss_norm, fill=Treatment)) +
  geom_boxplot() + 
  theme_classic() + labs(y = "% Decline in Fv/Fm", x="Colony") +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) + 
  facet_wrap(~Species, scales = "free_x")
#ggsave("treatment_normalized_genotypes_boxplot.pdf")
```

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
treat_model_2 <- lm(fvfm_loss_norm ~ Treatment*Species*Colony + Tank, data = treatment_normalized_bothspecies)
treat_model_metrics_2 <- augment(treat_model_2)
plot(treat_model_2)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics_2$.resid) # not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species*Colony, data = treat_model_metrics_2) #not significant

# identifying outliers
treat_model_metrics_2 %>% 
  filter(abs(.std.resid) > 3) -> treat_outliers_2
# 2 outliers

#remove outliers
treatment_normalized_bothspecies_3wayanova <- treatment_normalized_bothspecies[!rownames(treatment_normalized_bothspecies) %in% treat_outliers_2$.rownames,]
```


Run 3-way ANOVA
```{r}
aov(fvfm_loss_norm ~ Treatment*Species*Colony, data = treatment_normalized_bothspecies_3wayanova)

summary(aov(fvfm_loss_norm ~ Treatment*Species*Colony, data = treatment_normalized_bothspecies_3wayanova))

#capture.output(summary(aov(fvfm_loss_norm ~ Treatment*Species*Colony, data = treatment_normalized_bothspecies_3wayanova)), file = "fvfm_loss_norm_bothspeciesgenet_anova_table.csv")

TukeyHSD(aov(fvfm_loss_norm ~ Treatment*Species*Colony, data = treatment_normalized_bothspecies_3wayanova))
         
#capture.output(TukeyHSD(aov(fvfm_loss_norm ~ Treatment*Species*Colony, data = treatment_normalized_bothspecies_3wayanova)), file = "fvfm_loss_norm_bothspeciesgenet_anova_tukey.csv")
```


# for figure: mean standard error dot plots
```{r}
p1<-treatment_normalized_bothspecies %>% 
  drop_na(fvfm_loss_norm) %>% 
  group_by(Species, Treatment) %>% 
  mutate(mean_fvfmlossnorm = mean(fvfm_loss_norm), se_fvfmlossnorm = std.error(fvfm_loss_norm)) %>% 
  ggplot(., aes(x = Treatment, y = mean_fvfmlossnorm)) +
  theme_classic()+
  geom_jitter(aes(y=fvfm_loss_norm, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Treatment, ymax = mean_fvfmlossnorm+se_fvfmlossnorm, ymin = mean_fvfmlossnorm-se_fvfmlossnorm, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Treatment, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Treatment")+
  ylab("% Decline in Fv/Fm")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15)) +
  ylim(-4,40) +
  theme(legend.position="none")
```

```{r}
p2<-treatment_normalized_bothspecies %>% 
  drop_na(fvfm_loss_norm) %>% 
  group_by(Species, Treatment, Colony) %>% 
  mutate(mean_fvfmlossnorm = mean(fvfm_loss_norm), se_fvfmlossnorm = std.error(fvfm_loss_norm)) %>% 
  ggplot(., aes(x = Colony, y = mean_fvfmlossnorm)) +
  theme_classic()+
  geom_jitter(aes(y=fvfm_loss_norm, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Colony, ymax = mean_fvfmlossnorm+se_fvfmlossnorm, ymin = mean_fvfmlossnorm-se_fvfmlossnorm, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Colony, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Colony")+
  ylab("% Decline in Fv/Fm")+
  facet_wrap(~Species, scales = "free_x") +
  ylim(-4,40) +
  theme(text = element_text(size = 15))
```


create combined plot
```{r}
plot_grid(p1, p2, rel_widths = c(1.5,2))
#ggsave("treatment_percentfvfmloss_dotplot.pdf", width = 12, height = 5)
```

