---
title: "BW_stresshardening2022"
author: "Allyson DeMerlis"
date: "4/11/2022"
output: html_document
---
# load libraries and data frames
```{r}
library(tidyverse)
library(gsw) #for density calculation
library(rstatix)
library(ggsignif)
library(dunn.test)
library(cowplot)
library(plotrix)

metadata <- read_csv("../metadata.csv")

buoyantweights<- metadata %>% 
    dplyr::select(Species, ID, Colony, Treatment, Treatment_Tank, Bw_initial_raw_mass:bw_final_temp) %>% 
    drop_na(Treatment)

#sample size
buoyantweights %>% 
  drop_na(bw_final_raw_mass) %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n())

#the final time point sample sizes are much lower because I didn't buoyant weigh any corals that got tissue sampled or FACS run (N=69 corals)
```
# Species x Treatment

## Calculating percent change of mass in air for both species, all genotypes together
```{r}
#determining seawater density using temperature and salinity (from Patrick Kiel's code)
#general equation:gsw_rho_t_exact(SA = salinity, t = temp,p = 10.1325)/1000

#initial and final corrected buoyant weight calculations
buoyantweights %>% 
  group_by(Species, ID, Treatment, bw_initial_date) %>% 
  mutate(SWrho_initial = gsw_rho_t_exact(SA = `bw_initial_salinity (ppt)`,
                         t = bw_initial_temp,
                         p = 10.1325)/1000) %>% 
  mutate(CalcMAir_initial = Bw_initial_raw_mass/(1-SWrho_initial/2.93)) %>%  #2.93 is the density of solid aragonite
  ungroup() %>% 
  group_by(Species, ID, Treatment, bw_final_date) %>% 
  mutate(SWrho_final = gsw_rho_t_exact(SA = bw_final_salinity,
                         t = bw_final_temp,
                         p = 10.1325)/1000) %>% 
  mutate(CalcMAir_final = bw_final_raw_mass/(1-SWrho_final/2.93)) %>% 
  filter(!(Species == "Acropora cervicornis" & ID == "94")) %>% #drop this one sample because it was tissue sampled
  ungroup() %>% 
  mutate(percent_change = ((CalcMAir_final - CalcMAir_initial)/ CalcMAir_initial)*100) %>% 
  drop_na() -> calc_bw_bothspecies_bothtimepoints

#write_csv(calc_bw_bothspecies_bothtimepoints, file = "calc_bw_bothspecies_bothtimepoints.csv")
```

## initial vs final masses for corals
```{r}
calc_bw_bothspecies_bothtimepoints %>% 
  dplyr::select(Species:Treatment_Tank, CalcMAir_initial, CalcMAir_final) %>% 
  pivot_longer(CalcMAir_initial:CalcMAir_final, names_to = "timepoint", values_to = "mass") %>% 
  mutate(timepoint = case_when(timepoint == "CalcMAir_initial" ~ "0",
                               timepoint == "CalcMAir_final" ~ "28")) %>% 
  ggplot(., aes(x=timepoint, y = mass, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() + 
   scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  facet_wrap(~Species+Colony+Treatment)
```


## boxplot of % growth - all genotypes together
```{r}
calc_bw_bothspecies_bothtimepoints %>% 
  group_by(Treatment) %>% 
  ggplot(., aes(x=Species, y = percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() + 
   scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(y= "% Growth") +
  theme(legend.position="none")
#ggsave("percentgrowth_speciestreatments.pdf")
```
## stats 

Change to factors
```{r}
str(calc_bw_bothspecies_bothtimepoints)
#make factors: Colony, Puck, Tank, Treatment, Species

calc_bw_bothspecies_bothtimepoints %>% 
  mutate_at(vars(Colony, ID, Treatment_Tank, Treatment, Species), factor) -> calc_bw_bothspecies_bothtimepoints
```

### species tested together

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
treat_model <- lm(percent_change ~ Treatment*Species + Colony, data = calc_bw_bothspecies_bothtimepoints)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics) #not significant
```

use non-parametric method
```{r}
#do nonparametric test with interaction term of colony and tank
calc_bw_bothspecies_bothtimepoints$species_treatment <- interaction(calc_bw_bothspecies_bothtimepoints$Species, calc_bw_bothspecies_bothtimepoints$Treatment)

kruskal_test(calc_bw_bothspecies_bothtimepoints, percent_change ~ species_treatment) #significant

dunn_results <- dunn.test(calc_bw_bothspecies_bothtimepoints$percent_change, calc_bw_bothspecies_bothtimepoints$species_treatment, method = "bonferroni")

bw_posthoc_results <- as.data.frame(dunn_results) 

#write_csv(bw_posthoc_results, "bw_posthoc_results.csv")
```

### species separate

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
# Acropora cervicornis
calc_bw_bothspecies_bothtimepoints %>% 
  filter(Species == "Acropora cervicornis") -> Acer_bw

Acer_treat_model <- lm(percent_change ~ Treatment + Colony, data = Acer_bw)
treat_model_metrics <- augment(Acer_treat_model)
plot(Acer_treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment, data = treat_model_metrics) #not significant


# Pseudodiploria clivosa
calc_bw_bothspecies_bothtimepoints %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_bw

Pcli_treat_model <- lm(percent_change ~ Treatment + Colony, data = Pcli_bw)
treat_model_metrics <- augment(Pcli_treat_model)
plot(Pcli_treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment, data = treat_model_metrics) #not significant
```

use non-parametric method
```{r}
kruskal_test(Acer_bw, percent_change ~ Treatment) #not significant

kruskal_test(Pcli_bw, percent_change ~ Treatment) #significant (0.0497)
```

# Colony x Treatment

## Calculating percent change of mass in air for both species and all genotypes separated
```{r}
#determining seawater density using temperature and salinity (from Patrick Kiel's code)
#general equation:gsw_rho_t_exact(SA = salinity, t = temp,p = 10.1325)/1000

#initial and final corrected buoyant weight calculations
buoyantweights %>% 
  group_by(Species, ID, Colony, Treatment, bw_initial_date) %>% 
  mutate(SWrho_initial = gsw_rho_t_exact(SA = `bw_initial_salinity (ppt)`,
                         t = bw_initial_temp,
                         p = 10.1325)/1000) %>% 
  mutate(CalcMAir_initial = Bw_initial_raw_mass/(1-SWrho_initial/2.93)) %>%  #2.93 is the density of solid aragonite
  ungroup() %>% 
  group_by(Species, ID, Colony, Treatment, bw_final_date) %>% 
  mutate(SWrho_final = gsw_rho_t_exact(SA = bw_final_salinity,
                         t = bw_final_temp,
                         p = 10.1325)/1000) %>% 
  mutate(CalcMAir_final = bw_final_raw_mass/(1-SWrho_final/2.93)) %>% 
  filter(!(Species == "Acropora cervicornis" & ID == "94")) %>% #drop this one sample because it was tissue sampled
  ungroup() %>% 
  mutate(percent_change = ((CalcMAir_final - CalcMAir_initial)/ CalcMAir_initial)*100) %>% 
  drop_na() -> calc_bw_bothspeciesgenet_bothtimepoints

#write_csv(calc_bw_bothspeciesgenet_bothtimepoints, file = "calc_bw_bothspeciesgenet_bothtimepoints.csv")
```

## boxplot of % growth - genotypes separated
```{r}
calc_bw_bothspeciesgenet_bothtimepoints %>% 
  group_by(Colony, Treatment) %>% 
  ggplot(., aes(x=Colony, y = percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() + 
  facet_wrap(~Species, scales= "free_x") +
    scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(y= "% Growth")
```


## stats 

Change to factors
```{r}
str(calc_bw_bothspeciesgenet_bothtimepoints)
#make factors: Colony, Puck, Tank, Treatment, Species

calc_bw_bothspeciesgenet_bothtimepoints %>% 
  mutate_at(vars(Colony, ID, Treatment_Tank, Treatment, Species), factor) -> calc_bw_bothspeciesgenet_bothtimepoints
```

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
# Acer colonies 

calc_bw_bothspeciesgenet_bothtimepoints %>% 
  filter(Species == "Acropora cervicornis") -> Acer_colony_bw

Acer_colony_treat_model <- lm(percent_change ~ Treatment*Colony, data = Acer_colony_bw)
treat_model_metrics <- augment(Acer_colony_treat_model)
plot(Acer_colony_treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = treat_model_metrics) #not significant

# Pcli colonies 

calc_bw_bothspeciesgenet_bothtimepoints %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_colony_bw

Pcli_colony_treat_model <- lm(percent_change ~ Treatment*Colony, data = Pcli_colony_bw)
treat_model_metrics <- augment(Pcli_colony_treat_model)
plot(Pcli_colony_treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) # significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = treat_model_metrics) #not significant
```
use non-parametric method
```{r}
# Acer colonies

#do nonparametric test with interaction term of colony and tank
Acer_colony_bw$treatment_colony <- interaction(Acer_colony_bw$Treatment, Acer_colony_bw$Colony)

kruskal_test(Acer_colony_bw, percent_change ~ treatment_colony) #significant

dunn_results_Acer_colony <- dunn.test(Acer_colony_bw$percent_change, Acer_colony_bw$treatment_colony, method = "bonferroni")

bw_posthoc_results_Acer_colony <- as.data.frame(dunn_results_Acer_colony) 

#write_csv(bw_posthoc_results_Acer_colony, "Acer_colony_bw_posthoc_results_bygenet.csv")

# Pcli colonies 
Pcli_colony_bw$treatment_colony <- interaction(Pcli_colony_bw$Treatment, Pcli_colony_bw$Colony)

kruskal_test(Pcli_colony_bw, percent_change ~ treatment_colony) #significant

dunn_results_Pcli_colony <- dunn.test(Pcli_colony_bw$percent_change, Pcli_colony_bw$treatment_colony, method = "bonferroni")

bw_posthoc_results_Pcli_colony <- as.data.frame(dunn_results_Pcli_colony) 

#write_csv(bw_posthoc_results_Pcli_colony, "Pcli_colony_bw_posthoc_results_bygenet.csv")
```


combined plot
```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5, 2))
ggsave("growthrates_species_genet.pdf", width = 10, height = 5)
```

# Archived code

## mean/standard error growth rate bar chart
```{r}
calc_bw_bothspecies_bothtimepoints %>% 
  group_by(Species, Treatment) %>% 
  summarise(mean = mean(percent_change), se = std.error(percent_change)) %>% 
  ggplot(., aes(x=Species, y = mean, fill = Treatment)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width =
                  .2, position=position_dodge(.9)) + 
  theme_classic() + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(y= "% Growth") +
  theme(legend.position="none")
```


## dotplot: mean/standard error 
```{r}
p1<- calc_bw_bothspecies_bothtimepoints %>% 
  group_by(Species, Treatment) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Treatment, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Treatment, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Treatment, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Treatment")+
  ylab("% Growth")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15)) +
  theme(legend.position="none")
```

```{r}
p2<- calc_bw_bothspecies_bothtimepoints %>% 
  group_by(Species, Treatment, Colony) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Colony, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Colony, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Colony, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Colony")+
  ylab("% Growth")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15))
```


combined plot
```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5, 2))
ggsave("growthrates_species_genet_dotplot.pdf", width = 12, height = 5)
```

 mean/standard error growth rate bar chart

```{r}
calc_bw_bothspecies_bothtimepoints %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(mean = mean(percent_change), se = std.error(percent_change)) %>% 
  ggplot(., aes(x=Colony, y = mean, fill = Treatment)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width =
                  .2, position=position_dodge(.9)) + 
  theme_classic() + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))  +
  theme(text = element_text(size = 15)) +
  labs(y= "% Growth") +
  facet_wrap(~Species, scales = "free_x")
```
