---
title: "tank8"
author: "allyson_demerlis"
date: "2024-01-14"
output: html_document
---

# Load libraries and data
```{r}
library(tidyverse)
library(plotrix) #for standard error function
library(cowplot)

feb4<- readxl::read_xlsx("Data from LabVIEW/20220204_TemperatureData edited.xlsx", col_types = c("date", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric"))
feb23 <- readxl::read_xlsx("Data from LabVIEW/20220223_TemperatureData edited.xlsx")
march7 <- readxl::read_xlsx("Data from LabVIEW/20220307_TemperatureData - Copy.xlsx")
march29 <- readxl::read_xlsx("Data from LabVIEW/20220329_TemperatureData.xlsx")
april15 <- readxl::read_xlsx("Data from LabVIEW/20220415_TemperatureData.xlsx")

april25 <- readxl::read_xlsx("Data from LabVIEW/20220425_TemperatureData (1).xlsx") #had to convert to csv because dates weren't reading in correctly

CBASS <- readxl::read_xlsx("Data from LabVIEW/CBASS_labview.xlsx")

may11 <- readxl::read_xlsx("Data from LabVIEW/20220511_TemperatureData.xlsx")
may20 <- readxl::read_xlsx("Data from LabVIEW/20220520_TemperatureData.xlsx")
june10 <- readxl::read_xlsx("Data from LabVIEW/20220610_TemperatureData - Copy.xlsx")
june24 <- readxl::read_xlsx("Data from LabVIEW/20220624_TemperatureData (1).xlsx")
july6 <- readxl::read_xlsx("Data from LabVIEW/20220706_TemperatureData.xlsx", 
    col_types = c("date", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric"))
july8 <- readxl::read_xlsx("Data from LabVIEW/20220708_TemperatureData edited.xlsx")
july20 <- read_csv("Data from LabVIEW/20220720_TemperatureData.csv") #had to convert to csv because dates weren't reading in correctly
  july20$Time <- lubridate::mdy_hms(july20$Time)
```

Creating subsets for different stages of experiment
```{r}
acclimation <- bind_rows(feb4, feb23) 
temptreatment<- bind_rows(march7, march29, april15) #treatment started march 24 and ended april 20
temptreatment %>% filter(Time >= as.Date("2022-03-24") & Time <= as.Date("2022-04-22")) -> temptreatment
#filter out one anomalous data point
  temptreatment %>% 
    filter(T1<23) 
# Set the specific data point to NA
temptreatment[temptreatment$T1 < 22.5, "T1"]= NA
  temptreatment %>% 
    filter(T1<23) #its gone now

CBASS #april 22
recovery <- bind_rows(april15, april25, may11, may20, june10) #recovery started april 23 and ended june 21
  recovery %>% filter(Time >= as.Date("2022-04-23") & Time <= as.Date("2022-06-22")) -> recovery
heatstressassay <- bind_rows(june10, june24, july6, july8, july20) #ramp up started june 21
```

# Treatment period (Mar 24 - Apr 20)
```{r}
#all tanks separated
temptreatment %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  ggplot(data = .,
       aes(x = Time, y = Temp, color = Tank)) +
  geom_point() + 
  theme_classic() +
  facet_wrap(~Tank) +
  labs(title = "Tank Temperatures",
       y = "Temperature (°C)",
       x = "Date")
#ggsave("temptreatment_28d.pdf", width = 10, height = 6)
```

There is something wrong with tank 8. It never reached 31ºC. 

You can see it in the standard deviation plot below. 

```{r}
#summary figure of average tank temp during 24-hr period over the 28 days
temptreatment$Time <- lubridate::round_date(temptreatment$Time, unit = "15 min")
temptreatment %>% drop_na() -> temptreatment

temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable", 
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control",
                               Tank == "T8" ~ "variable")) %>% 
  dplyr::group_by(Time, treatment) %>% 
  dplyr::summarise(avg_temp = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = treatment)) + 
  theme_classic() +
  geom_ribbon(aes(fill=treatment, group = treatment, ymin = avg_temp - sd, ymax = avg_temp + sd), alpha = 0.2) +
   geom_line(aes(group = treatment)) +
  scale_x_discrete(breaks = c("03:00:00", "09:00:00", "15:00:00", "21:00:00")) +
  labs(y="Temperature (ºC)") +
  scale_color_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34"))+
  scale_fill_manual(labels=c("Untreated", "Treated"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 15)) 
#ggsave("summarytemptreatment_stdeviation.pdf", width = 10, height = 6)
```

Separate out tanks
```{r}
temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  # mutate(treatment = case_when(Tank == "T1" ~ "control",
  #                              Tank == "T2" ~ "variable",
  #                              Tank == "T3" ~ "variable",
  #                              Tank == "T4" ~ "control",
  #                              Tank == "T5" ~ "variable", 
  #                              Tank == "T6" ~ "control",
  #                              Tank == "T7" ~ "control",
  #                              Tank == "T8" ~ "variable")) %>% 
  dplyr::group_by(Time, Tank) %>% 
  dplyr::summarise(avg_temp = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = Tank)) + 
  theme_classic() +
  geom_ribbon(aes(fill=Tank, group = Tank, ymin = avg_temp - sd, ymax = avg_temp + sd), alpha = 0.2) +
   geom_line(aes(group = Tank)) +
  scale_x_discrete(breaks = c("03:00:00", "09:00:00", "15:00:00", "21:00:00")) +
  labs(y="Temperature (ºC)") +
 # scale_color_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34"))+
 # scale_fill_manual(labels=c("Control", "Variable"), values = c( "#60DBDB", "#F54A34")) +
  theme(text = element_text(size = 15)) 
#ggsave("summarytemptreatment_alltanks_sd.pdf", width = 10, height = 6)
```


Standard deviation with tank 8 separated
```{r}
temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable",
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control")) %>%
  dplyr::group_by(Time, treatment) %>% 
  dplyr::summarise(avg_temp = mean(Temp), sd = sd(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = treatment)) + 
  theme_classic() +
  geom_ribbon(aes(fill=treatment, group = treatment, ymin = avg_temp - sd, ymax = avg_temp + sd), alpha = 0.2) +
   geom_line(aes(group = treatment)) +
  scale_x_discrete(breaks = c("03:00:00", "09:00:00", "15:00:00", "21:00:00")) +
  labs(y="Temperature (C)") +
  scale_color_manual(labels=c("Untreated", "Treated", "Tank 8"), values = c( "#60DBDB", "#F54A34", "grey"))+
  scale_fill_manual(labels=c("Untreated", "Treated", "Tank 8"), values = c( "#60DBDB", "#F54A34", "grey")) +
  theme(text = element_text(size = 15)) 
#ggsave("summarytemptreatment_tank8separate_sd.pdf", width = 10, height = 6)
```

Standard error with tank 8 separated
```{r}
#with standard error instead of standard deviation

temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable", 
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control")) %>% 
  dplyr::group_by(Time, treatment) %>% 
  dplyr::summarise(avg_temp = mean(Temp), se = std.error(Temp)) %>% 
  ggplot(., aes(x=Time, y=avg_temp, color = treatment)) + 
  theme_classic() +
  geom_ribbon(aes(fill=treatment, group = treatment, ymin = avg_temp - se, ymax = avg_temp + se), alpha = 0.2) +
   geom_line(aes(group = treatment)) +
  scale_x_discrete(breaks = c("03:00:00", "09:00:00", "15:00:00", "21:00:00")) +
  labs(y="Temperature (C)") +
  scale_color_manual(labels=c("Untreated", "Treated", "Tank 8"), values = c( "#60DBDB", "#F54A34", "grey"))+
  scale_fill_manual(labels=c("Untreated", "Treated", "Tank 8"), values = c( "#60DBDB", "#F54A34", "grey")) +
  theme(text = element_text(size = 20)) 
#ggsave("summarytemptreatment_tank8separate_stderr.pdf", width = 10, height = 6)
```

Calculate the mean daily min, max, and variance for each tank/treatment over the 28-d period.
```{r}
temptreatment %>% 
  filter(Time >= "2022-03-24 09:00:00") %>% 
  filter(Time <= "2022-04-20 09:00:00") %>% 
    #remove tissue sampling time points because during those times, you modified the tank temps to all hold at  28 C
  filter(!Time >= "2022-03-30 09:00:00" | !Time <= "2022-03-31 09:00:00") %>% 
  dplyr::select(Time:T8) %>% 
  pivot_longer(T1:T8, names_to="Tank", values_to ="Temp") %>% 
  tidyr::separate(Time, c("Date", "Time"), sep = "\\ ") %>% 
  mutate(Treatment = case_when(Tank == "T1" ~ "control",
                               Tank == "T2" ~ "variable",
                               Tank == "T3" ~ "variable",
                               Tank == "T4" ~ "control",
                               Tank == "T5" ~ "variable", 
                               Tank == "T6" ~ "control",
                               Tank == "T7" ~ "control",
                               Tank == "T8" ~ "tank_8")) %>% 
  mutate(Species = case_when(Tank == "T1" ~ "Pclivosa",
                             Tank == "T2" ~ "Pclivosa",
                               Tank == "T3" ~ "Pclivosa",
                               Tank == "T4" ~ "Pclivosa",
                               Tank == "T5" ~ "Acervicornis", 
                               Tank == "T6" ~ "Acervicornis",
                               Tank == "T7" ~ "Acervicornis",
                               Tank == "T8" ~ "Acervicornis")) %>% 
  group_by(Tank, Date, Treatment, Species) %>% 
  summarise(daily_max = max(Temp), daily_min = min(Temp), daily_range = (max(Temp)-min(Temp))) %>% 
  ungroup() %>% 
  group_by(Tank, Treatment, Species) %>% 
  summarise(mean_daily_max=mean(daily_max),sd_daily_max=sd(daily_max), mean_daily_min=mean(daily_min),sd_daily_min=sd(daily_min),mean_daily_range=mean(daily_max - daily_min), sd_daily_range=sd(daily_max - daily_min)) -> meanmaxminrange_summarytable

meanmaxminrange_summarytable %>% 
  group_by(Treatment) %>% summarise(mean_max=mean(mean_daily_max),sd_max=sd(mean_daily_max), mean_min=mean(mean_daily_min),sd_min=sd(mean_daily_min),mean_range=mean(mean_daily_max - mean_daily_min), sd_daily_range=sd(mean_daily_max - mean_daily_min))

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_max, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_max - sd_daily_max, ymax = mean_daily_max + sd_daily_max), width = 0.2) +
  theme_classic()

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_min, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_min - sd_daily_min, ymax = mean_daily_min + sd_daily_min), width = 0.2) +
  theme_classic()

ggplot(meanmaxminrange_summarytable, aes(x=Treatment, y=mean_daily_range, color=Tank)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_daily_range - sd_daily_range, ymax = mean_daily_range + sd_daily_range), width = 0.2) +
  theme_classic()
```