---
title: "colorscores"
author: "allyson_demerlis"
date: "2022-11-30"
output: html_document
---

# Read in data and R packages
```{r}
library(tidyverse)
library(readxl)
library(rstatix)
library(plotrix)
library(ggpubr)
library(cowplot)
library(FSA) #for Summarize
library(lme4)
library(car)
library(MASS)
library(emmeans)

Rscoredata <- read_xlsx("data_rscores.xlsx") 

Rscoredata %>% 
  rename(`Photo ID` = `Image #`) %>% 
  rename(Date = `Date (Folder Name)`) -> Rscoredata

# Only use the R-score data from the treatment period dates
initialphotos <- read_xlsx("photo_metadata.xlsx")
april18 <- read_xlsx("photo_metadata.xlsx", sheet = "april 18")

# june21 <- read_xlsx("photo_metadata.xlsx", sheet = "june 21")
# july8 <- read_xlsx("photo_metadata.xlsx", sheet = "july 8")

coral_metadata <- read_csv("../metadata.csv")

coral_metadata %>% 
  dplyr::select(Species:CBASS_tank) -> coral_metadata
```

# Tidy data
```{r}
str(initialphotos) 
initialphotos %>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> initialphotos

str(april18)
april18 %>% 
  dplyr::select(`Photo ID`:`Coral Position`) -> april18

# str(june21) 
# june21%>% 
#   dplyr::select(`Photo ID`:`Coral Position`) -> june21
# 
# str(july8)
# july8 %>% 
#   dplyr::select(`Photo ID`:`Coral Position`) -> july8

full_join(initialphotos, april18) -> allRscores

allRscores

#Tank column depends on time point, as the tank assignments changed throughout timeline. But, should be able to use the Puck ID + Species as the unique identifier

allRscores %>% 
  mutate(Species = case_when(Species == "Acer" ~ "Acropora cervicornis",
                             Species == "Pcli" ~ "Pseudodiploria clivosa")) -> allRscores

allRscores %>% 
  rename(ID = `Puck ID`) -> allRscores #821 rows

full_join(allRscores, coral_metadata, by = c("Species", "ID")) %>% 
  drop_na(`Photo ID`)  -> allmetadata #821 rows 

full_join(allmetadata, Rscoredata, by = c("Date", "Photo ID", "Coral Position")) -> Rintensity_withmetadata

Rintensity_withmetadata %>% 
  dplyr::select(!Genotype) -> Rintensity_withmetadata

Rintensity_withmetadata$month <- format(Rintensity_withmetadata$Date, "%m")

Rintensity_withmetadata$Date <- as.factor(Rintensity_withmetadata$Date)

Rintensity_withmetadata$month <- as.factor(Rintensity_withmetadata$month)

Rintensity_withmetadata %>% 
  group_by(ID, Species, Colony, Treatment, month) %>% 
  mutate(mean_Rintensity_percoral = mean(Mean)) -> Rintensity_withmetadata

# Rintensity_withmetadata %>% 
#   filter(is.na(Species)) #photo 4709 is a duplicate from photo 4711 so I already have these values in the spreadsheet

Rintensity_withmetadata %>% 
  drop_na(ID) -> Rintensity_withmetadata

#write_csv(Rintensity_withmetadata, "Rintensity_withmetadata.csv")

 # Calculate days relative to start of treatment
Rintensity_withmetadata %>% 
  ungroup() %>% 
  dplyr::select(Date:Species, Colony:Treatment_Tank, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  mutate(Date = as.Date(as.character(Date), format = "%Y-%m-%d")) %>% 
  mutate(Days_Relative = as.numeric(Date - as.Date("2022-03-24"))) -> Rintensity_tidydata
``` 

# Boxplot of raw values
```{r}
  ggplot(Rintensity_withmetadata,aes(x = month, y= mean_Rintensity_percoral, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  ylim(0, 255) +
  labs(y = "R Intensity") +
  theme(text = element_text(size = 13)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
#ggsave("Rintensityovertime_byspecies.pdf")

  ggplot(Rintensity_withmetadata,aes(x = month, y= mean_Rintensity_percoral, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species+Colony, scales = "free") +
  ylim(0, 255) +
  labs(y = "R Intensity") +
  theme(text = element_text(size = 13)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```
# boxplot for percent change
```{r}
Rintensity_withmetadata %>% 
  ungroup() %>% 
  filter(month == "03" | month == "04") %>% 
  dplyr::select(ID, Species, Colony, Treatment, Treatment_Tank, month, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  pivot_wider(names_from = "month", values_from = "mean_Rintensity_percoral") %>% 
  mutate(percent_change = ((`04`-`03`)/`03`)*100) -> percentchange

percentchange %>% 
  drop_na() -> percentchange

plot1<- ggplot(percentchange,aes(x = Species, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  facet_wrap(~Species, scales = "free_x", ncol = 1) +
  theme(legend.position="none") 
```

## mixed effects linear model
```{r}
percentchange %>% 
  filter(Species == "Acropora cervicornis") -> Acer_colorscore

Acer_colorscore_model <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore)
summary(Acer_colorscore_model)
anova(Acer_colorscore_model)

percentchange %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_colorscore

Pcli_colorscore_model <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Pcli_colorscore)
summary(Pcli_colorscore_model)
anova(Pcli_colorscore_model)
```


### Model fitting and assumptions diagnostic

Acer
```{r}
x = residuals(Acer_colorscore_model)
plot(fitted(Acer_colorscore_model), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test  -- significant aka not normal
car::outlierTest(Acer_colorscore_model) #row 30
Acer_colorscore[-30, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
car::outlierTest(Acer_colorscore_model_2) #row 61
Acer_colorscore_noout[-61, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
car::outlierTest(Acer_colorscore_model_2) #row 62
Acer_colorscore_noout[-62, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
car::outlierTest(Acer_colorscore_model_2) #row 36
Acer_colorscore_noout[-36, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
car::outlierTest(Acer_colorscore_model_2) #row 48
Acer_colorscore_noout[-48, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
car::outlierTest(Acer_colorscore_model_2) #row 36
Acer_colorscore_noout[-36, ] -> Acer_colorscore_noout

Acer_colorscore_model_2 <- lmer(percent_change ~ Treatment*Colony + (1|Treatment_Tank), data = Acer_colorscore_noout)
x = residuals(Acer_colorscore_model_2)
plot(fitted(Acer_colorscore_model_2), x) 
leveneTest(x ~ Treatment*Colony, data=Acer_colorscore_noout, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test
#that's the best we're gonna get

Anova(Acer_colorscore_model_2,  type = "II", test.statistic = "F") #you need to manually specify Satterthwaite's method, because otherwise it will give you the Wald/Chi-squared approximation and that is only appropriate for binomial or poisson distributions. Satterthwaite's method is for normal/gaussian

#capture.output(Anova(Acer_colorscore_model_2,  type = "II", test.statistic = "F"), file = "Acer_colorscore_percentchange_ANOVA.csv")
```

Pcli
```{r}
x = residuals(Pcli_colorscore_model)
plot(fitted(Pcli_colorscore_model), x) 
leveneTest(x ~ Treatment*Colony, data=Pcli_colorscore, center=mean) # formal statistical test for homogenity of variance #not significant
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
shapiro.test(x) # formal statistical test 

Anova(Pcli_colorscore_model,  type = "II", test.statistic = "F") 

capture.output(Anova(Pcli_colorscore_model,  type = "II", test.statistic = "F"), file = "Pcli_colorscore_percentchange_ANOVA.csv")
```


### Statistical pairwise comparisons

Acer
```{r}
#colony + treatment
colorscore_Acer_emms <- emmeans(Acer_colorscore_model_2, pairwise ~ Colony*Treatment)

pairs(colorscore_Acer_emms) 
capture.output(pairs(colorscore_Acer_emms), file="Acer_percentchange_treatment_colony_emmeans.csv")
```

Pcli
```{r}
#colony + treatment
colorscore_Pcli_emms <- emmeans(Pcli_colorscore_model, pairwise ~ Colony*Treatment)

pairs(colorscore_Pcli_emms) 
capture.output(pairs(colorscore_Pcli_emms), file="Pcli_percentchange_treatment_colony_emmeans.csv")
```

# Line + scatterplot over time 

## Species x Treatment
```{r}
plot2<-ggplot(Rintensity_tidydata, aes(x=Days_Relative, y=mean_Rintensity_percoral, color = Treatment)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species,ncol = 1, nrow = 2) +
  ylab("R Intensity")+
  theme_classic() +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) +
  labs(x="Days in Treatment") +
  theme(legend.position="none") +
  scale_x_continuous(limits = c(-16, 25), breaks = seq(-15, 25, by = 5))
```

## Colony x Treatment
```{r}
plot3<-ggplot(Rintensity_tidydata, aes(x=Days_Relative, y=mean_Rintensity_percoral, color = Treatment)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species+Colony) +
  ylab("R Intensity")+
  theme_classic() +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(text = element_text(size = 15)) +
  labs(x="Days in Treatment") +
  scale_x_continuous(limits = c(-16, 25), breaks = seq(-15, 25, by = 5)) +
  theme(legend.position="none")
```


```{r}
plot_grid(plot1, plot2,plot3, nrow=1, rel_widths = c(1, 1, 2))
ggsave("colorscore_boxplot_species_genet_lines.pdf", width = 12, height = 5)
```


## Mixed effects GLMM (fixed: treatment colony, time point; random: tank, coral ID)

Test species separately because it is not part of your research question whether Acer vs. Pcli has different color scores.

For the time component, we really just want to compare pre-treatment vs. post-treatment to reflect the effect of time and treatment. So, all the dates in March will be encoded as "T0", and all the end measurements will be "T1". It needs to be a factor to be included as a fixed effect for the post-hoc comparisons.
```{r}
Rintensity_tidydata %>% 
  filter(Species == "Acropora cervicornis") -> Acer_tidydata

Acer_tidydata %>% 
  mutate(time_point = case_when(Date < as.Date("2022-03-30") ~ "T0",
                                Date > as.Date("2022-04-17") ~ "T1")) -> Acer_tidydata

Rintensity_tidydata %>% 
  filter(Species == "Pseudodiploria clivosa") %>% 
  mutate(time_point = case_when(Date < as.Date("2022-03-30") ~ "T0",
                                Date > as.Date("2022-04-17") ~ "T1")) -> Pcli_tidydata
```


```{r}
#repeated measures -- include ID in the formula as a random effect
Acer_model <- lmer(mean_Rintensity_percoral ~ Treatment*Colony*time_point + (1|Treatment_Tank) + (1|ID), data = Acer_tidydata) 
summary(Acer_model)
Anova(Acer_model, type = "II", test.statistic = "F")

Pcli_model <- lmer(mean_Rintensity_percoral ~ Treatment*Colony*time_point + (1|Treatment_Tank) + (1|ID), data = Pcli_tidydata)
summary(Pcli_model)
Anova(Pcli_model, type = "II", test.statistic = "F")
```

### Model fitting and assumptions diagnostic

Acer
```{r}
x = residuals(Acer_model)
plot(fitted(Acer_model), x) 
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
# Acer data meets assumptions, yay

Anova(Acer_model, type = "II", test.statistic = "F")
#capture.output(Anova(Acer_model, type = "II", test.statistic = "F"), file = "Acer_colorscore_model_ANOVA.csv")
```

Pcli
```{r}
x = residuals(Pcli_model)
plot(fitted(Pcli_model), x) 
qqnorm(x); qqline(x) # qq plot to check for normal distribution of residuals
hist(x) # histogram of residuals to check for normal distribution of residuals
# Pcli data meets assumptions, yay

Anova(Pcli_model, type = "III", test.statistic = "F")
#capture.output(Anova(Pcli_model, type = "II", test.statistic = "F"), file = "Pcli_colorscore_model_ANOVA.csv")
```

### Statistical pairwise comparisons

Acer
```{r}
#colony + treatment + timepoint
Acer_emms <- emmeans(Acer_model, pairwise ~ Treatment*Colony*time_point)
capture.output(Acer_emms$contrasts, file = "Acer_Tukey_colonytreatmenttimepoint.csv")
```

Pcli
```{r}
#colony + treatment + timepoint
Pcli_emms <- emmeans(Pcli_model, pairwise ~ Treatment*Colony*time_point)
capture.output(Pcli_emms$contrasts, file = "Pcli_colonytreatmenttimepoint.csv")
```


# Archive
## Calculate percent change for the treatment period 

```{r}
Rintensity_withmetadata %>% 
  ungroup() %>% 
  filter(month == "03" | month == "04") %>% 
  dplyr::select(ID, Species, Colony, Treatment, Treatment_Tank, month, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  pivot_wider(names_from = "month", values_from = "mean_Rintensity_percoral") %>% 
  mutate(percent_change = ((`03`-`04`)/`03`)*100) -> percentchange

percentchange %>% 
  drop_na() -> percentchange #7 corals were not measured in the second time point 

# percentchange %>% 
#   select(ID:Treatment, percent_change) %>% 
#   write_csv("treatment_Rscore.csv")

#sample sizes
percentchange %>% 
  group_by(Species, Colony, Treatment) %>% 
  summarise(count = n())
```


```{r}
p1<-ggplot(percentchange,aes(x = Species, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  labs(y = "Percent Decline in Color Score") +
  theme(text = element_text(size = 15)) +
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) 
#ggsave("percentdeclinecolorscore_species.pdf")
```

```{r}
p2<-ggplot(percentchange,aes(x = Colony, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Decline in Color Score") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```

```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5,2))

#ggsave("percentchangeRintensity_species_genets_treatment.pdf", width = 12, height = 5)
```


dotplots
```{r}
percentchange %>% 
  group_by(Species, Treatment) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Treatment, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Treatment, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Treatment, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Treatment")+
  ylab("% Change R-Intensity")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15)) +
  theme(legend.position="none")
```


```{r}
percentchange %>% 
  group_by(Species, Treatment,Colony) %>% 
  mutate(mean_percentchange = mean(percent_change), se_percentchange = std.error(percent_change)) %>% 
  ggplot(., aes(x = Colony, y = mean_percentchange)) +
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.1, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Colony, ymax = mean_percentchange+se_percentchange, ymin = mean_percentchange-se_percentchange, color = Treatment), width = .1, position = position_dodge(width=0.8)) +
  scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Colony, fill = Treatment), size = 2.5, pch = 21, color = "black", position = position_dodge(width=0.8))+
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Colony")+
  ylab("% Change R-Intensity")+
  facet_wrap(~Species, scales = "free_x") +
  theme(text = element_text(size = 15))
```


## Statistics on percent change

### species x treatment

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
str(percentchange) #275 rows

percentchange %>% 
  mutate_at(vars(Colony, Treatment, Species), factor) -> percentchange

treat_model_1 <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = percentchange)
treat_model_metrics_1 <- augment(treat_model_1)
plot(treat_model_1)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics_1$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics_1) #significant

# identifying outliers
treat_model_metrics_1 %>% 
  filter(abs(.std.resid) > 3)
# 7 outliers 

percentchange %>% 
  arrange(desc(percent_change)) # everything greater than 180 is an outlier

#remove outliers
percentchange  %>% 
  filter(percent_change > -180) -> percentchange_edited

#rerun
treat_model_1 <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = percentchange_edited)
treat_model_metrics_1 <- augment(treat_model_1)

shapiro_test(treat_model_metrics_1$.resid) #still significant
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics_1) #still significant

#run kruskal-wallis test

#make interaction term
percentchange_edited$Treatment_Species <- interaction(percentchange_edited$Treatment, percentchange_edited$Species)

kruskal_test(percent_change ~ Treatment_Species, data = percentchange_edited) #significant

#posthoc test
as.data.frame(dunn.test(percentchange_edited$percent_change, percentchange_edited$Treatment_Species, method = "bonferroni")) #%>% 
  #write_csv("changeinRintensity_posthoc_species.csv")

```


### within species

#### Acer x treatment
```{r}
percentchange %>% 
  filter(Species == "Acropora cervicornis") -> Acer_R_percentchange

Acer_Rtreat_model <- lm(percent_change ~ Treatment + Colony + Treatment_Tank, data = Acer_R_percentchange)
Acer_Rtreat_model_metrics <- augment(Acer_Rtreat_model)
plot(Acer_Rtreat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_Rtreat_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment, data = Acer_Rtreat_model_metrics) #significant

# identifying outliers
Acer_Rtreat_model_metrics %>% 
  filter(abs(.std.resid) > 3) #1 outlier

Acer_R_percentchange %>% 
  filter(percent_change > -252) -> Acer_R_percentchange

Acer_Rtreat_model <- lm(percent_change ~ Treatment + Colony + Treatment_Tank, data = Acer_R_percentchange)
Acer_Rtreat_model_metrics <- augment(Acer_Rtreat_model)
plot(Acer_Rtreat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_Rtreat_model_metrics$.resid) #still significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment, data = Acer_Rtreat_model_metrics) #still significant


kruskal_test(percent_change ~ Treatment, data = Acer_R_percentchange)#significant

#posthoc test
as.data.frame(dunn.test(Acer_R_percentchange$percent_change, Acer_R_percentchange$Treatment, method = "bonferroni")) 
```


#### Acer colony X treatment
```{r}
percentchange %>% 
  filter(Species == "Acropora cervicornis") -> Acer_R_percentchange

Acer_Rtreat_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Acer_R_percentchange)
Acer_Rtreat_model_metrics <- augment(Acer_Rtreat_model)
plot(Acer_Rtreat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_Rtreat_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Acer_Rtreat_model_metrics) #significant

# identifying outliers
Acer_Rtreat_model_metrics %>% 
  filter(abs(.std.resid) > 3) #0 outliers

#run non parametric test

#make interaction term
Acer_R_percentchange$Treatment_Colony <- interaction(Acer_R_percentchange$Treatment, Acer_R_percentchange$Colony)

kruskal_test(percent_change ~ Treatment_Colony, data = Acer_R_percentchange)#significant

#posthoc test
Acer_colony_colorscore_posthoc <- as.data.frame(dunn.test(Acer_R_percentchange$percent_change, Acer_R_percentchange$Treatment_Colony, method = "bonferroni")) 

write_csv(Acer_colony_colorscore_posthoc, "Acer_colony_colorscore_posthoc.csv")
```

#### Pcli x treatment
```{r}
percentchange %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_R_percentchange

Pcli_Rtreat_model <- lm(percent_change ~ Treatment + Colony + Treatment_Tank, data = Pcli_R_percentchange)
Pcli_Rtreat_model_metrics <- augment(Pcli_Rtreat_model)
plot(Pcli_Rtreat_model)

# identifying outliers
Pcli_Rtreat_model_metrics %>% 
  filter(abs(.std.resid) > 3) #0 outliers

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Pcli_Rtreat_model_metrics$.resid) #significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment, data = Pcli_Rtreat_model_metrics) #not significant

kruskal_test(percent_change ~ Treatment, data = Pcli_R_percentchange)# not significant
```


#### Pcli colony x treatment
```{r}
percentchange %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_R_percentchange

Pcli_Rtreat_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Pcli_R_percentchange)
Pcli_Rtreat_model_metrics <- augment(Pcli_Rtreat_model)
plot(Pcli_Rtreat_model)

# identifying outliers
Pcli_Rtreat_model_metrics %>% 
  filter(abs(.std.resid) > 3) #0 outliers

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Pcli_Rtreat_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Pcli_Rtreat_model_metrics) #not significant

# two-way anova
summary(aov(percent_change ~ Treatment*Colony, data = Pcli_R_percentchange)) #significant

capture.output(summary(aov(percent_change ~ Treatment*Colony, data = Pcli_R_percentchange)), file = "pclivosa_Rintensity_bycolony_anova.csv")


capture.output(TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Pcli_R_percentchange)), file = "pclivosa_Rintensity_bycolony_posthoc.csv")

as.data.frame(pclivosa_Rintensity_bycolony_posthoc)
  write_csv(., "pclivosa_Rintensity_bycolony_posthoc.csv")

```


```{r}
percentchange %>% 
  dplyr::group_by(Species, Treatment) %>% 
  dplyr::mutate(mean_percent_change = mean(percent_change), se_percent_change = sd(percent_change)) %>% 
  ggplot(., aes(x = Species, y = mean_percent_change))+
  theme_classic()+
  geom_jitter(aes(y=percent_change, color = Treatment, fill = Treatment),
              position=position_dodge(width=0.8),
              alpha=0.2, pch = 21,
              color = "black") +
  geom_errorbar(aes(x = Species, ymax = mean_percent_change+se_percent_change, ymin = mean_percent_change-se_percent_change, color = Treatment), width = .2, position = position_dodge(width=0.8)) +
  #scale_color_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  geom_point(mapping = aes(x = Species, fill = Treatment), size = 2, pch = 21, color = "black", position = position_dodge(width=0.8))+
  #scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  xlab("Day")+
  #facet_wrap(~Species, scales = "free_x", dir = "v") +
  theme(text = element_text(size = 15))
```



## Calculate percent change for the slow-burn heat stress assay
```{r}
Rintensity_withmetadata %>% 
  ungroup() %>% 
  filter(month == "06" | month == "07") %>% 
  dplyr::select(ID, Species, Colony, Treatment, Treatment_Tank, month, mean_Rintensity_percoral) %>% 
  distinct() %>% 
  pivot_wider(names_from = "month", values_from = "mean_Rintensity_percoral") %>% 
  mutate(percent_change = ((`07`-`06`)/`06`)*100) -> slowburn_Rintensity

slowburn_Rintensity %>% 
  drop_na() -> slowburn_Rintensity #~60 corals were not measured in the second time point 
```


```{r}
p1<- ggplot(slowburn_Rintensity,aes(x = Treatment, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB")) +
  theme(legend.position="none")
```

```{r}
p2<-ggplot(slowburn_Rintensity,aes(x = Colony, y= percent_change, fill = Treatment)) + 
  geom_boxplot() + 
  theme_classic() +
  facet_wrap(~Species, scales = "free") +
  labs(y = "Percent Change in R-Intensity") +
  theme(text = element_text(size = 15)) + 
  scale_fill_manual(labels=c("Treated", "Untreated"), values = c("#F54A34", "#60DBDB"))
```

```{r}
plot_grid(p1, p2, ncol = 2, rel_widths = c(1.5,2))

#ggsave("percentchangeRintensity_species_genets_slowburn.pdf", width = 12, height = 5)
```


## Statistics on percent change

### species x treatment

Create linear model to test normality of residuals and homogeneity of variance. Remove outliers
```{r}
str(slowburn_Rintensity) #99 rows

slowburn_Rintensity %>% 
  mutate_at(vars(Colony, Treatment, Species), factor) -> slowburn_Rintensity

treat_model <- lm(percent_change ~ Treatment*Species + Colony + Treatment_Tank, data = slowburn_Rintensity)
treat_model_metrics <- augment(treat_model)
plot(treat_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(treat_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Species, data = treat_model_metrics) #not significant

# anova

summary(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity)) #significant for species

TukeyHSD(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity))

#capture.output(TukeyHSD(aov(percent_change ~ Treatment*Species, data = slowburn_Rintensity)), file = "slowburn_Rintensity_treatmentspecies.csv")
```


### within species colony X treatment

#### Acer
```{r}
slowburn_Rintensity %>% 
  filter(Species == "Acropora cervicornis") -> Acer_slowburn_Rintensity

Acer_R_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Acer_slowburn_Rintensity)
Acer_R_model_model_metrics <- augment(Acer_R_model)
plot(Acer_R_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Acer_R_model_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Acer_R_model_model_metrics) #not significant

# anova

summary(aov(percent_change ~ Treatment*Colony, data = Acer_slowburn_Rintensity)) 

TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Acer_slowburn_Rintensity))
```

#### Pcli
```{r}
slowburn_Rintensity %>% 
  filter(Species == "Pseudodiploria clivosa") -> Pcli_slowburn_Rintensity

Pcli_R_model <- lm(percent_change ~ Treatment*Colony + Treatment_Tank, data = Pcli_slowburn_Rintensity)
Pcli_R_model_metrics <- augment(Pcli_R_model)
plot(Pcli_R_model)

# assess normality of residuals using Shapiro-Wilk test
shapiro_test(Pcli_R_model_metrics$.resid) #not significant

# assess homogeneity of variances using Levene's Test
levene_test(.resid ~ Treatment*Colony, data = Pcli_R_model_metrics) #not significant

# two-way anova
summary(aov(percent_change ~ Treatment*Colony, data = Pcli_slowburn_Rintensity)) #significant

TukeyHSD(aov(percent_change ~ Treatment*Colony, data = Pcli_slowburn_Rintensity))
```



 