---
title: "GO_term_venndiagram"
author: "allyson_demerlis"
date: "2024-02-23"
output: html_document
---

code adapted from [Dr. Hannah Aichelman](https://github.com/hannahaichelman/Oculina_Host_Sym_GE/blob/main/05_Dixon_2020_Comparisons/Dixon_comparison_oculina.R). 

# load libraries and data from Dixon et al 2020
```{r}
### Comparing the GO Delta Ranks to the stress module by Dixon et al. 2020
library(tidyverse)
library(patchwork)
library(cowplot)
library(ggpubr)
library(plotly)

# Read in Dixon's 'type A' or 'type B' cluster for each GO category
# downloaded from here: https://github.com/grovesdixon/Acropora_gene_expression_meta/tree/master/go_mwu
Dixon_BP_typeA = read.table("MWU_BP_clusterAstress_For_MWU.csv", header = T)
Dixon_BP_typeB = read.table("MWU_BP_clusterBstress_For_MWU.csv", header = T)
```

# all BP GO terms


## Variable vs Control
This one has the strongest correlation and provides the most interesting comparison, so I am going to use this data for the manuscript.
```{r}
MWU_result_BP_VariablevControl=read.csv("lpv/MWU_BP_Treated_vs_Untreated_lpv.csv", sep = "")

# Type A 

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
VariablevControl_typeA = intersect(MWU_result_BP_VariablevControl$term, Dixon_BP_typeA$term)
VariablevControl_typeA_GO_BP_set = MWU_result_BP_VariablevControl[MWU_result_BP_VariablevControl$term %in% VariablevControl_typeA,]
Dixon_BP_A_set = Dixon_BP_typeA[Dixon_BP_typeA$term %in% VariablevControl_typeA,]

# Combine them
VariablevControl_BP_typeA_data = merge(VariablevControl_typeA_GO_BP_set, Dixon_BP_A_set, by=c("term", "name"))

p1<-ggplot(VariablevControl_BP_typeA_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type A") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black", se =FALSE) +
  theme_bw() +
  theme(text = element_text(size = 12))

# look at which categories have greatest negative correlation
# can id delta ranks using this plotly:
plotly_typeA = ggplot(VariablevControl_BP_typeA_data, aes(x = delta.rank.x, y = delta.rank.y)) +
  geom_point()
ggplotly(plotly_typeA)

VariablevControl_BP_typeA_data %>% 
  filter(delta.rank.x > 1780 & delta.rank.y < -1840)
# GO:0006271 DNA strand elongation involved in DNA replication

VariablevControl_BP_typeA_data %>% 
  filter(delta.rank.x < -1340 & delta.rank.y > 1810)
# GO:0002474 antigen processing and presentation of peptide antigen via MHC class I


# Type B

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
VariablevControl_typeB = intersect(MWU_result_BP_VariablevControl$term, Dixon_BP_typeB$term)
VariablevControl_typeB_GO_BP_set = MWU_result_BP_VariablevControl[MWU_result_BP_VariablevControl$term %in% VariablevControl_typeB,]
Dixon_BP_B_set = Dixon_BP_typeB[Dixon_BP_typeB$term %in% VariablevControl_typeB,]

# Combine them
VariablevControl_BP_typeB_data = merge(VariablevControl_typeB_GO_BP_set, Dixon_BP_B_set, by=c("term", "name"))

p2<-ggplot(VariablevControl_BP_typeB_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type B") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black", linetype = "dashed", se =FALSE) +
  theme_bw() +
  theme(text = element_text(size = 12))

# look at which categories have greatest negative correlation
# can id delta ranks using this plotly:
plotly_typeB = ggplot(VariablevControl_BP_typeB_data, aes(x = delta.rank.x, y = delta.rank.y)) +
  geom_point()
ggplotly(plotly_typeB)

VariablevControl_BP_typeB_data %>% 
  filter(delta.rank.x > 1780 & delta.rank.y > 1850)
# GO:0006271 DNA strand elongation involved in DNA replication

VariablevControl_BP_typeB_data %>% 
  filter(delta.rank.x < -1540 & delta.rank.y < -2250)
# GO:0070972	protein localization to endoplasmic reticulum
```


```{r}
plot_grid(p1, p2, nrow=1,
       labels = c("A", "B"),       # Add panel labels
  label_size = 14)
ggsave("Dixon2020Comparison.pdf", width = 12, height = 5)
```




## Variable vs Initial
```{r}
MWU_result_BP_VariablevInitial=read.csv("lpv/MWU_BP_Treated_vs_Initial_lpv.csv", sep = "")

# Type A 

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
VariablevInitial_typeA = intersect(MWU_result_BP_VariablevInitial$term, Dixon_BP_typeA$term)
VariablevInitial_typeA_GO_BP_set = MWU_result_BP_VariablevInitial[MWU_result_BP_VariablevInitial$term %in% VariablevInitial_typeA,]
Dixon_BP_A_set = Dixon_BP_typeA[Dixon_BP_typeA$term %in% VariablevInitial_typeA,]

# Combine them
VariablevInitial_BP_typeA_data = merge(VariablevInitial_typeA_GO_BP_set, Dixon_BP_A_set, by=c("term", "name"))

ggplot(VariablevInitial_BP_typeA_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type A") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black", se =FALSE) +
  theme_bw()

# Type B

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
VariablevInitial_typeB = intersect(MWU_result_BP_VariablevInitial$term, Dixon_BP_typeB$term)
VariablevInitial_typeB_GO_BP_set = MWU_result_BP_VariablevInitial[MWU_result_BP_VariablevInitial$term %in% VariablevInitial_typeB,]
Dixon_BP_B_set = Dixon_BP_typeB[Dixon_BP_typeB$term %in% VariablevInitial_typeB,]

# Combine them
VariablevInitial_BP_typeB_data = merge(VariablevInitial_typeB_GO_BP_set, Dixon_BP_B_set, by=c("term", "name"))

ggplot(VariablevInitial_BP_typeB_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type B") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black", linetype = "dashed", se =FALSE) +
  theme_bw()
```


# ARCHIVE - Only significant GO terms for variable corals
This isn't what Aichelman and Wuitchik did, so I am not going to use this. You don't get any useful info from these plots because it's filtering out all the other terms. 

## load lpv datasets
```{r}
# Variable vs. Initial
MWU_result_BP_VariablevInitial=read.csv("lpv/MWU_BP_Treated_vs_Initial_lpv.csv", sep = "")
MWU_result_BP_VariablevInitial %>% 
  filter(p.adj < 0.05) %>% 
  mutate(GO_category = "BP", Contrast = "Variable_vs_Initial") -> MWU_result_BP_VariablevInitial_05

# Variable vs. Control
MWU_result_BP_VariablevControl=read.csv("lpv/MWU_BP_Treated_vs_Untreated_lpv.csv", sep = "")
MWU_result_BP_VariablevControl %>% 
  filter(p.adj < 0.05) %>% 
  mutate(GO_category = "BP", Contrast = "Variable_vs_Control") -> MWU_result_BP_VariablevControl_05

# Control vs. Initial
MWU_result_BP_ControlvInitial=read.csv("lpv/MWU_BP_Untreated_vs_Initial_lpv.csv", sep = "")
MWU_result_BP_ControlvInitial %>% 
  filter(p.adj < 0.05) %>% 
  mutate(GO_category = "BP", Contrast = "Control_vs_Initial") -> MWU_result_BP_ControlvInitial_05
```


## isolate variable GO terms (removing control vs. initial overlaps)
```{r}
# code from venn diagram 

#BP
MWU_result_BP_VariablevInitial_05_name <- MWU_result_BP_VariablevInitial_05$name
MWU_result_BP_VariablevControl_05_name <- MWU_result_BP_VariablevControl_05$name
MWU_result_BP_ControlvInitial_05_name <- MWU_result_BP_ControlvInitial_05$name

pairwise_bp=list("Variable vs. Initial" = MWU_result_BP_VariablevInitial_05_name, "Variable vs. Control" = MWU_result_BP_VariablevControl_05_name, "Control vs. Initial" = MWU_result_BP_ControlvInitial_05_name)

 v.table <- gplots::venn(pairwise_bp)

print(v.table)

    all_intersections <- attr(v.table, "intersections")
    for (comparison in names(all_intersections)) {
      cat(comparison, "\n")
      print(all_intersections[[comparison]])
    }
    
length(all_intersections$`Variable vs. Control`) #60

GO_BP_VvC <- all_intersections$`Variable vs. Control`

length(all_intersections$`Variable vs. Initial`) #237

GO_BP_VvI <- all_intersections$`Variable vs. Initial`

length( all_intersections$`Variable vs. Initial:Variable vs. Control`) #9 

GO_BP_V <- all_intersections$`Variable vs. Initial:Variable vs. Control`

variable_GO_BP <- c(GO_BP_V, GO_BP_VvC, GO_BP_VvI) %>% 
  as.data.frame() %>%  #583 total 
  dplyr::rename(name = ".")

variable_GO_BP_stats <-rbind(MWU_result_BP_VariablevInitial_05, MWU_result_BP_VariablevControl_05, MWU_result_BP_ControlvInitial_05) %>% 
  right_join(., variable_GO_BP) 
```


## Sig BP Variable terms only - c orrelation plots 
```{r}
# Type A 

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
variable_typeA = intersect(variable_GO_BP_stats$term, Dixon_BP_typeA$term)
variable_typeA_GO_BP_set = variable_GO_BP_stats[variable_GO_BP_stats$term %in% variable_typeA,]
Dixon_BP_A_set = Dixon_BP_typeA[Dixon_BP_typeA$term %in% variable_typeA,]

# Combine them
host_BP_typeA_data = merge(variable_typeA_GO_BP_set, Dixon_BP_A_set, by=c("term", "name"))

p1<-ggplot(host_BP_typeA_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,10)) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type A") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black", se =FALSE) +
  theme_bw()

# look for most correlated categories
# can id delta ranks using this plotly:
typeA_plotly = ggplot(host_BP_typeA_data, aes(x = delta.rank.x, y = delta.rank.y)) +
  geom_point()
ggplotly(typeA_plotly)



# Type B

# intersect() makes comparisons are made row-wise, so that in the data frame case, intersect(A,B) is a data frame with those rows that are both in A and in B.
variable_typeB = intersect(variable_GO_BP_stats$term, Dixon_BP_typeB$term)
variable_typeB_GO_BP_set = variable_GO_BP_stats[variable_GO_BP_stats$term %in% variable_typeB,]
Dixon_BP_B_set = Dixon_BP_typeB[Dixon_BP_typeB$term %in% variable_typeB,]

# Combine them
host_BP_typeB_data = merge(variable_typeB_GO_BP_set, Dixon_BP_B_set, by=c("term", "name"))

p2<-ggplot(host_BP_typeB_data, aes(delta.rank.x, delta.rank.y)) +
  scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0,10)) +
  geom_hex() +
  labs( x = "Variable Temp-Treated Acer",
        y = "Dixon Type B") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  geom_smooth(method=lm, color="black",  linetype = "dashed", se =FALSE) +
  theme_bw()

plot_grid(p1, p2)
```




