---
title: "GO_terms_tidying"
author: "allyson_demerlis"
date: "2024-02-22"
output: html_document
---

```{r}
library(tidyverse)
library(DESeq2)
library(cowplot)
```

# BP Treated vs. Untreated
```{r}
bp_TreatedvUntreated <- read_tsv("BP_Treated_vs_Untreated_fc.csv")

bp_TreatedvUntreated %>% 
  rename(lfc = value) %>% 
  rename (level = lev) -> bp_TreatedvUntreated

MWU_result_BP_TreatedvUntreated <- read.csv("MWU_BP_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_BP_TreatedvUntreated %>%
  filter(p.adj <0.05) %>%  #114 GO terms
  mutate(GO_category = "BP") -> MWU_result_BP_TreatedvUntreated_sig

full_join(bp_TreatedvUntreated, MWU_result_BP_TreatedvUntreated_sig, by = c("level", "name")) %>% 
  drop_na(delta.rank) %>% 
  filter(p.adj < 0.05) -> genelists_GO_BP_TreatedvUntreated

# now top 5 upregulated GO terms

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "regulation of DNA-dependent DNA replication initiation")
# Acropora_004501 = regulation of DNA-dependent DNA replication initiation

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "DNA unwinding involved in DNA replication")
# Acropora_015598 = Belongs to the MCM family

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "mitotic DNA replication")
# Acropora_010848 = Tower 

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "positive regulation of chondrocyte differentiation")
# Acropora_006256 = Lysyl oxidase-like 2 

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "regulation of mesonephros development")
# Acropora_026270 = mesenchymal stem cell proliferation

# downregulated GO terms top 5

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "protein exit from endoplasmic reticulum")
# Acropora_022173 = Rhomboid domain containing 1

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "protein folding in endoplasmic reticulum")
#Acropora_001602 = Belongs to the heat shock protein 70 family (this is a repeat)
# Acropora_029829 = heat shock protein 90kDa beta (Grp94), member 1

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "germ-band shortening")
# Acropora_009447 = RAB30, member RAS oncogene family

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "regulation of translational initiation in response to stress")
# Acropora_020594 = positive regulation of translation initiation in response to endoplasmic reticulum stress

genelists_GO_BP_TreatedvUntreated %>% 
  filter(name == "PERK-mediated unfolded protein response")
#Acropora_001602 = Belongs to the heat shock protein 70 family LFC=-1.5537950	
```


```{r}
load("../Rdata_files/vsd.RData")

metadata <- read_csv("../../../physiology/metadata.csv")

metadata %>% 
  dplyr::select(Species, ID, Treatment) %>% 
  filter(Species == "Acropora cervicornis") %>% 
  dplyr::select(!Species) %>% 
  dplyr::rename(Sample_ID = ID) -> metadata

metadata$Formatted <- sprintf("Acer.%03d", metadata$Sample_ID)

metadata %>% 
  drop_na() %>%  
  dplyr::select(!Sample_ID) %>% 
  dplyr::rename(Sample_ID = Formatted) -> metadata
```

## Acropora_004501 = regulation of DNA-dependent DNA replication initiation
```{r}
plot1<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_004501) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_004501, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_004501 = regulation of DNA-dependent DNA replication initiation") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_015598 = Belongs to the MCM family
```{r}
plot2<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_015598) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_015598, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_015598 = Belongs to the MCM family") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_010848 = Tower
```{r}
plot3<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_010848) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_010848, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_010848 = Tower") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_006256 = Lysyl oxidase-like 2 
```{r}
plot4<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_006256) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_006256, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_006256 = Lysyl oxidase-like 2") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_026270 = mesenchymal stem cell proliferation
```{r}
plot5<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_026270) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_026270, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_026270 = mesenchymal stem cell proliferation") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_022173 = Rhomboid domain containing 1 
```{r}
plot6<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_022173) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_022173, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_022173 = Rhomboid domain containing 1 ") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none")
```


## Acropora_029829 = heat shock protein 90kDa beta (Grp94), member 1
```{r}
plot7<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_029829) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_029829, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_029829 = heat shock protein 90kDa beta (Grp94), member 1") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none") 
```


## Acropora_009447 = RAB30, member RAS oncogene family
```{r}
plot8<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_009447) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_009447, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_009447 = RAB30, member RAS oncogene family") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none") 
```



## Acropora_020594 = positive regulation of translation initiation in response to endoplasmic reticulum stress
```{r}
plot9<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_020594) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_020594, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_020594 = Positive regulation of translation initiation in response to endoplasmic reticulum stress") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none") 
```

## Acropora_001602 = Belongs to the heat shock protein 70 family 
```{r}
plot10<-t(vsd) %>% 
  as.data.frame() %>% 
  dplyr::select(Acropora_001602) %>% 
  rownames_to_column() %>% 
  separate(rowname, into = c("Sample_ID", "Genotype", "Expression", "Treatment"), sep ="_") %>% 
  unite("Genotype", Genotype,Expression, sep = "_") %>% 
  mutate(Day = case_when(Treatment == "Initial" ~ 0,
                         Treatment == "Treated" ~ 28,
                         Treatment == "Untreated" ~ 28)) %>% 
  dplyr::select(!Treatment) %>% 
  full_join(., metadata) %>% 
  drop_na() %>% 
  ggplot(., aes(x=Day, y=Acropora_001602, color = Treatment, fill = Treatment)) +
  geom_point(alpha=0.2) +
  theme_classic() +
  geom_smooth(method = "lm", mapping = aes(color = Treatment), se=FALSE) +
  scale_color_manual(values = c( "#FF3333", "#00CCCC"))+
  labs(y = "Vst-Normalized Expression Level", title = "Acropora_001602 = Belongs to the heat shock protein 70 family") +
  theme(text = element_text(size = 13))+
  theme(legend.position="none") 
```

# plot all 10 genes 
```{r}
plot_grid(plot1,plot2,plot3,plot4,plot5,plot6,plot7,plot8,plot9,plot10, ncol=1)
ggsave("attempt1.pdf", width = 3, height= 25)
```



# CC Treated vs. Initial
```{r}
cc_TreatedvInitial <- read.csv("CC_Treated_vs_Initial_fc.csv", sep = "")

cc_TreatedvInitial %>% 
  rename(lfc = value) %>% 
  rename (level = lev) -> cc_TreatedvInitial

MWU_result_CC_TreatedvInitial <- read.csv("MWU_CC_Treated_vs_Initial_fc.csv", sep = "")

full_join(cc_TreatedvInitial, MWU_result_CC_TreatedvInitial, by = c("level", "term", "name")) %>% view()
  filter(p.adj < 0.05) %>% 
  distinct(seq)
  
# these tables are going to be way to big -- CC for Treated v Initial alone is like 44,000 rows...
```

Instead lets just make a table for each dds contrast with significant (p.adj < 0.05) BP, MF, and CC terms in one table

Treated vs Initial
```{r}
MWU_result_CC_TreatedvInitial <- read.csv("MWU_CC_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_CC_TreatedvInitial %>% 
  filter(p.adj <0.05) %>%  #111 GO terms
  mutate(GO_category = "CC") -> MWU_result_CC_TreatedvInitial_sig

MWU_result_MF_TreatedvInitial <- read.csv("MWU_MF_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_MF_TreatedvInitial %>% 
  filter(p.adj <0.05) %>%  #94 GO terms
  mutate(GO_category = "MF") -> MWU_result_MF_TreatedvInitial_sig

MWU_result_BP_TreatedvInitial <- read.csv("MWU_BP_Treated_vs_Initial_fc.csv", sep = "")
MWU_result_BP_TreatedvInitial %>%
  filter(p.adj <0.05) %>%  #543 GO terms
  mutate(GO_category = "BP") -> MWU_result_BP_TreatedvInitial_sig

full_join(MWU_result_CC_TreatedvInitial_sig, MWU_result_MF_TreatedvInitial_sig) %>% 
  full_join(., MWU_result_BP_TreatedvInitial_sig) %>% 
  write_csv("TreatedvsInitial_GOMWU_sigterms.csv")
```


Treated vs Untreated
```{r}
MWU_result_CC_TreatedvUntreated <- read.csv("MWU_CC_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_CC_TreatedvUntreated %>% 
  filter(p.adj <0.05) %>%  #74 GO terms
  mutate(GO_category = "CC") -> MWU_result_CC_TreatedvUntreated_sig

MWU_result_MF_TreatedvUntreated <- read.csv("MWU_MF_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_MF_TreatedvUntreated %>% 
  filter(p.adj <0.05) %>%  #10 GO terms
  mutate(GO_category = "MF") -> MWU_result_MF_TreatedvUntreated_sig

MWU_result_BP_TreatedvUntreated <- read.csv("MWU_BP_Treated_vs_Untreated_fc.csv", sep = "")
MWU_result_BP_TreatedvUntreated %>%
  filter(p.adj <0.05) %>%  #114 GO terms
  mutate(GO_category = "BP") -> MWU_result_BP_TreatedvUntreated_sig

full_join(MWU_result_CC_TreatedvUntreated_sig, MWU_result_MF_TreatedvUntreated_sig) %>% 
  full_join(., MWU_result_BP_TreatedvUntreated_sig) %>% 
  write_csv("TreatedvsUntreated_GOMWU_sigterms.csv")
```

Untreated vs Initial

```{r}
MWU_result_CC_UntreatedvInitial <- read.csv("MWU_CC_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_CC_UntreatedvInitial %>% 
  filter(p.adj <0.05) %>%  #104 GO terms
  mutate(GO_category = "CC") -> MWU_result_CC_UntreatedvInitial_sig

MWU_result_MF_UntreatedvInitial <- read.csv("MWU_MF_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_MF_UntreatedvInitial %>% 
  filter(p.adj <0.05) %>%  #70 GO terms
  mutate(GO_category = "MF") -> MWU_result_MF_UntreatedvInitial_sig

MWU_result_BP_UntreatedvInitial <- read.csv("MWU_BP_Untreated_vs_Initial_fc.csv", sep = "")
MWU_result_BP_UntreatedvInitial %>%
  filter(p.adj <0.05) %>%  #432 GO terms
  mutate(GO_category = "BP") -> MWU_result_BP_UntreatedvInitial_sig

full_join(MWU_result_CC_UntreatedvInitial_sig, MWU_result_MF_UntreatedvInitial_sig) %>% 
  full_join(., MWU_result_BP_UntreatedvInitial_sig) %>% 
  write_csv("UntreatedvsInitial_GOMWU_sigterms.csv")
```

